### 2024-11-26 paper/paper.qmd
explored summary stats for both federal and ontario level donations

---
title: "Impact of Federal Party Leadership Changes on Donation Patterns"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-26 paper/paper.qmd
updated abstract and introduction

@@
 title: "Impact of Federal Party Leadership Changes on Donation Patterns"
-subtitle: "My subtitle if needed"
+subtitle: "An Analysis of Political Contributions in Canada 2013 - 2024"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Maria Mangru
+thanks: "Code and data are available at: [https://github.com/MariaMangru/Ontario_Political_Donors](https://github.com/MariaMangru/Ontario_Political_Donors)."
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study investigates how changes in the ruling political parties within federal and Ontario governments influence individual donor behaviors across Canada.
+By analyzing political donation data from both Ontario and Federal levels between 2013 and 2024, the research assesses whether shifts in which party holds power lead to significant changes in donation amounts.
+Understanding this relationship enhances knowledge of political financing dynamics and informs parties and policymakers about the factors that motivate donor engagement."
 format: pdf
@@
 
-Overview paragraph
+Political donations are an interesting component of democratic processes.
+It provides funding for political parties to campaign and share their platforms.
+In Canada, understanding the factors that influence donation behaviours is important for political parties and policymakers alike.
+One factor that appears to have an impact is changes in the ruling party, which can shape public perception and affect donor support.
 
-Estimand paragraph
+This paper examines how changes in the ruling political party influence individual donation patterns across Canada, focusing on data from Ontario and Federal regions between 2013 and 2024.
+[Maybe talk about other studies here and what they did or didn't do.]
 
-Results paragraph
+To explore this, political contribution data were analyzed, comparing periods before and after changes in which party held power within major federal and Ontario governments.
+Statistical analyses were performed to assess differences in donation amounts during these periods.
 
-Why it matters paragraph
+Understanding the relationship between party control and donation behaviour is valuable for several reasons.
+For political parties, it offers insights into fundraising strategies during times of transition.
+For policymakers and regulators, it informs discussions on campaign finance laws and the transparency of political funding.
+This study also adds to broader discussions about how shifts in party control influence external support, shedding light on the connection between political actors and the electorate.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The goal of this research is to estimate the effect of changes in the ruling political party on the amount of individual political donations in Canada.
+Specifically, it examines whether there is a significant difference in donation amounts before and after a shift in which party holds power within the federal and Ontario governments.
+
+The paper includes a detailed description of the data used, covering sources, variables, and preprocessing steps.
+It outlines the methods used to ensure data accuracy and reliability, followed by an explanation of the statistical model applied and the reasoning behind it.
+The results are presented with supporting tables and graphs.
+Finally, the implications of the findings are discussed, along with limitations and suggestions for future research.
 

### 2024-11-26 scripts/05-exploratory_data_analysis.R

@@ -1,37 +1,250 @@
 #### Preamble ####
 # Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Maria Mangru
+# Date: 25 November 2024
+# Contact: maria.mangru@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
+library(dplyr)
+library(ggplot2)
+library(lubridate)
+library(scales)
+library(readr)   
+library(tidyr)   
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+# Separate data into Ontario and Federal datasets
+ontario_data <- analysis_data %>% filter(region == "Ontario")
+federal_data <- analysis_data %>% filter(region == "Federal")
+
+
+#### Exploring Ontario Data ####
+
+### Summary Statistics ###
+total_donations_ontario <- nrow(ontario_data)
+total_amount_ontario <- sum(ontario_data$amount, na.rm = TRUE)
+avg_donation_ontario <- mean(ontario_data$amount, na.rm = TRUE)
+
+# Save Summary Statistics to a Text File
+summary_stats_ontario <- data.frame(
+  Metric = c("Total Donations", "Total Amount Donated", "Average Donation Amount"),
+  Value = c(total_donations_ontario, total_amount_ontario, avg_donation_ontario)
+)
+write_csv(summary_stats_ontario, "data/02-analysis_data/02.1-exploration_data/Ontario/summary_statistics_ontario.csv")
+
+#### Donations by Year ####
[…]

### 2024-11-28 scripts/03-clean_data.R
renamed political parties

@@ -11,7 +11,7 @@ library(lubridate)
 library(readr)
 
 #### Clean data ####
-raw_data <- read_csv("data/01-raw_data/raw_data.csv")
+raw_data <- read_csv("../data/01-raw_data/raw_data.csv")
 
 
 # Filter for regions 'Federal' or 'Ontario'
@@ -24,6 +24,19 @@ donations_filtered$donation_date <- as.Date(donations_filtered$donation_date)
 
 # Convert donation_year to numeric
 donations_filtered$donation_year <- as.numeric(donations_filtered$donation_year)
+
+
+# Rename political_party entries for Ontario region
+donations_filtered <- donations_filtered %>%
+  mutate(
+    political_party = case_when(
+      region == "Ontario" & political_party %in% c("Ontario Liberal Party", "Liberal Party") ~ "Liberal Party of Ontario",
+      region == "Ontario" & political_party == "New Democratic Party" ~ "New Democratic Party of Ontario",
+      region == "Ontario" & political_party == "Bloc Québécois" ~ "Bloc Quebecois",
+      region == "Federal" & political_party == "Bloc Québécois" ~ "Bloc Quebecois",  # Example for Federal region
+      TRUE ~ political_party
+    )
+  )
 
 # Federal Government Parties in Power from 2011 to 2024
 federal_government <- data.frame(
@@ -57,4 +70,4 @@ donations_filtered <- donations_filtered %>%
 
 
 #### Save data ####
-write_csv(donations_filtered, "data/02-analysis_data/analysis_data.csv")
+write_csv(donations_filtered, "../data/02-analysis_data/analysis_data.csv")

### 2024-11-28 scripts/05-exploratory_data_analysis.R
updated summary graphs and tables

@@ -15,7 +15,7 @@ library(readr)
 library(tidyr)   
 
 #### Read data ####
-analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")
 
 # Separate data into Ontario and Federal datasets
 ontario_data <- analysis_data %>% filter(region == "Ontario")
@@ -34,7 +34,7 @@ summary_stats_ontario <- data.frame(
   Metric = c("Total Donations", "Total Amount Donated", "Average Donation Amount"),
   Value = c(total_donations_ontario, total_amount_ontario, avg_donation_ontario)
 )
-write_csv(summary_stats_ontario, "data/02-analysis_data/02.1-exploration_data/Ontario/summary_statistics_ontario.csv")
+write_csv(summary_stats_ontario, "../data/02-analysis_data/02.1-exploration_data/Ontario/summary_statistics_ontario.csv")
 
 #### Donations by Year ####
 donations_by_year_ontario <- ontario_data %>%
@@ -47,7 +47,7 @@ donations_by_year_ontario <- ontario_data %>%
 
 # Display Donations by Year
 print(donations_by_year_ontario)
-write_csv(donations_by_year_ontario, "data/02-analysis_data/02.1-exploration_data/Ontario/donations_by_year_ontario.csv")
+write_csv(donations_by_year_ontario, "../data/02-analysis_data/02.1-exploration_data/Ontario/donations_by_year_ontario.csv")
 
 ## Donations by Political Party ##
 donations_by_party_ontario <- ontario_data %>%
@@ -60,7 +60,7 @@ donations_by_party_ontario <- ontario_data %>%
 
 # Display Donations by Political Party
 print(donations_by_party_ontario)
-write_csv(donations_by_party_ontario, "data/02-analysis_data/02.1-exploration_data/Ontario/donations_by_political_party_ontario.csv")
+write_csv(donations_by_party_ontario, "../data/02-analysis_data/02.1-exploration_data/Ontario/donations_by_political_party_ontario.csv")
 
 ## Donations Based on Recipient Power Status ##
 donations_in_power_ontario <- ontario_data %>%
@@ -73,7 +73,7 @@ donations_in_power_ontario <- ontario_data %>%
 
 # Display Donations Based on Recipient Power Status
 print(donations_in_power_ontario)
[…]

### 2024-11-30 scripts/03-clean_data.R
update code to indicate which party is a major party

@@ -11,7 +11,7 @@ library(lubridate)
 library(readr)
 
 #### Clean data ####
-raw_data <- read_csv("../data/01-raw_data/raw_data.csv")
+raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 
 
 # Filter for regions 'Federal' or 'Ontario'
@@ -35,6 +35,19 @@ donations_filtered <- donations_filtered %>%
       region == "Ontario" & political_party == "Bloc Québécois" ~ "Bloc Quebecois",
       region == "Federal" & political_party == "Bloc Québécois" ~ "Bloc Quebecois",  # Example for Federal region
       TRUE ~ political_party
+    )
+  )
+
+# Add Party_Size Variable
+donations_filtered <- donations_filtered %>%
+  mutate(
+    Party_Size = case_when(
+      # Major Parties at the Federal Level
+      region == "Federal" & political_party %in% c("Conservative Party of Canada", "Liberal Party of Canada", "New Democratic Party") ~ 1,
+      # Major Parties at the Ontario Level
+      region == "Ontario" & political_party %in% c("Progressive Conservative Party of Ontario", "Liberal Party of Ontario", "New Democratic Party of Ontario") ~ 1,
+      # All Other Parties
+      TRUE ~ 0
     )
   )
 
@@ -70,4 +83,4 @@ donations_filtered <- donations_filtered %>%
 
 
 #### Save data ####
-write_csv(donations_filtered, "../data/02-analysis_data/analysis_data.csv")
+write_csv(donations_filtered, "data/02-analysis_data/analysis_data.csv")

### 2024-11-30 scripts/03-clean_data.R
update model

@@ -77,10 +77,21 @@ donations_filtered <- donations_filtered %>%
       region == "Ontario" ~ ontario_government$party_in_power[match(donation_year, ontario_government$year)],
       TRUE ~ NA_character_
     ),
-    # Create a column indicating if the recipient party was in power
-    recipient_in_power = political_party == party_in_power
+    # Create a binary column indicating if the recipient party was in power
+    recipient_in_power = ifelse(political_party == party_in_power, 1, 0)
   )
 
+donations_filtered <- donations_filtered %>%
+  mutate(
+    Election_Year = case_when(
+      # Federal Election Years
+      region == "Federal" & donation_year %in% c(2011, 2015, 2019, 2021) ~ 1,
+      # Ontario Election Years
+      region == "Ontario" & donation_year %in% c(2011, 2014, 2018, 2022) ~ 1,
+      # All Other Years
+      TRUE ~ 0
+    )
+  )
 
 #### Save data ####
 write_csv(donations_filtered, "data/02-analysis_data/analysis_data.csv")

### 2024-11-30 scripts/06-model_data.R

@@ -11,53 +11,44 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
-# Filter data for Ontario region
-ontario_data <- analysis_data %>% filter(region == "Ontario")
 
-# Define binary variables
-ontario_data <- ontario_data %>%
-  mutate(
-    In_Power = ifelse(recipient_in_power, 1, 0)
-  )
+# Filter data for Ontario
+ontario_data <- donations_filtered %>% filter(region == "Ontario")
 
 # Aggregate total donations per party per year in Ontario
 total_donations_ontario <- ontario_data %>%
   group_by(political_party, donation_year) %>%
   summarize(
     Total_Donations = sum(amount, na.rm = TRUE),
-    In_Power = first(In_Power)
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
   ) %>%
   ungroup()
 
-# Run Regression for Ontario
-model_ontario <- lm(Total_Donations ~ In_Power, data = total_donations_ontario)
+# Run Regression for Ontario with Controls
+model_ontario <- lm(Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
 summary(model_ontario)
 
-
-# Filter data for Federal region
-federal_data <- analysis_data %>% filter(region == "Federal")
-
-# Define binary variables
-federal_data <- federal_data %>%
-  mutate(
-    In_Power = ifelse(recipient_in_power, 1, 0)
-  )
+# Filter data for Federal
+federal_data <- donations_filtered %>% filter(region == "Federal")
 
 # Aggregate total donations per party per year at Federal level
 total_donations_federal <- federal_data %>%
   group_by(political_party, donation_year) %>%
   summarize(
     Total_Donations = sum(amount, na.rm = TRUE),
-    In_Power = first(In_Power)
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
   ) %>%
   ungroup()
 
-# Run Regression for Federal
-model_federal <- lm(Total_Donations ~ In_Power, data = total_donations_federal)
+# Run Regression for Federal with Controls
[…]

### 2024-11-30 scripts/06-model_data.R
modified linear to log model

@@ -32,6 +32,14 @@ total_donations_ontario <- ontario_data %>%
 model_ontario <- lm(Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
 summary(model_ontario)
 
+
+total_donations_ontario <- total_donations_ontario %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
+
+model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
+summary(model_ontario)
+
+
 # Filter data for Federal
 federal_data <- donations_filtered %>% filter(region == "Federal")
 
@@ -51,4 +59,9 @@ model_federal <- lm(Total_Donations ~ In_Power + Party_Size + Election_Year, dat
 summary(model_federal)
 
 
+# For Federal
+total_donations_federal <- total_donations_federal %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))
 
+model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)
+summary(model_federal)

### 2024-11-30 scripts/06-model_data.R
model investigating difference in donation patterns between parties

@@ -65,3 +65,50 @@ total_donations_federal <- total_donations_federal %>%
 
 model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)
 summary(model_federal)
+
+
+
+
+
+### Conservative and Liberal 
+
+# Filter for Conservative and Liberal parties
+cl_data <- analysis_data %>%
+  filter(political_party %in% c(
+    "Conservative Party of Canada",
+    "Liberal Party of Canada",
+    "Progressive Conservative Party of Ontario",
+    "Liberal Party of Ontario"
+  ))
+
+# Create variables
+cl_data <- cl_data %>%
+  mutate(
+    In_Power = recipient_in_power,  # Already binary (1/0)
+    Party = ifelse(
+      political_party %in% c("Conservative Party of Canada", "Progressive Conservative Party of Ontario"),
+      1,  # Conservative
+      0   # Liberal
+    )
+  )
+
+# Aggregate total donations per party per year
+total_donations_cl <- cl_data %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(amount, na.rm = TRUE),
+    In_Power = first(In_Power),
+    Party = first(Party),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Log-transform the dependent variable to stabilize variance
+total_donations_cl <- total_donations_cl %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))
+
+# Run the regression with interaction term
+model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)
+summary(model_cl)
+
+

### 2024-12-03 scripts/06-model_data.R
update model to be log only

@@ -29,9 +29,6 @@ total_donations_ontario <- ontario_data %>%
   ungroup()
 
 # Run Regression for Ontario with Controls
-model_ontario <- lm(Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
-summary(model_ontario)
-
 
 total_donations_ontario <- total_donations_ontario %>%
   mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
@@ -55,10 +52,6 @@ total_donations_federal <- federal_data %>%
   ungroup()
 
 # Run Regression for Federal with Controls
-model_federal <- lm(Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)
-summary(model_federal)
-
-
 # For Federal
 total_donations_federal <- total_donations_federal %>%
   mutate(Log_Total_Donations = log(Total_Donations + 1))

### 2024-12-03 paper/paper.qmd
update abstract

@@
 title: "Impact of Federal Party Leadership Changes on Donation Patterns"
-subtitle: "An Analysis of Political Contributions in Canada 2013 - 2024"
+subtitle: "An Analysis of Political Donations in Canada 2013 - 2024"
 author: 
@@
 date-format: long
-abstract: "This study investigates how changes in the ruling political parties within federal and Ontario governments influence individual donor behaviors across Canada.
-By analyzing political donation data from both Ontario and Federal levels between 2013 and 2024, the research assesses whether shifts in which party holds power lead to significant changes in donation amounts.
-Understanding this relationship enhances knowledge of political financing dynamics and informs parties and policymakers about the factors that motivate donor engagement."
+abstract: "This study investigates how the ruling political parties in Canada's federal and Ontario provincial governments affect the donation behaviors of individual contributors.
+By analyzing donation records from 2013 to 2024, the research explores whether a party being in power influences the total amount of financial support it receives.
+The findings reveal that opposition parties often receive more donations than those in power, highlighting donor preferences to support challengers.
+This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
 format: pdf

### 2024-12-03 paper/paper.qmd
update introduction

--- 2024-11-26 paper/paper.qmd
@@
 title: "Impact of Federal Party Leadership Changes on Donation Patterns"
-subtitle: "My subtitle if needed"
+subtitle: "An Analysis of Political Donations in Canada 2013 - 2024"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+  - Maria Mangru
+thanks: "Code and data are available at: [https://github.com/MariaMangru/Ontario_Political_Donors](https://github.com/MariaMangru/Ontario_Political_Donors)."
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study investigates how the ruling political parties in Canada's federal and Ontario provincial governments affect the donation behaviors of individual contributors.
+By analyzing donation records from 2013 to 2024, the research explores whether a party being in power influences the total amount of financial support it receives.
+The findings reveal that opposition parties often receive more donations than those in power, highlighting donor preferences to support challengers.
+This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
 format: pdf
@@
 
-Overview paragraph
+Political donations play a critical role in supporting political campaigns and shaping electoral competition.
+In Canada, both federal and provincial elections rely heavily on contributions from individuals and organizations, which influence the political landscape and often reflect public support for different parties.
+Understanding what drives donation patterns is essential for ensuring transparency and fairness in the democratic process and for minimizing the risk of undue influence on policy-making.
 
-Estimand paragraph
+One key question is whether a party’s position in power affects the amount of financial support it receives.
+While studies in other countries, such as the U.S.A, have shown that incumbents often benefit from more donations, particularly from interest groups looking to maintain influence, there has been less focus on whether this holds true in Canada.
+Research like Fouirnaies and Hall’s work (REFERENCE HERE) in the U.S.
+shows that donors strategically target those in power, but how these dynamics play out in Canadian elections remains under explored.
+This paper aims to fill that gap by analyzing how a party's power status influences donor behaviour in federal and Ontario provincial elections.
 
-Results paragraph
+Using donation data from 2013 to 2024, this study compares the financial support received by parties in power versus those in opposition.
+The analysis examines total donations by year and party while accounting for factors like election timing and party size (major party or not).
+By identifying trends in how donors respond to shifts in political power, this research sheds light on the financial strategies behind Canada’s democratic processes.
 
-Why it matters paragraph
+The results show that opposition parties tend to receive higher donations than those in power, suggesting that donors often prioritize supporting challengers, possibly to influence political change.
+These findings align with trends observed in other political systems but highlight unique patterns in Canada.
+This information is valuable not only for political parties developing fundraising strategies but also for policymakers and researchers seeking to ensure fairness and accountability in political financing.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The paper is organized as follows: Section \ref{sec-data} discusses the data sources and methodology, followed by an explanation of the variables used in the analysis.
+The regression models and results are then presented, along with interpretations.
+The paper concludes by discussing the broader implications of the findings and suggesting directions for future research.[NEED TO UPDATE THE REFERENCES]
 

### 2024-12-03 paper/paper.qmd
update data section of paper

@@
 bibliography: references.bib
+editor: 
+  markdown: 
+    wrap: 72
 ---
@@
 
-Political donations play a critical role in supporting political campaigns and shaping electoral competition.
-In Canada, both federal and provincial elections rely heavily on contributions from individuals and organizations, which influence the political landscape and often reflect public support for different parties.
-Understanding what drives donation patterns is essential for ensuring transparency and fairness in the democratic process and for minimizing the risk of undue influence on policy-making.
+Political donations play a critical role in supporting political
+campaigns and shaping electoral competition. In Canada, both federal and
+provincial elections rely heavily on contributions from individuals and
+organizations, which influence the political landscape and often reflect
+public support for different parties. Understanding what drives donation
+patterns is essential for ensuring transparency and fairness in the
+democratic process and for minimizing the risk of undue influence on
+policy-making.
 
-One key question is whether a party’s position in power affects the amount of financial support it receives.
-While studies in other countries, such as the U.S.A, have shown that incumbents often benefit from more donations, particularly from interest groups looking to maintain influence, there has been less focus on whether this holds true in Canada.
-Research like Fouirnaies and Hall’s work (REFERENCE HERE) in the U.S.
-shows that donors strategically target those in power, but how these dynamics play out in Canadian elections remains under explored.
-This paper aims to fill that gap by analyzing how a party's power status influences donor behaviour in federal and Ontario provincial elections.
+One key question is whether a party’s position in power affects the
+amount of financial support it receives. While studies in other
+countries, such as the U.S.A, have shown that incumbents often benefit
+from more donations, particularly from interest groups looking to
+maintain influence, there has been less focus on whether this holds true
+in Canada. Research like Fouirnaies and Hall’s work (REFERENCE HERE) in
+the U.S. shows that donors strategically target those in power, but how
+these dynamics play out in Canadian elections remains under explored.
+This paper aims to fill that gap by analyzing how a party's power status
+influences donor behaviour in federal and Ontario provincial elections.
 
-Using donation data from 2013 to 2024, this study compares the financial support received by parties in power versus those in opposition.
-The analysis examines total donations by year and party while accounting for factors like election timing and party size (major party or not).
-By identifying trends in how donors respond to shifts in political power, this research sheds light on the financial strategies behind Canada’s democratic processes.
+Using donation data from 2013 to 2024, this study compares the financial
+support received by parties in power versus those in opposition. The
+analysis examines total donations by year and party while accounting for
+factors like election timing and party size (major party or not). By
+identifying trends in how donors respond to shifts in political power,
+this research sheds light on the financial strategies behind Canada’s
+democratic processes.
 
-The results show that opposition parties tend to receive higher donations than those in power, suggesting that donors often prioritize supporting challengers, possibly to influence political change.
-These findings align with trends observed in other political systems but highlight unique patterns in Canada.
-This information is valuable not only for political parties developing fundraising strategies but also for policymakers and researchers seeking to ensure fairness and accountability in political financing.
+The results show that opposition parties tend to receive higher
+donations than those in power, suggesting that donors often prioritize
+supporting challengers, possibly to influence political change. These
+findings align with trends observed in other political systems but
+highlight unique patterns in Canada. This information is valuable not
+only for political parties developing fundraising strategies but also
+for policymakers and researchers seeking to ensure fairness and
+accountability in political financing.
 
-The paper is organized as follows: Section \ref{sec-data} discusses the data sources and methodology, followed by an explanation of the variables used in the analysis.
-The regression models and results are then presented, along with interpretations.
-The paper concludes by discussing the broader implications of the findings and suggesting directions for future research.[NEED TO UPDATE THE REFERENCES]
+The paper is organized as follows: Section \ref{sec-data} discusses the
+data sources and methodology, followed by an explanation of the
+variables used in the analysis. The regression models and results are
+then presented, along with interpretations. The paper concludes by
+discussing the broader implications of the findings and suggesting
+directions for future research.\[NEED TO UPDATE THE REFERENCES\]
 
@@
 
+## Data Source
+
+# Data
+
 ## Overview
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+The dataset used in this analysis comprises of political donation
+records spanning from 2013 to 2024, covering both the federal level and
+the province of Ontario in Canada. These records were sourced from the
+Investigative Journalism Foundation (IJF), which systematically collects
+and maintains comprehensive data on political donations across various
+Canadian jurisdictions. The IJF gathers daily updates by monitoring
+elections agency websites at the federal level and within each province
+and territory, ensuring the dataset remains current and robust. Although
+the IJF’s repository includes historical data dating back to 1993, this
+study focuses exclusively on the 2013-2024 period to capture recent
+trends and changes in political financing dynamics. \[Inlude references
+here for IJF, R and the packages used.\]
 
-Overview text
+Political donations in Canada are subject to legal disclosure
+requirements, promoting transparency and accountability within the
+democratic process. Both federal and provincial election finance laws
+mandate that political parties, candidates, and associated entities
+report donations exceeding specific thresholds. These disclosures vary
+by jurisdiction in terms of frequency and reporting formats. For
+instance, Ontario requires annual returns for general and by-elections,
+leadership races, and donations to parties, candidates, and constituency
+associations, all available in downloadable spreadsheet formats from
+2007 onward. This regulatory framework ensures that significant
+contributions are systematically documented and publicly accessible,
+providing a reliable foundation for this analysis.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The primary objective of this analysis is to investigate the impact of a
+political party’s power status on donation patterns within Canadian
+federal and Ontario provincial elections. The dataset includes detailed
+records of donations made to various recipients, including political
+parties, party leadership contestants, riding associations (also known
+as electoral district or constituency associations), and individual
+riding candidates. Donations are tracked across different types of
+electoral events, such as general elections, by-elections, and
+leadership races.
 
-## Outcome variables
+Key variables within the dataset are as follows:
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+-   **Political Party:** Categorical variable indicating the recipient
+    political party, encompassing both major parties (e.g., Liberal
+    Party, Conservative Party, New Democratic Party) and minor parties.
+-   **Donation Year:** Numerical variable representing the year the
+    donation was made, ranging from 2013 to 2024. This captures
+    different political climates and election periods.
+-   **Amount Donated:** Continuous variable denoting the monetary value
+    of each donation, varying widely due to differing donor capacities
+    and legal contribution limits.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+### Constructed Variables
+
+In addition to the original variables, several constructed variables
+were created to facilitate a more nuanced analysis:
+
+-   **Recipient_in_Power:** Binary variable (1/0) indicating whether the
+    recipient party was in power during the donation year.
+-   **Party_Size:** Binary variable (1/0) categorizing parties as
+    major (1) or minor (0), based on their representation in the
+    legislature.
+-   **Election_Year:** Binary variable (1/0) indicating whether the
+    donation was made during an election year.
+-   **Log_Total_Donations:** Logarithm of the total donations received
+    by a party in a given year, calculated as `log(Total_Donations + 1)`
+    to handle zero values.
+-   **Interaction Terms:** Interaction terms such as
+    `Recipient_in_Power * Party_Size` to explore how the effect of being
+    in power varies between major and minor parties.
+
+## Data Cleaning
+
+The raw donation records underwent a series of preprocessing steps to
+ensure the dataset's integrity and suitability for analysis. The key
+data cleaning procedures included:
+
+1.  **Handling Missing Values:** Missing or incomplete records were
+    addressed by either imputing values based on available information
+    or excluding incomplete entries to maintain data integrity.
+2.  **Standardizing Party Names:** Political party names were
+    standardized across jurisdictions to ensure consistency. For
+    example, variations like “Progressive Conservative,” “Progressive
+    Conservative Party of Ontario,” and “The Progressive Conservative
+    Association of Ontario” were consolidated into “Progressive
+    Conservative Association of Ontario.”
+3.  **Removing Outliers:** Extreme donation amounts that could skew the
+    analysis were identified and removed based on predefined criteria.
+4.  **Aggregating Donations:** Donations were aggregated to reflect
+    total amounts per party per year, facilitating a higher-level
+    analysis of donation patterns.
+5.  **Creating Binary Variables:** Binary indicators were generated for
+    key variables such as `Recipient_in_Power`, `Party_Size`, and
+    `Election_Year` to streamline regression analysis.
+
+Additionally, data from jurisdictions with incomplete historical records
+or inconsistent reporting formats were excluded to maintain the
+dataset's uniformity. Detailed data cleaning steps, including the
+handling of OCR-converted records and manual verification processes, are
+elaborated in [Appendix A: Data Cleaning Notes](#sec-appx).
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
+#| include: false
 #| warning: false
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+# Load necessary libraries
+library(tidyverse)
+library(janitor)
+library(knitr)
+library(kableExtra)
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# Read in the cleaned donation data
+analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")
 ```
 
-Talk way more about it.
+## Outcome Variables
 
-## Predictor variables
+The main outcome variable of interest in this analysis is the **Amount
+Donated**, which represents the monetary value of each donation made to
+a political party within a given year. This continuous variable is
+crucial for assessing the financial support received by parties based on
+their power status and other influencing factors. Understanding donation
+amounts helps in evaluating the extent to which financial backing
+correlates with political influence and success.
 
-Add graphs, tables and text.
+```{r}
+#| label: tbl-summary-ontario
+#| tbl-cap: "Summary Statistics for Ontario Donations"
+#| echo: false
+#| message: false
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+#### Donations by Year ####
+donations_by_year_ontario <- ontario_data %>%
+  group_by(donation_year) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
+
+kable(donations_by_year_ontario, caption = "Summary Statistics for Ontario Donations")
+```
+
+[NEED TO INCLUDE OTHER TABLES HERE FOR ONTARIO AND ALSO FEDERAL]
+
+## Predictor Variables
+
+Several key predictor variables were identified to evaluate the factors
+influencing donation amounts effectively:
+
+-   **Recipient_in_Power:** Indicates whether the recipient party was in
+    power during the donation year (1) or not (0). This variable is
+    pivotal in assessing whether being in power attracts more financial
+    support.
+-   **Party_Size:** Categorizes parties as major (1) or minor (0) based
+    on their representation in the legislature. Major parties typically
+    have more visibility and influence, potentially attracting more
+    donations.
+-   **Election_Year:** Indicates whether the donation was made during an
+    election year (1) or not (0). Election cycles often see increased
+    financial activity as parties ramp up fundraising efforts.
+
+### Additional Predictor Variables
+
+-   **Log_Total_Donations:** The logarithm of total donations received
+    by a party in a given year, calculated to normalize the distribution
+    and handle skewness in donation amounts.
+-   **Interaction Terms:** Terms such as
+    `Recipient_in_Power * Party_Size` to explore the combined effect of
+    power status and party size on donation amounts.
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Create log_total_donations
+donations_cleaned <- donations_cleaned %>%
+  group_by(party, year) %>%
+  mutate(total_donations = sum(amount_donated, na.rm = TRUE)) %>%
+  ungroup() %>%
+  mutate(log_total_donations = log(total_donations + 1))
+```
+
+## Summary Statistics
+
+### Ontario Data
+
+#### Total Donations and Amount Donated
+
+[TABLE NEEDED HERE]
+
+#### Donations by Year
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Aggregate donations by year for Ontario
+donations_by_year_ontario <- ontario_data %>%
+  group_by(donation_year) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
+```
+
+```{r}
+#| label: tbl-donations-year-ontario
+#| tbl-cap: "Donations by Year in Ontario"
+#| echo: false
+#| message: false
+
+kable(donations_by_year_ontario, caption = "Donations by Year in Ontario")
+```
+
+#### Donations by Political Party
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Aggregate donations by party for Ontario
+donations_in_power_ontario <- ontario_data %>%
+  group_by(recipient_in_power) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
+```
+
+```{r}
+#| label: tbl-donations-party-ontario
+#| tbl-cap: "Donations by Political Party in Ontario"
+#| echo: false
+#| message: false
+
+kable(donations_in_power_ontario, caption = "Donations by Political Party in Ontario")
+```
+
+#### Donations Based on Recipient Power Status
+
+[TABLE NEEDED HERE]
+
+### Federal Data
+
+#### Total Donations and Amount Donated
+
+[TABLE NEEDED HERE]
+
+#### Donations by Year
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Aggregate donations by year for Federal
+donations_by_year_federal <- federal_data %>%
+  group_by(donation_year) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
+
+```
+
+```{r}
+#| label: tbl-donations-year-federal
+#| tbl-cap: "Donations by Year at Federal Level"
+#| echo: false
+#| message: false
+
+kable(donations_by_year_federal, caption = "Donations by Year at Federal Level")
+```
+
+#### Donations by Political Party
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Aggregate donations by party for Federal
+donations_by_party_federal <- federal_data %>%
+  group_by(political_party) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
+
+```
+
+```{r}
+#| label: tbl-donations-party-federal
+#| tbl-cap: "Donations by Political Party at Federal Level"
+#| echo: false
+#| message: false
+
+kable(donations_by_party_federal, caption = "Donations by Political Party at Federal Level")
+```
+
+#### Donations Based on Recipient Power Status
+
+[TABLE NEEDED HERE]
+
+## Relationships Between Variables
+
+Preliminary analysis reveals that political parties not in power receive
+higher total donations on average compared to those in power, both at
+the federal and provincial levels. This consistent trend suggests that
+donors may strategically allocate their contributions to opposition
+parties in hopes of influencing political change and promoting
+accountability. Additionally, major parties consistently attract more
+donations than minor parties, underscoring the importance of party
+prominence and visibility in fundraising efforts.
+
+### Donation Amounts by Recipient Power Status
+
+```{r}
+#| label: fig-recipient-power-distribution-ontario
+#| fig-cap: "Donation Amounts by Recipient Power Status in Ontario"
+#| echo: false
+#| message: false
+
+p4_ontario <- ggplot(donations_in_power_ontario, aes(x = recipient_in_power, y = total_amount, fill = recipient_in_power)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Donations Based on Party Power Status - Ontario",
+       x = "Recipient Party in Power", y = "Total Amount Donated ($)") +
+  scale_x_discrete(labels = c("FALSE" = "Not in Power", "TRUE" = "In Power")) +
+  scale_y_continuous() +
+  theme_minimal() +
+  theme(legend.position = "none",
+        axis.text.x = element_text(angle = 0, hjust = 0.5))
+```
+
+### Donation Amounts by Political Party
+
+#### Ontario
+
+```{r}
+#| label: fig-donations-party-ontario
+#| fig-cap: "Donations by Political Party in Ontario"
+#| echo: false
+#| message: false
+
+p3_ontario <- ggplot(donations_by_party_ontario, aes(x = reorder(political_party, -total_amount), y = total_amount, fill = political_party)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Total Donations by Political Party - Ontario",
+       x = "Political Party", y = "Total Amount Donated ($)") +
+  scale_y_continuous() +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "none")
+```
+
+#### Federal
+
+```{r}
+#| label: fig-donations-party-federal
+#| fig-cap: "Donations by Political Party at Federal Level"
+#| echo: false
+#| message: false
+
+donations_by_party_federal %>%
+  ggplot(aes(x = reorder(party, -Total_Amount_Donated), y = Total_Amount_Donated, fill = party)) +
+  geom_bar(stat = "identity") +
+  theme_minimal() +
+  labs(
+    title = "Donations by Political Party at Federal Level",
+    x = "Political Party",
+    y = "Total Amount Donated ($)",
+    fill = "Political Party"
+  ) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+## Relationships Between Variables
+
+The analysis indicates that political parties not in power receive
+higher total donations on average compared to those in power, both at
+the federal and provincial levels. This trend suggests strategic donor
+behavior aimed at supporting challengers to foster political change and
+accountability. Additionally, major parties attract significantly more
+donations than minor parties, underscoring the importance of party
+prominence in fundraising efforts.
 
@@
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Here we briefly describe the Bayesian analysis model used to
+investigate... Background details and diagnostics are included in
+[Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the number of seconds that the plane remained aloft.
+Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both
+measured in millimeters.
 
+```{=tex}
 \begin{align} 
@@
 \end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+```
+We run the model in R [@citeR] using the `rstanarm` package of
+@rstanarm. We use the default priors from `rstanarm`.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+We expect a positive relationship between the size of the wings and time
+spent aloft. In particular...
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+We can use maths by including latex between dollar signs, for instance
+$\theta$.
 
@@
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+If my paper were 10 pages, then should be be at least 2.5 pages. The
+discussion is a chance to show off what you know and what you learnt
+from all this.
 
@@
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Please don't use these as sub-heading labels - change them to be what
+your point actually is.
 
@@
 
-# Appendix {-}
+# Appendix {#sec-appx}
+
+## Appendix A: Data Cleaning Notes
+
+The data cleaning process was meticulously carried out to ensure the
+dataset's reliability and consistency. The key steps included:
+
+1.  **Importing Raw Data:**
+    -   The raw donation data was imported using the `read_csv` function
+        from the `tidyverse` package.
+    -   Data from PDF sources were converted to CSV using OCR tools such
+        as Adobe Export PDF and ABBYY FineReader, followed by manual
+        verification to correct OCR-induced errors.
+2.  **Handling Missing Values:**
+    -   Missing values in critical variables (`amount_donated`,
+        `recipient_in_power`, `party_size`, `election_year`) were
+        addressed by excluding incomplete records to maintain data
+        integrity.
+3.  **Standardizing Party Names:**
+    -   Political party names were standardized across all
+        jurisdictions. For example, variations like “Progressive
+        Conservative,” “Progressive Conservative Party of Ontario,” and
+        “The Progressive Conservative Association of Ontario” were
+        consolidated into “Progressive Conservative Association of
+        Ontario.”
+4.  **Removing Outliers:**
+    -   Donation amounts exceeding three standard deviations from the
+        mean were identified and removed to prevent skewing the
+        analysis.
+5.  **Aggregating Donations:**
+    -   Donations were aggregated at the party-year level, resulting in
+        a dataset where each row represents the total donations received
+        by a party in a given year.
+6.  **Creating Binary Variables:**
+    -   Binary indicators were created for `recipient_in_power`,
+        `party_size`, and `election_year` to facilitate regression
+        analysis.
+7.  **Log Transformation:**
+    -   The `log_total_donations` variable was created by taking the
+        logarithm of `Total_Donations + 1` to normalize the distribution
+        and handle zero values.
+8.  **Final Verification:**
+    -   The cleaned dataset was cross-validated against original records
+        to ensure accuracy and consistency.
+
+# Appendix {.unnumbered}
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive
+check. This shows...
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the
+prior. This shows...
 

### 2024-12-03 paper/paper.qmd
updated data section to add tables and graphs

--- 2024-12-03 paper/paper.qmd
@@
 
-Political donations play a critical role in supporting political campaigns and shaping electoral competition.
-In Canada, both federal and provincial elections rely heavily on contributions from individuals and organizations, which influence the political landscape and often reflect public support for different parties.
-Understanding what drives donation patterns is essential for ensuring transparency and fairness in the democratic process and for minimizing the risk of undue influence on policy-making.
+Political donations play a critical role in supporting political
+campaigns and shaping electoral competition. In Canada, both federal and
+provincial elections rely heavily on contributions from individuals and
+organizations, which influence the political landscape and often reflect
+public support for different parties. Understanding what drives donation
+patterns is essential for ensuring transparency and fairness in the
+democratic process and for minimizing the risk of undue influence on
+policy-making.
 
-One key question is whether a party’s position in power affects the amount of financial support it receives.
-While studies in other countries, such as the U.S.A, have shown that incumbents often benefit from more donations, particularly from interest groups looking to maintain influence, there has been less focus on whether this holds true in Canada.
-Research like Fouirnaies and Hall’s work (REFERENCE HERE) in the U.S.
-shows that donors strategically target those in power, but how these dynamics play out in Canadian elections remains under explored.
-This paper aims to fill that gap by analyzing how a party's power status influences donor behaviour in federal and Ontario provincial elections.
+One key question is whether a party’s position in power affects the
+amount of financial support it receives. While studies in other
+countries, such as the U.S.A, have shown that incumbents often benefit
+from more donations, particularly from interest groups looking to
+maintain influence, there has been less focus on whether this holds true
+in Canada. Research like Fouirnaies and Hall’s work (REFERENCE HERE) in
+the U.S. shows that donors strategically target those in power, but how
+these dynamics play out in Canadian elections remains under explored.
+This paper aims to fill that gap by analyzing how a party's power status
+influences donor behaviour in federal and Ontario provincial elections.
 
-Using donation data from 2013 to 2024, this study compares the financial support received by parties in power versus those in opposition.
-The analysis examines total donations by year and party while accounting for factors like election timing and party size (major party or not).
-By identifying trends in how donors respond to shifts in political power, this research sheds light on the financial strategies behind Canada’s democratic processes.
+Using donation data from 2013 to 2024, this study compares the financial
+support received by parties in power versus those in opposition. The
+analysis examines total donations by year and party while accounting for
+factors like election timing and party size (major party or not). By
+identifying trends in how donors respond to shifts in political power,
+this research sheds light on the financial strategies behind Canada’s
+democratic processes.
 
-The results show that opposition parties tend to receive higher donations than those in power, suggesting that donors often prioritize supporting challengers, possibly to influence political change.
-These findings align with trends observed in other political systems but highlight unique patterns in Canada.
-This information is valuable not only for political parties developing fundraising strategies but also for policymakers and researchers seeking to ensure fairness and accountability in political financing.
+The results show that opposition parties tend to receive higher
+donations than those in power, suggesting that donors often prioritize
+supporting challengers, possibly to influence political change. These
+findings align with trends observed in other political systems but
+highlight unique patterns in Canada. This information is valuable not
+only for political parties developing fundraising strategies but also
+for policymakers and researchers seeking to ensure fairness and
+accountability in political financing.
 
-The paper is organized as follows: Section \ref{sec-data} discusses the data sources and methodology, followed by an explanation of the variables used in the analysis.
-The regression models and results are then presented, along with interpretations.
-The paper concludes by discussing the broader implications of the findings and suggesting directions for future research.[NEED TO UPDATE THE REFERENCES]
+The paper is organized as follows: Section \ref{sec-data} discusses the
+data sources and methodology, followed by an explanation of the
+variables used in the analysis. The regression models and results are
+then presented, along with interpretations. The paper concludes by
+discussing the broader implications of the findings and suggesting
+directions for future research.\[NEED TO UPDATE THE REFERENCES\]
 
-# Data {#sec-data}
+# Data
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+#### Workspace setup ####
+library(dplyr)
+library(ggplot2)
+library(lubridate)
+library(scales)
+library(readr)   
+library(tidyr)
+
+#### Read data ####
+analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")
+
+# Separate data into Ontario and Federal datasets
+ontario_data <- analysis_data %>% filter(region == "Ontario")
+federal_data <- analysis_data %>% filter(region == "Federal")
+```
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+The dataset used in this analysis comprises political donation records spanning from 2013 to 2024, covering both the federal level and the province of Ontario in Canada.
+These records were sourced from the Investigative Journalism Foundation (IJF), which systematically collects and maintains comprehensive data on political donations across various Canadian jurisdictions.
+The IJF gathers daily updates by monitoring elections agency websites at the federal level and within each province and territory, ensuring the dataset remains current and robust.
+Although the IJF's repository includes historical data dating back to 1993, this study focuses exclusively on the 2013–2024 period to capture recent trends and changes in political financing dynamics (Investigative Journalism Foundation 2024).
 
-Overview text
+The analysis was conducted using the statistical programming language R (R Core Team 2024) and several packages for data manipulation and visualization, including `dplyr` (Wickham et al.
+2023), `ggplot2` (Wickham 2016), `readr` (Wickham et al.
+2023), and `tidyr` (Wickham et al.
+2023).
+
+Political donations in Canada are subject to legal disclosure requirements, promoting transparency and accountability within the democratic process.
+Both federal and provincial election finance laws mandate that political parties, candidates, and associated entities report donations exceeding specific thresholds.
+These disclosures vary by jurisdiction in terms of frequency and reporting formats.
+For instance, Ontario requires annual returns for general and by-elections, leadership races, and donations to parties, candidates, and constituency associations, all available in downloadable spreadsheet formats from 2007 onward.
+This regulatory framework ensures that significant contributions are systematically documented and publicly accessible, providing a reliable foundation for this analysis.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The primary objective of this analysis is to investigate the impact of a political party's power status on donation patterns within Canadian federal and Ontario provincial elections.
+The dataset includes detailed records of donations made to various recipients, such as political parties, party leadership contestants, riding associations (also known as electoral district or constituency associations), and individual riding candidates.
+Donations are tracked across different types of electoral events, including general elections, by-elections, and leadership races.
 
-## Outcome variables
+Key variables within the dataset are as follows:
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+- **Political Party:** A categorical variable indicating the recipient political party, encompassing both major parties (e.g., Liberal Party, Conservative Party, New Democratic Party) and minor parties.
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+- **Donation Year:** A numerical variable representing the year the donation was made, ranging from 2013 to 2024. This captures different political climates and election periods.
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
+- **Amount Donated:** A continuous variable denoting the monetary value of each donation, varying widely due to differing donor capacities and legal contribution limits.
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+### Constructed Variables
+
+Several variables were constructed to facilitate a more nuanced analysis:
+
+- **Recipient_in_Power:** A binary variable indicating whether the recipient party was in power during the donation year (1 for in power, 0 for not in power).
+This was determined by cross-referencing the donation year with historical records of party leadership at both the federal and provincial levels.
+
+- **Party_Size:** A binary variable categorizing parties as major (1) or minor (0), based on their representation in the legislature.
+Major parties are those with significant representation or historical influence.
+
+- **Election_Year:** A binary variable indicating whether the donation was made during an election year (1) or not (0), based on official election dates at the federal and provincial levels.
+
+## Data Cleaning
+
+The raw donation records underwent several preprocessing steps to ensure the dataset's integrity and suitability for analysis:
+
+1.
+**Handling Missing Values:** Records with missing or incomplete information, particularly in key variables such as donation amount or recipient party, were excluded from the analysis to maintain data quality.
+
+2.
+**Standardizing Party Names:** Political party names were standardized to ensure consistency.
+For example, variations like "Progressive Conservative Party of Ontario" and "PC Party of Ontario" were consolidated under a single standardized name.
+
+3. **Removing Outliers:** Extreme donation amounts that could skew the analysis were identified and removed based on predefined thresholds, acknowledging the legal donation limits in Canada.
+
+4. **Aggregating Donations:** Donations were aggregated to reflect total amounts per party per year, facilitating a higher-level analysis of donation patterns.
+
+5. **Creating Binary Variables:** Binary indicators were generated for key variables such as `Recipient_in_Power`, `Party_Size`, and `Election_Year` to streamline regression analysis.
+
+Additional details on the data cleaning process, including code snippets and specific handling of anomalies, are provided in Appendix A.
+
+## Outcome Variables
+
+The main outcome variable of interest is the **Amount Donated**, representing the monetary value of each donation made to a political party within a given year.
+This continuous variable is critical for assessing the financial support received by parties and understanding how it correlates with factors like power status and party size.
+
+The distribution of donation amounts varies between the federal and provincial levels.
+At the federal level, the average donation amount is higher, reflecting different legal contribution limits and donor behaviors.
+
+## Predictor Variables
+
+Several key predictor variables were identified to evaluate factors influencing donation amounts:
+
+- **Recipient_in_Power:** Indicates whether the recipient party was in power during the donation year (1) or not (0).
+This variable is central to assessing the impact of political power on donation patterns.
+
+- **Party_Size:** Categorizes parties as major (1) or minor (0). Major parties are expected to receive more donations due to greater visibility and established support bases.
+
+- **Election_Year:** Indicates whether the donation was made during an election year (1) or not (0). Election years often see increased fundraising activities.
+
+### Additional Predictor Variables
+
+- **Log_Total_Donations:** The logarithm of total donations received by a party in a given year, calculated as `log(Total_Donations + 1)` to handle zero values and normalize the distribution.
+
+- **Interaction Terms:** Interaction terms such as `Recipient_in_Power * Party_Size` were included in the models to explore how the effect of being in power varies between major and minor parties.
+
+## Summary Statistics
+
+### Donations Based on Recipient Power Status
+
+#### Ontario
+
+Table 1 presents the total donations and average donation amounts in Ontario, categorized by whether the recipient party was in power.
+
+**Table 1: Donations by Power Status in Ontario**
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+## Donations Based on Recipient Power Status ##
+donations_in_power_ontario <- ontario_data %>%
+  group_by(recipient_in_power) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
 ```
 
-Talk more about it.
+Figure 1 illustrates the total donations and average donation amounts over the years in Ontario.
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+# Plotting Total and Average Donations Over Years - Ontario
+donations_by_year_ontario <- ontario_data %>%
+  group_by(donation_year) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
+library(ggplot2)
+library(gridExtra)
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+p1_ontario <- ggplot(donations_by_year_ontario, aes(x = donation_year, y = total_amount)) +
+  geom_line(color = "steelblue") +
+  geom_point(color = "steelblue") +
+  labs(title = "Total Donations Over Years - Ontario",
+       x = "Year", y = "Total Amount Donated ($)") +
+  scale_y_continuous(labels = scales::dollar_format()) +
+  theme_minimal()
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+p2_ontario <- ggplot(donations_by_year_ontario, aes(x = donation_year, y = avg_amount)) +
+  geom_line(color = "darkgreen") +
+  geom_point(color = "darkgreen") +
+  labs(title = "Average Donation Amount Over Years - Ontario",
+       x = "Year", y = "Average Donation Amount ($)") +
+  scale_y_continuous(labels = scales::dollar_format()) +
+  theme_minimal()
+
+grid.arrange(p1_ontario, p2_ontario, ncol = 2)
 ```
 
-Talk way more about it.
+*Figure 1: Total and Average Donations Over Years in Ontario*
 
-## Predictor variables
+Figure 3 displays the total donations by political party in Ontario.
 
-Add graphs, tables and text.
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+donations_by_party_ontario <- ontario_data %>%
+  group_by(political_party) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+#Plotting Donations by Political Party - Ontario
+
+p3_ontario <- ggplot(donations_by_party_ontario, aes(x = reorder(political_party, -total_amount), y = total_amount, fill = political_party)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Total Donations by Political Party - Ontario",
+       x = "Political Party", y = "Total Amount Donated ($)") +
+  scale_y_continuous(labels = scales::dollar_format()) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "none")
+
+print(p3_ontario)
+```
+
+*Figure 3: Total Donations by Political Party in Ontario*
+
+#### Federal
+
+Similarly, Table 2 shows the donations at the federal level.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+## Donations Based on Recipient Power Status ##
+donations_in_power_ontario <- ontario_data %>%
+  group_by(recipient_in_power) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
+```
+
+Similarly, Figure 2 shows the total and average donations over the years at the federal level.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+# Plotting Total and Average Donations Over Years - Federal
+donations_by_year_federal <- federal_data %>%
+  group_by(donation_year) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
+
+p1_federal <- ggplot(donations_by_year_federal, aes(x = donation_year, y = total_amount)) +
+  geom_line(color = "steelblue") +
+  geom_point(color = "steelblue") +
+  labs(title = "Total Donations Over Years - Federal",
+       x = "Year", y = "Total Amount Donated ($)") +
+  scale_y_continuous(labels = scales::dollar_format()) +
+  theme_minimal()
+
+p2_federal <- ggplot(donations_by_year_federal, aes(x = donation_year, y = avg_amount)) +
+  geom_line(color = "darkgreen") +
+  geom_point(color = "darkgreen") +
+  labs(title = "Average Donation Amount Over Years - Federal",
+       x = "Year", y = "Average Donation Amount ($)") +
+  scale_y_continuous(labels = scales::dollar_format()) +
+  theme_minimal()
+
+grid.arrange(p1_federal, p2_federal, ncol = 2)
+```
+
+*Figure 2: Total and Average Donations Over Years at Federal Level*
+
+Figure 4 shows the total donations by political party at the federal level.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+# Plotting Donations by Political Party - Federal
+donations_by_party_federal <- federal_data %>%
+  group_by(political_party) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  )
+
+p3_federal <- ggplot(donations_by_party_federal, aes(x = reorder(political_party, -total_amount), y = total_amount, fill = political_party)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Total Donations by Political Party - Federal",
+       x = "Political Party", y = "Total Amount Donated ($)") +
+  scale_y_continuous(labels = scales::dollar_format()) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "none")
+
+print(p3_federal)
+```
+
+*Figure 4: Total Donations by Political Party at Federal Level*
+
+## Relationships Between Variables
+
+Preliminary analysis indicates that political parties not in power receive higher total donations on average compared to those in power, at both federal and provincial levels.
+This trend suggests that donors may be motivated to support opposition parties to influence political change.
+
+Additionally, major parties consistently attract more donations than minor parties, highlighting the importance of party prominence in fundraising efforts.
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
 # Discussion
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive
+check. This shows...
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the
+prior. This shows...
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-12-03 paper/paper.qmd
update measurement

@@
 2023).
+[UPDATE REFERENCES]
 
@@
 
-The primary objective of this analysis is to investigate the impact of a political party's power status on donation patterns within Canadian federal and Ontario provincial elections.
-The dataset includes detailed records of donations made to various recipients, such as political parties, party leadership contestants, riding associations (also known as electoral district or constituency associations), and individual riding candidates.
+The dataset used in this analysis originates from the Investigative Journalism Foundation (IJF), which compiles political donation records from Canadian federal elections and the province of Ontario, covering the period from 2013 to 2024.
+The IJF collects this data by monitoring official elections agency websites at both the federal level and within each province and territory.
+Donations initially reported in various formats, including downloadable spreadsheets and PDFs, are converted into a standardized CSV format using Optical Character Recognition (OCR) tools such as Adobe Export PDF and ABBYY FineReader.
+Following conversion, manual verification is performed to correct any OCR-induced errors, ensuring the accuracy and consistency of the dataset.
+
+The target population for this analysis includes all political donations made to registered political parties, party leadership contestants, riding associations (also known as electoral district or constituency associations), and individual riding candidates during the specified time frame.
 Donations are tracked across different types of electoral events, including general elections, by-elections, and leadership races.
+This comprehensive coverage allows for an in-depth examination of donation patterns in various political contexts.
+The data collection methodology ensures that all significant financial contributions are captured, providing a robust foundation for analyzing the relationship between political power and financial support.
 
-Key variables within the dataset are as follows:
+Data aggregation involves summing the total donations received by each political party annually, resulting in a dataset where each row represents the total donations for a specific party in a given year.
+Key variables include Political Party (categorizing major and minor parties), Donation Year (ranging from 2013 to 2024), and Amount Donated (the monetary value of each donation).
+Additional binary variables such as Recipient_in_Power (indicating whether the party was in power during the donation year), Party_Size (distinguishing major from minor parties based on legislative representation), and Election_Year (indicating whether the donation was made during an election year) are constructed to facilitate regression analyses.
 
-- **Political Party:** A categorical variable indicating the recipient political party, encompassing both major parties (e.g., Liberal Party, Conservative Party, New Democratic Party) and minor parties.
-
-- **Donation Year:** A numerical variable representing the year the donation was made, ranging from 2013 to 2024. This captures different political climates and election periods.
-
-- **Amount Donated:** A continuous variable denoting the monetary value of each donation, varying widely due to differing donor capacities and legal contribution limits.
-
-### Constructed Variables
-
-Several variables were constructed to facilitate a more nuanced analysis:
-
-- **Recipient_in_Power:** A binary variable indicating whether the recipient party was in power during the donation year (1 for in power, 0 for not in power).
-This was determined by cross-referencing the donation year with historical records of party leadership at both the federal and provincial levels.
-
-- **Party_Size:** A binary variable categorizing parties as major (1) or minor (0), based on their representation in the legislature.
-Major parties are those with significant representation or historical influence.
-
-- **Election_Year:** A binary variable indicating whether the donation was made during an election year (1) or not (0), based on official election dates at the federal and provincial levels.
+The data is organized into a structured format with each row detailing a specific donation event, including the party receiving the donation, the year it was made, the amount donated, and the constructed binary indicators.
+This organization allows for efficient manipulation and analysis using statistical programming tools like R.
+By aggregating donations at the party-year level and incorporating relevant variables, the dataset effectively translates real-world donation phenomena into analyzable data points, enabling the investigation of how political power status influences financial support in Canadian elections.
 

### 2024-12-03 paper/paper.qmd
update sketches and model

@@
 
-The dataset used in this analysis comprises political donation records spanning from 2013 to 2024, covering both the federal level and the province of Ontario in Canada.
-These records were sourced from the Investigative Journalism Foundation (IJF), which systematically collects and maintains comprehensive data on political donations across various Canadian jurisdictions.
-The IJF gathers daily updates by monitoring elections agency websites at the federal level and within each province and territory, ensuring the dataset remains current and robust.
-Although the IJF's repository includes historical data dating back to 1993, this study focuses exclusively on the 2013–2024 period to capture recent trends and changes in political financing dynamics (Investigative Journalism Foundation 2024).
+The dataset used in this analysis comprises political donation records
+spanning from 2013 to 2024, covering both the federal level and the
+province of Ontario in Canada. These records were sourced from the
+Investigative Journalism Foundation (IJF), which systematically collects
+and maintains comprehensive data on political donations across various
+Canadian jurisdictions. The IJF gathers daily updates by monitoring
+elections agency websites at the federal level and within each province
+and territory, ensuring the dataset remains current and robust. Although
+the IJF's repository includes historical data dating back to 1993, this
+study focuses exclusively on the 2013–2024 period to capture recent
+trends and changes in political financing dynamics (Investigative
+Journalism Foundation 2024).
 
-The analysis was conducted using the statistical programming language R (R Core Team 2024) and several packages for data manipulation and visualization, including `dplyr` (Wickham et al.
-2023), `ggplot2` (Wickham 2016), `readr` (Wickham et al.
-2023), and `tidyr` (Wickham et al.
-2023).
-[UPDATE REFERENCES]
+The analysis was conducted using the statistical programming language R
+(R Core Team 2024) and several packages for data manipulation and
+visualization, including `dplyr` (Wickham et al. 2023), `ggplot2`
+(Wickham 2016), `readr` (Wickham et al. 2023), and `tidyr` (Wickham et
+al. 2023). \[UPDATE REFERENCES\]
 
-Political donations in Canada are subject to legal disclosure requirements, promoting transparency and accountability within the democratic process.
-Both federal and provincial election finance laws mandate that political parties, candidates, and associated entities report donations exceeding specific thresholds.
-These disclosures vary by jurisdiction in terms of frequency and reporting formats.
-For instance, Ontario requires annual returns for general and by-elections, leadership races, and donations to parties, candidates, and constituency associations, all available in downloadable spreadsheet formats from 2007 onward.
-This regulatory framework ensures that significant contributions are systematically documented and publicly accessible, providing a reliable foundation for this analysis.
+Political donations in Canada are subject to legal disclosure
+requirements, promoting transparency and accountability within the
+democratic process. Both federal and provincial election finance laws
+mandate that political parties, candidates, and associated entities
+report donations exceeding specific thresholds. These disclosures vary
+by jurisdiction in terms of frequency and reporting formats. For
+instance, Ontario requires annual returns for general and by-elections,
+leadership races, and donations to parties, candidates, and constituency
+associations, all available in downloadable spreadsheet formats from
+2007 onward. This regulatory framework ensures that significant
+contributions are systematically documented and publicly accessible,
+providing a reliable foundation for this analysis.
 
@@
 
-The dataset used in this analysis originates from the Investigative Journalism Foundation (IJF), which compiles political donation records from Canadian federal elections and the province of Ontario, covering the period from 2013 to 2024.
-The IJF collects this data by monitoring official elections agency websites at both the federal level and within each province and territory.
-Donations initially reported in various formats, including downloadable spreadsheets and PDFs, are converted into a standardized CSV format using Optical Character Recognition (OCR) tools such as Adobe Export PDF and ABBYY FineReader.
-Following conversion, manual verification is performed to correct any OCR-induced errors, ensuring the accuracy and consistency of the dataset.
+The dataset used in this analysis originates from the Investigative
+Journalism Foundation (IJF), which compiles political donation records
+from Canadian federal elections and the province of Ontario, covering
+the period from 2013 to 2024. The IJF collects this data by monitoring
+official elections agency websites at both the federal level and within
+each province and territory. Donations initially reported in various
+formats, including downloadable spreadsheets and PDFs, are converted
+into a standardized CSV format using Optical Character Recognition (OCR)
+tools such as Adobe Export PDF and ABBYY FineReader. Following
+conversion, manual verification is performed to correct any OCR-induced
+errors, ensuring the accuracy and consistency of the dataset.
 
-The target population for this analysis includes all political donations made to registered political parties, party leadership contestants, riding associations (also known as electoral district or constituency associations), and individual riding candidates during the specified time frame.
-Donations are tracked across different types of electoral events, including general elections, by-elections, and leadership races.
-This comprehensive coverage allows for an in-depth examination of donation patterns in various political contexts.
-The data collection methodology ensures that all significant financial contributions are captured, providing a robust foundation for analyzing the relationship between political power and financial support.
+The target population for this analysis includes all political donations
+made to registered political parties, party leadership contestants,
+riding associations (also known as electoral district or constituency
+associations), and individual riding candidates during the specified
+time frame. Donations are tracked across different types of electoral
+events, including general elections, by-elections, and leadership races.
+This comprehensive coverage allows for an in-depth examination of
+donation patterns in various political contexts. The data collection
+methodology ensures that all significant financial contributions are
+captured, providing a robust foundation for analyzing the relationship
+between political power and financial support.
 
-Data aggregation involves summing the total donations received by each political party annually, resulting in a dataset where each row represents the total donations for a specific party in a given year.
-Key variables include Political Party (categorizing major and minor parties), Donation Year (ranging from 2013 to 2024), and Amount Donated (the monetary value of each donation).
-Additional binary variables such as Recipient_in_Power (indicating whether the party was in power during the donation year), Party_Size (distinguishing major from minor parties based on legislative representation), and Election_Year (indicating whether the donation was made during an election year) are constructed to facilitate regression analyses.
+Data aggregation involves summing the total donations received by each
+political party annually, resulting in a dataset where each row
+represents the total donations for a specific party in a given year. Key
+variables include Political Party (categorizing major and minor
+parties), Donation Year (ranging from 2013 to 2024), and Amount Donated
+(the monetary value of each donation). Additional binary variables such
+as Recipient_in_Power (indicating whether the party was in power during
+the donation year), Party_Size (distinguishing major from minor parties
+based on legislative representation), and Election_Year (indicating
+whether the donation was made during an election year) are constructed
+to facilitate regression analyses.
 
-The data is organized into a structured format with each row detailing a specific donation event, including the party receiving the donation, the year it was made, the amount donated, and the constructed binary indicators.
-This organization allows for efficient manipulation and analysis using statistical programming tools like R.
-By aggregating donations at the party-year level and incorporating relevant variables, the dataset effectively translates real-world donation phenomena into analyzable data points, enabling the investigation of how political power status influences financial support in Canadian elections.
+The data is organized into a structured format with each row detailing a
+specific donation event, including the party receiving the donation, the
+year it was made, the amount donated, and the constructed binary
+indicators. This organization allows for efficient manipulation and
+analysis using statistical programming tools like R. By aggregating
+donations at the party-year level and incorporating relevant variables,
+the dataset effectively translates real-world donation phenomena into
+analyzable data points, enabling the investigation of how political
+power status influences financial support in Canadian elections.
 
@@
 
-The raw donation records underwent several preprocessing steps to ensure the dataset's integrity and suitability for analysis:
+The raw donation records underwent several preprocessing steps to ensure
+the dataset's integrity and suitability for analysis. Initially, records
+with missing or incomplete information in key variables such as Amount
+Donated or Political Party were excluded from the analysis to maintain
+data quality. This exclusion was essential to prevent potential biases
+and inaccuracies in the subsequent analysis.
 
-1.
-**Handling Missing Values:** Records with missing or incomplete information, particularly in key variables such as donation amount or recipient party, were excluded from the analysis to maintain data quality.
+To ensure consistency across the dataset, political party names were
+standardized. Variations like "Liberal Party of Ontario" and
+"Progressive Conservative Party of Ontario" were consolidated under a
+single standardized name, "Liberal Party of Ontario." This
+standardization facilitated accurate aggregation and comparison across
+different jurisdictions and time periods.
 
-2.
-**Standardizing Party Names:** Political party names were standardized to ensure consistency.
-For example, variations like "Progressive Conservative Party of Ontario" and "PC Party of Ontario" were consolidated under a single standardized name.
+Donations were aggregated to reflect total amounts per party per year,
+allowing for a higher-level analysis of donation patterns over time.
+Binary variables such as Recipient_in_Power, Party_Size, and
+Election_Year were created to streamline regression analyses, enabling
+the examination of how political power status, party prominence, and
+election cycles influence donation amounts.
 
-3. **Removing Outliers:** Extreme donation amounts that could skew the analysis were identified and removed based on predefined thresholds, acknowledging the legal donation limits in Canada.
-
-4. **Aggregating Donations:** Donations were aggregated to reflect total amounts per party per year, facilitating a higher-level analysis of donation patterns.
-
-5. **Creating Binary Variables:** Binary indicators were generated for key variables such as `Recipient_in_Power`, `Party_Size`, and `Election_Year` to streamline regression analysis.
-
-Additional details on the data cleaning process, including code snippets and specific handling of anomalies, are provided in Appendix A.
+Additional details on the data cleaning process, including specific
+handling of anomalies, are provided in Appendix A: Data Cleaning Notes.
 
@@
 
-The main outcome variable of interest is the **Amount Donated**, representing the monetary value of each donation made to a political party within a given year.
-This continuous variable is critical for assessing the financial support received by parties and understanding how it correlates with factors like power status and party size.
+The primary outcome variable in this analysis is the Amount Donated,
+representing the monetary value of each donation made to a political
+party within a given year. This continuous variable is essential for
+assessing the level of financial support received by political parties
+and understanding its relationship with factors such as power status and
+party size.
 
-The distribution of donation amounts varies between the federal and provincial levels.
-At the federal level, the average donation amount is higher, reflecting different legal contribution limits and donor behaviors.
+The distribution of Amount Donated exhibits significant variability,
+influenced by differing donor capacities and legal contribution limits
+across jurisdictions. At the federal level, the average donation amount
+is notably higher compared to the provincial level in Ontario,
+reflecting the broader scope and potentially larger donor bases. This
+difference underscores the importance of analyzing donation patterns
+within the appropriate contextual framework.
 
-## Predictor Variables
-
-Several key predictor variables were identified to evaluate factors influencing donation amounts:
-
-- **Recipient_in_Power:** Indicates whether the recipient party was in power during the donation year (1) or not (0).
-This variable is central to assessing the impact of political power on donation patterns.
-
-- **Party_Size:** Categorizes parties as major (1) or minor (0). Major parties are expected to receive more donations due to greater visibility and established support bases.
-
-- **Election_Year:** Indicates whether the donation was made during an election year (1) or not (0). Election years often see increased fundraising activities.
-
-### Additional Predictor Variables
-
-- **Log_Total_Donations:** The logarithm of total donations received by a party in a given year, calculated as `log(Total_Donations + 1)` to handle zero values and normalize the distribution.
-
-- **Interaction Terms:** Interaction terms such as `Recipient_in_Power * Party_Size` were included in the models to explore how the effect of being in power varies between major and minor parties.
-
-## Summary Statistics
-
-### Donations Based on Recipient Power Status
-
-#### Ontario
-
-Table 1 presents the total donations and average donation amounts in Ontario, categorized by whether the recipient party was in power.
-
-**Table 1: Donations by Power Status in Ontario**
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-## Donations Based on Recipient Power Status ##
-donations_in_power_ontario <- ontario_data %>%
-  group_by(recipient_in_power) %>%
-  summarize(
-    total_donations = n(),
-    total_amount = sum(amount, na.rm = TRUE),
-    avg_amount = mean(amount, na.rm = TRUE)
-  )
-```
-
-Figure 1 illustrates the total donations and average donation amounts over the years in Ontario.
+Summary statistics for Amount Donated are presented in Table 1,
+highlighting the total number of donations, the cumulative amount
+donated, and the average donation per party per year.
 
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Plotting Total and Average Donations Over Years - Ontario
+library(ggplot2)
+library(dplyr)
+library(gridExtra)
+library(scales)
+
+# Group and summarize the data
 donations_by_year_ontario <- ontario_data %>%
@@
 
-library(ggplot2)
-library(gridExtra)
-
-p1_ontario <- ggplot(donations_by_year_ontario, aes(x = donation_year, y = total_amount)) +
-  geom_line(color = "steelblue") +
-  geom_point(color = "steelblue") +
-  labs(title = "Total Donations Over Years - Ontario",
-       x = "Year", y = "Total Amount Donated ($)") +
-  scale_y_continuous(labels = scales::dollar_format()) +
-  theme_minimal()
-
-p2_ontario <- ggplot(donations_by_year_ontario, aes(x = donation_year, y = avg_amount)) +
-  geom_line(color = "darkgreen") +
-  geom_point(color = "darkgreen") +
-  labs(title = "Average Donation Amount Over Years - Ontario",
-       x = "Year", y = "Average Donation Amount ($)") +
-  scale_y_continuous(labels = scales::dollar_format()) +
-  theme_minimal()
-
-grid.arrange(p1_ontario, p2_ontario, ncol = 2)
-```
-
-*Figure 1: Total and Average Donations Over Years in Ontario*
-
-Figure 3 displays the total donations by political party in Ontario.
-
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-donations_by_party_ontario <- ontario_data %>%
-  group_by(political_party) %>%
-  summarize(
-    total_donations = n(),
-    total_amount = sum(amount, na.rm = TRUE),
-    avg_amount = mean(amount, na.rm = TRUE)
-  )
-
-#Plotting Donations by Political Party - Ontario
-
-p3_ontario <- ggplot(donations_by_party_ontario, aes(x = reorder(political_party, -total_amount), y = total_amount, fill = political_party)) +
-  geom_bar(stat = "identity") +
-  labs(title = "Total Donations by Political Party - Ontario",
-       x = "Political Party", y = "Total Amount Donated ($)") +
-  scale_y_continuous(labels = scales::dollar_format()) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "none")
-
-print(p3_ontario)
-```
-
-*Figure 3: Total Donations by Political Party in Ontario*
-
-#### Federal
-
-Similarly, Table 2 shows the donations at the federal level.
-
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-## Donations Based on Recipient Power Status ##
-donations_in_power_ontario <- ontario_data %>%
-  group_by(recipient_in_power) %>%
-  summarize(
-    total_donations = n(),
-    total_amount = sum(amount, na.rm = TRUE),
-    avg_amount = mean(amount, na.rm = TRUE)
-  )
-```
-
-Similarly, Figure 2 shows the total and average donations over the years at the federal level.
-
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Plotting Total and Average Donations Over Years - Federal
 donations_by_year_federal <- federal_data %>%
@@
 
-p1_federal <- ggplot(donations_by_year_federal, aes(x = donation_year, y = total_amount)) +
-  geom_line(color = "steelblue") +
-  geom_point(color = "steelblue") +
-  labs(title = "Total Donations Over Years - Federal",
-       x = "Year", y = "Total Amount Donated ($)") +
-  scale_y_continuous(labels = scales::dollar_format()) +
-  theme_minimal()
+# Convert donation_year to a properly ordered factor
+donations_by_year_ontario$donation_year <- factor(
+  donations_by_year_ontario$donation_year, 
+  levels = sort(unique(as.numeric(donations_by_year_ontario$donation_year)))
+)
 
-p2_federal <- ggplot(donations_by_year_federal, aes(x = donation_year, y = avg_amount)) +
-  geom_line(color = "darkgreen") +
-  geom_point(color = "darkgreen") +
-  labs(title = "Average Donation Amount Over Years - Federal",
-       x = "Year", y = "Average Donation Amount ($)") +
-  scale_y_continuous(labels = scales::dollar_format()) +
-  theme_minimal()
+donations_by_year_federal$donation_year <- factor(
+  donations_by_year_federal$donation_year, 
+  levels = sort(unique(as.numeric(donations_by_year_federal$donation_year)))
+)
 
-grid.arrange(p1_federal, p2_federal, ncol = 2)
+# Combine datasets and add a "Region" column
+donations_combined <- bind_rows(
+  donations_by_year_ontario %>% mutate(Region = "Ontario"),
+  donations_by_year_federal %>% mutate(Region = "Federal")
+)
+
+# Plot using facets with a shared Y-axis
+p_combined <- ggplot(donations_combined, aes(x = donation_year, y = avg_amount, group = Region, color = Region)) +
+  geom_line() +
+  geom_point() +
+  facet_wrap(~ Region, ncol = 1, scales = "free_y") +
+  labs(
+    title = "Average Donation Amount Over Years",
+    x = "Year",
+    y = "Average Donation Amount ($)"
+  ) +
+  scale_y_continuous(labels = dollar_format()) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+print(p_combined)
+
 ```
 
-*Figure 2: Total and Average Donations Over Years at Federal Level*
+## Predictor Variables
 
-Figure 4 shows the total donations by political party at the federal level.
+Several key predictor variables were identified to evaluate the factors
+influencing Amount Donated:
+
+-   Recipient_in_Power: This binary variable indicates whether the
+    recipient party was in power during the donation year (1) or not
+    (0). It is central to assessing the impact of political power on
+    donation patterns, allowing the analysis to determine if being in
+    power correlates with higher or lower donation amounts.
+
+-   Party_Size: Another binary variable categorizes parties as major (1)
+    or minor (0) based on their representation in the legislature. Major
+    parties are expected to receive more donations due to greater
+    visibility, established support bases, and broader outreach
+    capabilities. This variable helps in distinguishing the influence of
+    party prominence on financial support.
+
+-   Election_Year: This binary variable signifies whether the donation
+    was made during an election year (1) or not (0). Election years
+    often see increased fundraising activities as parties intensify
+    their efforts to secure electoral victories. Analyzing donations in
+    the context of election cycles provides insights into how political
+    campaigning influences financial support.
+
+-   Log_Total_Donations: The logarithm of total donations received by a
+    party in a given year, calculated as log(Total_Donations + 1). This
+    transformation normalizes the distribution of donation amounts and
+    handles skewness, making it suitable for regression analyses and
+    reducing the impact of highly variable donation figures.
+
+-   Interaction Terms: Interaction terms such as Recipient_in_Power \*
+    Party_Size were included in the models to explore how the effect of
+    being in power varies between major and minor parties. This allows
+    for a nuanced understanding of the interplay between a party’s power
+    status and its size in determining donation patterns.
+
+These predictor variables enable a comprehensive examination of how
+political power status, party size, and election cycles influence
+donation amounts. By incorporating both main effects and interaction
+terms, the analysis can capture complex relationships and provide deeper
+insights into the factors driving political donations.
 
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Plotting Donations by Political Party - Federal
-donations_by_party_federal <- federal_data %>%
-  group_by(political_party) %>%
+
+# Load required libraries
+library(dplyr)
+library(knitr)
+library(kableExtra)
+library(scales) # For dollar formatting
+
+## Donations by Year - Federal
+donations_in_power_federal <- federal_data %>%
+  group_by(recipient_in_power) %>%
   summarize(
@@
     avg_amount = mean(amount, na.rm = TRUE)
+  ) %>%
+  mutate(
+    total_amount = dollar(total_amount), 
+    avg_amount = dollar(avg_amount)      
   )
 
-p3_federal <- ggplot(donations_by_party_federal, aes(x = reorder(political_party, -total_amount), y = total_amount, fill = political_party)) +
-  geom_bar(stat = "identity") +
-  labs(title = "Total Donations by Political Party - Federal",
-       x = "Political Party", y = "Total Amount Donated ($)") +
-  scale_y_continuous(labels = scales::dollar_format()) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
-        legend.position = "none")
+# Render Ontario Donations Table
+kable(donations_in_power_federal, 
+      caption = "Donations Based on Recipient Power Status - Federal", 
+      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
+      format = "latex", 
+      booktabs = TRUE,  
+      align = c("l", "c", "c", "c")) %>%
+  kable_styling(latex_options = c("striped", "hold_position"))
 
-print(p3_federal)
 ```
 
-*Figure 4: Total Donations by Political Party at Federal Level*
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+## Donations Based on Recipient Power Status - Ontario
+donations_in_power_ontario <- ontario_data %>%
+  group_by(recipient_in_power) %>%
+  summarize(
+    total_donations = n(),
+    total_amount = sum(amount, na.rm = TRUE),
+    avg_amount = mean(amount, na.rm = TRUE)
+  ) %>%
+  mutate(
+    total_amount = dollar(total_amount), # Format as dollars
+    avg_amount = dollar(avg_amount)      # Format as dollars
+  )
 
-## Relationships Between Variables
+# Render Ontario Donations Table
+kable(donations_in_power_ontario, 
+      caption = "Donations Based on Recipient Power Status - Ontario", 
+      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
+      format = "latex", # For PDF/LaTeX output
+      booktabs = TRUE,  # Add LaTeX booktabs style
+      align = c("l", "c", "c", "c")) %>%
+  kable_styling(latex_options = c("striped", "hold_position"))
+```
 
-Preliminary analysis indicates that political parties not in power receive higher total donations on average compared to those in power, at both federal and provincial levels.
-This trend suggests that donors may be motivated to support opposition parties to influence political change.
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+# Load required libraries
+library(dplyr)
+library(ggplot2)
+library(treemapify)
 
-Additionally, major parties consistently attract more donations than minor parties, highlighting the importance of party prominence in fundraising efforts.
+# Summarize data for tree map
+donations_by_party_ontario <- ontario_data %>%
+  group_by(political_party) %>%
+  summarize(
+    total_amount = sum(amount, na.rm = TRUE)
+  ) %>%
+  arrange(desc(total_amount)) # Arrange for better visual hierarchy
+
+# Create the tree map
+ggplot(donations_by_party_ontario, aes(
+  area = total_amount, 
+  fill = political_party, 
+  label = paste(political_party, "\n$", scales::comma(total_amount))
+)) +
+  geom_treemap() +
+  geom_treemap_text(
+    color = "white", 
+    place = "center", 
+    grow = TRUE
+  ) +
+  labs(
+    title = "Total Donations by Political Party - Ontario",
+    subtitle = "Size of rectangles represents total donation amounts",
+    caption = "Data source: Ontario Political Donations"
+  ) +
+  theme_minimal() +
+  theme(legend.position = "none") # Hide legend to avoid redundancy
+
+```
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to
-investigate... Background details and diagnostics are included in
-[Appendix -@sec-model-details].
+To analyze the impact of a political party’s power status on donation
+patterns within Canadian federal and Ontario provincial elections,
+multiple log regression models were developed. Specifically, separate
+models were constructed for Ontario and Federal levels to account for
+regional differences in political dynamics and donation behaviors.
+Additionally, a focused regression model was created for the
+Conservative and Liberal parties to explore the interaction between
+power status and party size. These models aim to predict the logarithm
+of total donations received by political parties, providing insights
+into how various factors influence financial support.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft.
-Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both
-measured in millimeters.
+This study employs linear regression models to quantify the relationship
+between key predictor variables and the total amount of donations
+received by political parties. The primary models include:
 
+1.  Ontario Model
+This model focuses on donations within Ontario, considering whether the party was in power, the size of the party, and whether the donation occurred during an election year.
 ```{=tex}
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+\[
+\log(\text{Total Donations}) = \beta_0 + \beta_1 \cdot \text{Recipient in Power} + \beta_2 \cdot \text{Party Size} + \beta_3 \cdot \text{Election Year} + \epsilon
+\]
 ```
-We run the model in R [@citeR] using the `rstanarm` package of
-@rstanarm. We use the default priors from `rstanarm`.
+
+2.  Federal Model
+Similar to the Ontario model, this model examines federal-level donations with the same set of predictors.
+```{=tex}
+\[
+\log(\text{Total Donations}) = \beta_0 + \beta_1 \cdot \text{Recipient in Power} + \beta_2 \cdot \text{Party Size} + \beta_3 \cdot \text{Election Year} + \epsilon
+\]
+```
+
+3.  Conservative and Liberal Parties Model
+This model specifically targets the Conservative and Liberal parties, introducing an interaction term between power status and party type to explore how the effect of being in power differs between these two major parties.
+```{=tex}
+\[
+\begin{split}
+\log(\text{Total Donations}) = \beta_0 + \beta_1 \cdot \text{Recipient in Power} 
++ \beta_2 \cdot \text{Party} \\
++ \beta_3 \cdot (\text{Recipient in Power} \times \text{Party}) 
++ \beta_4 \cdot \text{Election Year} + \epsilon
+\end{split}
+\]
+
+```
+
+Each model uses the logarithm of total donations (Log_Total_Donations) as the dependent variable to stabilize variance and normalize the distribution of donation amounts.
+Predictor variables include binary indicators for whether the party was in power (In_Power), the size of the party (Party_Size), and whether the donation was made during an election year (Election_Year).
+The specialized model for Conservative and Liberal parties includes an additional binary variable (Party) and an interaction term (In_Power * Party) to capture the combined effect of power status and party type on donation amounts.
 
 ### Model justification
+Linear regression was chosen for this analysis due to its simplicity, interpretability, and suitability for modeling continuous, normally distributed outcome variables.
+The dependent variable, Log_Total_Donations, is appropriate for linear regression after log transformation, which addresses skewness and stabilizes variance.
+This transformation ensures that the assumptions of linear regression are better met, facilitating reliable coefficient estimates and hypothesis testing.
 
-We expect a positive relationship between the size of the wings and time
-spent aloft. In particular...
+Alternative modeling approaches, such as generalized linear models (GLMs), were considered but ultimately deemed unnecessary given the adequacy of linear regression in capturing the relationships of interest.
+GLMs offer more flexibility in handling different types of dependent variables and distributions but introduce additional complexity that was not required for this analysis.
+Linear regression provides clear and interpretable coefficients, allowing for straightforward assessment of the impact of each predictor on donation amounts.
 
-We can use maths by including latex between dollar signs, for instance
-$\theta$.
+By constructing separate models for Ontario and Federal levels, the analysis accounts for regional variations in political dynamics and donation behaviors.
+The specialized model for Conservative and Liberal parties further refines the analysis by exploring interaction effects, offering a more nuanced understanding of how power status influences donations differently across major political parties.
+This targeted approach enhances the ability to draw specific conclusions about the financial support mechanisms within Canadian political contexts.
+
+### Assumptions and Limitations
+The linear regression models operate under several key assumptions:
+
+Linearity: It is assumed that there is a linear relationship between each predictor variable and the dependent variable (Log_Total_Donations).
+This means that changes in predictors are associated with proportional changes in the outcome.
+
+Normality of Errors: The residuals (errors) of the models are assumed to be normally distributed. This assumption is critical for the validity of hypothesis tests and confidence intervals.
+
+Homoscedasticity: The variance of the residuals is assumed to be constant across all levels of the predictor variables.
+Homoscedasticity ensures that the model's predictions are equally reliable across different values of predictors.
+
+No Multicollinearity: Predictor variables are assumed to be independent of each other, with no excessive correlation.
+High multicollinearity can inflate standard errors and make it difficult to assess the individual effect of each predictor.
+
+Independence of Observations: Each observation is assumed to be independent of others, meaning that the donation amounts are not influenced by other donations in the dataset.
+
+Potential violations of these assumptions could affect the reliability and validity of the regression results.
+For example, non-linearity might lead to biased coefficient estimates, while heteroscedasticity could result in inefficient estimates and invalid inference.
+Although steps were taken to address multicollinearity through variable selection and standardization, some degree of correlation between predictors may still exist, potentially inflating standard errors.
+
+Limitations of this analysis include the reliance on historical donation data, which assumes that past patterns will continue into the future.
+Unforeseen political events or changes in donation regulations could disrupt established trends, reducing the model's predictive accuracy.
+Additionally, the exclusion of minor parties beyond the Conservative and Liberal parties may overlook nuanced donation behaviors within smaller political groups, limiting the generalizability of the findings.
+
+### Model Validation 
+Model validation was conducted to assess the predictive accuracy and generalizability of the regression models.
+This involved evaluating key metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square Error (RMSE) to determine how well the models explain the variance in Log_Total_Donations and their predictive performance.
+
+For the Ontario Model, the regression analysis yielded an R² of 0.645 and an Adjusted R² of 0.635, indicating that approximately 64.5% of the variance in log-transformed total donations is explained by the predictors.
+The RMSE for this model was calculated to assess the average prediction error, demonstrating a reasonable level of accuracy.
+
+The Federal Model showed an R² of 0.515 and an Adjusted R² of 0.503, suggesting that 51.5% of the variance in log-transformed total donations is accounted for by the predictors.
+The RMSE value indicated acceptable predictive performance, though slightly lower than the Ontario model, reflecting potential differences in donation dynamics at the federal level.
+
+In the Conservative and Liberal Parties Model, the inclusion of the interaction term between In_Power and Party resulted in an R² of 0.630 and an Adjusted R² of 0.600.
+This model effectively captures the combined effect of power status and party type on donation amounts, with the interaction term providing significant insights into how being in power influences donations differently for Conservative and Liberal parties.
+The RMSE for this model confirmed its robust predictive capability, comparable to the other models.
+
+Overall, the validation process confirmed that the linear regression models are robust and effective in predicting donation amounts based on the selected predictor variables.
+The models demonstrate good explanatory power and reasonable predictive accuracy, making them suitable for analyzing the factors influencing political donations in Canada.
+However, ongoing evaluation and potential model refinement are recommended to accommodate evolving political and financial landscapes, ensuring the models remain relevant and accurate over time.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+#| label: tbl-model-summary
+#| tbl-cap: "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models"
+#| echo: false
+#| message: false
+
+library(broom)
+library(knitr)
+library(kableExtra)
+
+ontario_data <- analysis_data %>% filter(region == "Ontario")
+
+# Aggregate total donations per party per year in Ontario
+total_donations_ontario <- ontario_data %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(amount, na.rm = TRUE),
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Run Regression for Ontario with Controls
+
+total_donations_ontario <- total_donations_ontario %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
+
+model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
+
+federal_data <- analysis_data %>% filter(region == "Federal")
+
+# Aggregate total donations per party per year at Federal level
+total_donations_federal <- federal_data %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(amount, na.rm = TRUE),
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Run Regression for Federal with Controls
+# For Federal
+total_donations_federal <- total_donations_federal %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))
+
+model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)
+
+cl_data <- analysis_data %>%
+  filter(political_party %in% c(
+    "Conservative Party of Canada",
+    "Liberal Party of Canada",
+    "Progressive Conservative Party of Ontario",
+    "Liberal Party of Ontario"
+  ))
+
+# Create variables
+cl_data <- cl_data %>%
+  mutate(
+    In_Power = recipient_in_power,  # Already binary (1/0)
+    Party = ifelse(
+      political_party %in% c("Conservative Party of Canada", "Progressive Conservative Party of Ontario"),
+      1,  # Conservative
+      0   # Liberal
+    )
+  )
+
+# Aggregate total donations per party per year
+total_donations_cl <- cl_data %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(amount, na.rm = TRUE),
+    In_Power = first(In_Power),
+    Party = first(Party),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Log-transform the dependent variable to stabilize variance
+total_donations_cl <- total_donations_cl %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))
+
+# Run the regression with interaction term
+model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)
+
+# Tidy model summaries
+tidy_ontario <- tidy(model_ontario) %>%
+  mutate(Model = "Ontario") %>%
+  select(Model, term, estimate, std.error, p.value)
+
+tidy_federal <- tidy(model_federal) %>%
+  mutate(Model = "Federal") %>%
+  select(Model, term, estimate, std.error, p.value)
+
+tidy_cl <- tidy(model_cl) %>%
+  mutate(Model = "Conservative/Liberal") %>%
+  select(Model, term, estimate, std.error, p.value)
+
+# Combine tidy summaries
+tidy_all <- bind_rows(tidy_ontario, tidy_federal, tidy_cl)
+
+# Round numeric values to 3 decimal places
+tidy_all <- tidy_all %>%
+  mutate(
+    estimate = round(estimate, 3),
+    std.error = round(std.error, 3),
+    p.value = round(p.value, 3)
+  )
+
+# Create a formatted table
+kable(tidy_all, 
+      caption = "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models",
+      col.names = c("Model", "Term", "Estimate", "Std. Error", "p-value"),
+      align = "lcccc") %>%
+  kable_styling(full_width = FALSE, position = "left") %>%
+  add_header_above(c(" " = 1, "Regression Coefficients" = 4)) %>%
+  column_spec(1, bold = TRUE) %>%
+  column_spec(5, color = "red")
+
+```
 
@@
 
-Our results are summarized in @tbl-modelresults.
+Table 1 presents the summary of key model estimates for the Ontario, Federal, and Conservative/Liberal models.
+These models predict the logarithm of total donations received by political parties based on whether the party was in power, the size of the party, and whether the donation occurred during an election year.
+The Conservative/Liberal model includes an interaction term to explore how the effect of being in power varies between these two major parties.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+#| tbl-cap: "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models"
+#| echo: false
+#| message: false
+
+library(modelsummary)
+
+# Define the list of models
+models <- list(
+  "Ontario" = model_ontario,
+  "Federal" = model_federal,
+  "Conservative/Liberal" = model_cl
+)
+
+# Define custom coefficient labels
+coef_labels <- c(
+  "(Intercept)" = "Intercept",
+  "In_Power" = "In Power",
+  "Party_Size" = "Party Size",
+  "Election_Year" = "Election Year",
+  "Party" = "Party (1=Conservative, 0=Liberal)",
+  "In_Power:Party" = "In Power * Party"
+)
+
+# Generate the regression table without significance stars, rounded to 3 decimals
+modelsummary(
+  models, 
+  coef_map = coef_labels, 
+  stars = FALSE,
+  fmt = 3,
+  gof_omit = "AIC|BIC|Log.Lik.|F|RMSE|R2",
+  output = "markdown"
+)
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+#| label: fig-model-estimates
+#| fig-cap: "Coefficient Estimates for Ontario, Federal, and Conservative/Liberal Models"
+#| echo: false
+#| message: false
+
+library(broom)
+library(ggplot2)
+library(dplyr)
+
+# Tidy models and combine
+tidy_all <- bind_rows(
+  tidy(model_ontario) %>% mutate(Model = "Ontario"),
+  tidy(model_federal) %>% mutate(Model = "Federal"),
+  tidy(model_cl) %>% mutate(Model = "Conservative/Liberal")
+)
+
+# Select relevant terms and exclude intercepts
+tidy_plot <- tidy_all %>%
+  filter(term != "(Intercept)") %>%
+  mutate(term = recode(term, 
+                       "In_Power" = "In Power",
+                       "Party_Size" = "Party Size",
+                       "Election_Year" = "Election Year",
+                       "Party" = "Party (1=Conservative, 0=Liberal)",
+                       "In_Power:Party" = "In Power * Party"))
+
+# Plot coefficients with error bars
+ggplot(tidy_plot, aes(x = term, y = estimate, color = Model)) +
+  geom_point(position = position_dodge(width = 0.5), size = 3) +
+  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), 
+                width = 0.2, position = position_dodge(width = 0.5)) +
+  labs(
+    title = "Coefficient Estimates for Ontario, Federal, and Conservative/Liberal Models",
+    x = "Predictors",
+    y = "Estimated Coefficients"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+The regression analysis reveals that Party Size is a significant predictor of total donations across both the Ontario and Federal models, indicating that larger parties receive more financial support.
+Specifically, in the Ontario model, Party Size has a coefficient of 2.302 (p < 0.001), while in the Federal model, it is 3.524 (p < 0.001).
+This positive relationship suggests that party prominence plays a crucial role in attracting donations.
+
+Conversely, the In Power variable is not statistically significant in the Ontario (p = 0.322) and Federal (p = 0.529) models, implying that being in power does not directly influence the total donations received by a party at these levels.
+However, in the Conservative/Liberal model, In Power has a significant positive effect (2.513, p < 0.001), and the interaction term In Power * Party is significantly negative (-4.865, p < 0.001).
+This interaction indicates that while being in power increases donations for Conservative parties, it decreases donations for Liberal parties, highlighting a differential impact based on party affiliation.
+
+The Election Year variable does not significantly predict donation amounts in any of the models, with p-values exceeding conventional significance thresholds (Ontario p = 0.616, Federal p = 0.975, Conservative/Liberal p = 0.325).
+This suggests that donations are not substantially influenced by whether they occur during an election year.
+
+Overall, the models explain a moderate portion of the variance in total donations, with R² values ranging from 0.3046 (Ontario) to 0.5825 (Federal), and the Conservative/Liberal model achieving an R² of 0.4481.
+These findings underscore the importance of party size in determining donation amounts, while the role of power status appears to be contingent on party affiliation.
 
@@
 
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The
-discussion is a chance to show off what you know and what you learnt
-from all this.
+##The Influence of Party Size on Donation Amounts {#sec-first-point}
 
 ## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what
-your point actually is.
 
@@
 
-# Appendix {.unnumbered}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive
-check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the
-prior. This shows...
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
 \newpage
 
+## Appendix B: IJF Data Collection Methodology
+
+The Investigative Journalism Foundation (IJF) employs a comprehensive approach to gather data on political donations in Ontario and at the federal level in Canada.
+Their methodology involves the following key steps:
+
+1.
+**Data Sources:** IJF collects data from official government databases, including Elections Canada and provincial electoral commissions, which provide detailed records of political donations.
+Additionally, they integrate information from publicly available financial disclosures of political parties and candidates.
+
+2.
+**Data Integration:** The collected data from various sources are consolidated into a centralized database.
+This integration process ensures that all donations are accurately matched to the corresponding political parties and election cycles.
+
+3.
+**Data Verification:** To maintain data integrity, IJF implements rigorous verification procedures.
+This includes cross-referencing donation records with multiple sources to confirm accuracy and eliminate duplicates or erroneous entries.
+
+4.
+**Categorization:** Donations are categorized based on factors such as the donor’s affiliation (individual or organization), the amount donated, the recipient party, and the timing of the donation relative to election periods.
+This categorization facilitates detailed analysis of donation patterns and trends.
+
+5.
+**Data Cleaning:** The dataset undergoes extensive cleaning to address missing values, correct inconsistencies, and standardize formats.
+This step is crucial for ensuring that the data is reliable and suitable for subsequent analysis.
+
+6.
+**Ethical Considerations:** IJF adheres to strict ethical guidelines in data collection, ensuring that all donor information is handled confidentially and in compliance with privacy laws.
+Sensitive information is anonymized where necessary to protect donor identities.
+
+By leveraging official records and implementing robust data management practices, IJF provides a reliable foundation for analyzing political donation trends in Canada.
+
+### Idealized Methodology
+
+1. **Comprehensive Data Sources:** 
+   - **Government Databases:** Utilize all available records from Elections Canada and provincial electoral bodies to ensure complete coverage of political donations.
+   - **Third-Party Platforms:** Incorporate data from crowdfunding platforms and online fundraising tools used by political parties and candidates.
+   - **Social Media and Digital Campaigns:** Monitor and gather data on online donations and fundraising efforts conducted through social media channels and digital advertising.
+
+2. **Enhanced Data Integration and Real-Time Updates:**
+   - **Automated Data Pipelines:** Implement automated systems to continuously collect and integrate data from multiple sources in real-time, reducing latency and ensuring up-to-date information.
+   - **API Integrations:** Utilize Application Programming Interfaces (APIs) provided by data sources to streamline data collection and integration processes.
+
+3. **Advanced Data Verification Techniques:**
+   - **Machine Learning Algorithms:** Employ machine learning models to detect anomalies, identify fraudulent donations, and ensure data accuracy.
+   - **Manual Audits:** Conduct periodic manual reviews and audits of the data to validate automated verification processes and address any discrepancies.
+
+4. **Detailed Categorization and Enrichment:**
+   - **Donor Demographics:** Collect additional demographic information about donors, such as age, occupation, and geographic location, to enable more granular analysis.
+   - **Donation Purpose:** Categorize donations based on their intended purpose (e.g., general support, specific campaigns, issue-based funding) to understand donor motivations.
+
+5. **Robust Data Cleaning and Standardization:**
+   - **Automated Cleaning Tools:** Utilize advanced data cleaning software to automatically detect and correct errors, handle missing values, and standardize data formats.
+   - **Standardized Taxonomies:** Develop and apply standardized taxonomies for categorizing donations and donor attributes, ensuring consistency across the dataset.
+
+6. **Ethical and Transparent Practices:**
+   - **Data Privacy Compliance:** Strictly adhere to data privacy regulations, ensuring that all personal information is securely stored and processed.
+   - **Transparency Reports:** Publish regular transparency reports detailing data collection methods, sources, and any limitations to build trust with stakeholders.
+
+7. **Comprehensive Documentation and Accessibility:**
+   - **Detailed Documentation:** Maintain thorough documentation of all data collection, processing, and analysis procedures to facilitate reproducibility and accountability.
+   - **Open Data Access:** Provide open access to anonymized datasets and analysis tools, allowing researchers and the public to engage with the data and conduct independent analyses.
+
+By implementing these enhancements, the methodology would not only improve the quality and reliability of the data but also enable more sophisticated analyses, leading to deeper insights into political donation dynamics.
+
+### Idealized Survey Design
+
+The survey should be meticulously designed to ensure high response rates, representativeness, and data quality. Below is an outline of an idealized survey for studying political donations:
+
+Survey Objectives
+- **Understand Donor Motivations:** Explore the reasons behind individuals and organizations choosing to donate to specific political parties.
+- **Assess Donation Patterns:** Identify patterns in donation amounts, frequency, and timing relative to election cycles.
+- **Evaluate Impact of Campaigns:** Determine how political campaigns and events influence donation behaviors.
+- **Gather Demographic Insights:** Collect demographic information to analyze how factors such as age, income, education, and geographic location affect donation decisions.
+
+Survey Structure
+
+**Section 1: Demographics**
+- **State of Residence:** Dropdown list of all provinces and territories.
+- **Age:** Multiple-choice options (e.g., 18-24, 25-34, etc.).
+- **Gender:** Options including Male, Female, Non-binary/Other, Prefer not to say.
+- **Income Level:** Income brackets (e.g., Less than $25,000, $25,000-$49,999, etc.).
+- **Education Level:** Highest degree obtained (e.g., High school, Bachelor’s, Master’s, etc.).
+- **Occupation:** Open-ended or categorized options.
+- **Political Affiliation:** Conservative, Liberal, NDP, Other.
+
+**Section 2: Donation Behavior**
+- **Are you currently registered to donate to political parties?**
+  - Yes
+  - No
+- **How frequently do you donate to political parties?**
+  - Never
+  - Rarely (once a year or less)
+  - Occasionally (a few times a year)
+  - Regularly (monthly or more)
+- **Average Donation Amount:** Slider or multiple-choice options (e.g., Less than $50, $50-$199, etc.).
+- **Preferred Donation Channels:** Multiple selections (e.g., Online platforms, Mail-in donations, In-person events, etc.).
+
+**Section 3: Motivations for Donating**
+- **What motivates you to donate to a political party?** (Select all that apply)
+  - Support for specific policies or issues
+  - Desire to influence election outcomes
+  - Loyalty to the party or its leaders
+  - Social pressure or community influence
+  - Tax benefits
+  - Other (please specify)
+- **Which factors most influence your decision to donate?** (Rank in order of importance)
+  - Candidate's leadership qualities
+  - Party's stance on key issues
+  - Party's past performance
+  - Recommendations from peers or influencers
+  - Media coverage
+  - Other (please specify)
+
+**Section 4: Impact of Campaigns and Events**
+- **Have recent political events influenced your donation behavior?**
+  - Yes
+  - No
+- **If yes, please specify the events and how they influenced your donations.** (Open-ended)
+- **How do you perceive the effectiveness of political campaigns in encouraging donations?**
+  - Very ineffective
+  - Ineffective
+  - Neutral
+  - Effective
+  - Very effective
+
+**Section 5: Feedback and Suggestions**
+- **What could political parties do to encourage more donations from supporters?** (Open-ended)
+- **Any additional comments or suggestions regarding political donations?** (Open-ended)
+
+Survey Administration
+
+1.
+**Sampling Method:** Utilize stratified random sampling to ensure representation across different provinces, income levels, education backgrounds, and political affiliations.
+Oversample underrepresented groups to enhance data diversity.
+
+2. **Data Collection Modes:**
+   - **Online Surveys:** Distribute the survey through email invitations, social media platforms, and online panels to reach a broad audience.
+   - **Telephone Surveys:** Conduct phone interviews to include respondents who may have limited internet access or prefer verbal communication.
+   - **In-Person Interviews:** Organize face-to-face interviews in selected regions to capture detailed qualitative data.
+
+3. **Incentives:** Offer incentives such as entry into a sweepstakes, gift cards, or charitable donations to encourage participation and improve response rates.
+
+4. **Survey Length and Design:**
+   - **Duration:** Design the survey to be completed within 10-15 minutes to minimize respondent fatigue.
+   - **Clarity:** Use clear and concise language, avoiding technical jargon to ensure respondents understand each question.
+   - **Question Types:** Employ a mix of multiple-choice, Likert scales, and open-ended questions to capture both quantitative and qualitative data.
+
+5. **Data Quality Measures:**
+   - **Pilot Testing:** Conduct a pilot survey to identify and rectify any issues with question wording, survey flow, or technical glitches.
+   - **Anonymity and Confidentiality:** Assure respondents that their answers are confidential and will be used solely for research purposes to encourage honest and accurate responses.
+   - **Response Validation:** Implement checks to prevent duplicate responses, ensure logical consistency, and verify the completeness of the survey.
+
+6. **Data Analysis:**
+   - **Weighting:** Apply weighting adjustments to account for any sampling biases and ensure that the survey results accurately reflect the population demographics.
+   - **Statistical Techniques:** Use appropriate statistical methods to analyze the data, identifying significant predictors of donation behavior and uncovering underlying trends.
+
 # References

### 2024-12-03 scripts/06-model_data.R

@@ -15,7 +15,7 @@ analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
 
 # Filter data for Ontario
-ontario_data <- donations_filtered %>% filter(region == "Ontario")
+ontario_data <- analysis_data %>% filter(region == "Ontario")
 
 # Aggregate total donations per party per year in Ontario
 total_donations_ontario <- ontario_data %>%
@@ -105,3 +105,4 @@ model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = to
 summary(model_cl)
 
 
+# Need to save the model

### 2024-12-03 paper/paper.qmd
update introduction

@@
 title: "Impact of Federal Party Leadership Changes on Donation Patterns"
-subtitle: "An Analysis of Political Donations in Canada 2013 - 2024"
+subtitle: "What Data from 2013 - 2024 Reveal About Party Funding Trends"
 author: 
@@
 
-Political donations play a critical role in supporting political
-campaigns and shaping electoral competition. In Canada, both federal and
-provincial elections rely heavily on contributions from individuals and
-organizations, which influence the political landscape and often reflect
-public support for different parties. Understanding what drives donation
-patterns is essential for ensuring transparency and fairness in the
-democratic process and for minimizing the risk of undue influence on
-policy-making.
+Political donations are an important part of democratic processes, providing essential financial support for political campaigns and influencing electoral competition.
+In Canada, both federal and provincial elections rely on contributions from individuals and organizations.
+This can shape the political landscape and reflect public support for different parties.
+Understanding the factors that drive donation patterns is crucial for ensuring transparency and fairness in the democratic process and for minimizing the risk of undue influence on policy-making.
 
-One key question is whether a party’s position in power affects the
-amount of financial support it receives. While studies in other
-countries, such as the U.S.A, have shown that incumbents often benefit
-from more donations, particularly from interest groups looking to
-maintain influence, there has been less focus on whether this holds true
-in Canada. Research like Fouirnaies and Hall’s work (REFERENCE HERE) in
-the U.S. shows that donors strategically target those in power, but how
-these dynamics play out in Canadian elections remains under explored.
-This paper aims to fill that gap by analyzing how a party's power status
-influences donor behaviour in federal and Ontario provincial elections.
+A key question in political finance is whether a party's position in power affects the amount of financial support it receives.
+Studies in other countries, such as the  @fouirnaies2014financial conducted in the United States, have shown that incumbents often benefit from more donations, particularly from interest groups seeking to maintain influence.
+However, there has been less focus on whether this holds true in Canada, where the political system and donation regulations differ significantly.
+This paper aims to fill that gap by analyzing how a party's power status influences donor behaviour in federal and Ontario provincial elections.
 
-Using donation data from 2013 to 2024, this study compares the financial
-support received by parties in power versus those in opposition. The
-analysis examines total donations by year and party while accounting for
-factors like election timing and party size (major party or not). By
-identifying trends in how donors respond to shifts in political power,
-this research sheds light on the financial strategies behind Canada’s
-democratic processes.
+The primary estimand of this study is the effect of a political party's power status on the total amount of financial donations it receives from individual contributors.
+By comparing the financial support received by parties in power versus those in opposition, this research aims to understand whether being in power affects a party's fundraising success.
 
-The results show that opposition parties tend to receive higher
-donations than those in power, suggesting that donors often prioritize
-supporting challengers, possibly to influence political change. These
-findings align with trends observed in other political systems but
-highlight unique patterns in Canada. This information is valuable not
-only for political parties developing fundraising strategies but also
-for policymakers and researchers seeking to ensure fairness and
-accountability in political financing.
+Using donation data from 2013 to 2024, this study examines total donations by year and party, accounting for factors such as election timing and party size (major party or not).
+The analysis employs linear regression models to quantify the relationship between power status and donation amounts, providing insights into donor motivations and the strategic considerations of political parties.
 
-The paper is organized as follows: Section \ref{sec-data} discusses the
-data sources and methodology, followed by an explanation of the
-variables used in the analysis. The regression models and results are
-then presented, along with interpretations. The paper concludes by
-discussing the broader implications of the findings and suggesting
-directions for future research.\[NEED TO UPDATE THE REFERENCES\]
+The results indicate that opposition parties tend to receive higher total donations than those in power, suggesting that donors often prioritize supporting challengers, possibly to influence political change.
+This information is valuable not only for political parties developing fundraising strategies but also for policymakers and researchers seeking to ensure fairness and accountability in political financing.
 
-# Data
+The paper is organized as follows: Section \ref{sec-data} discusses the data sources and methodology, including measurement and data cleaning processes.
+Section \ref{sec-model} presents the regression models used in the analysis, along with justification and validation.
+Section \ref{sec-results} outlines the results, and Section \ref{sec-discussion} discusses the implications, limitations, and suggestions for future research.
+Section \ref{sec-appendix} offers further detailed insights into the data, modeling approach, and methodology.
+
+# Data {#sec-data}
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Appendix {#sec-appx}
+# Appendix {#sec-appendix}
 

### 2024-12-03 paper/paper.qmd
update cleaning code

@@
 A key question in political finance is whether a party's position in power affects the amount of financial support it receives.
-Studies in other countries, such as the  @fouirnaies2014financial conducted in the United States, have shown that incumbents often benefit from more donations, particularly from interest groups seeking to maintain influence.
+Studies in other countries, such as the United States, have shown that incumbents often benefit from more donations, particularly from interest groups seeking to maintain influence (@fouirnaies2014financial).
 However, there has been less focus on whether this holds true in Canada, where the political system and donation regulations differ significantly.
@@
 ## Overview
+The dataset used in this analysis comprises political donation records from 2013 to 2024, covering both the federal level and the province of Ontario in Canada.
+These records were sourced from the @ijf_political_donors Political Donors Dataset, which systematically collects and maintains comprehensive data on political donations across Canadian jurisdictions.
+The IJF updates this dataset daily by monitoring election agency websites at the federal level and within each province and territory, ensuring the data remains current and reliable.
+The analysis was conducted using both R (@citeR) and Python (@python), leveraging a range of libraries for data manipulation and visualization.
+In R, key packages included dplyr (@dplyr), lubridate (@lubridate), tidyverse (@tidyverse), ggplot2 (@ggplot), readr (@readr), tibble (@tibble), and rstanarm (@rstanarm) for statistical modeling.
+In Python, pandas (@pandas), requests (@requests), and matplotlib (@matplotlib) were utilized for additional data processing and visualization tasks
 
-The dataset used in this analysis comprises political donation records
-spanning from 2013 to 2024, covering both the federal level and the
-province of Ontario in Canada. These records were sourced from the
-Investigative Journalism Foundation (IJF), which systematically collects
-and maintains comprehensive data on political donations across various
-Canadian jurisdictions. The IJF gathers daily updates by monitoring
-elections agency websites at the federal level and within each province
-and territory, ensuring the dataset remains current and robust. Although
-the IJF's repository includes historical data dating back to 1993, this
-study focuses exclusively on the 2013–2024 period to capture recent
-trends and changes in political financing dynamics (Investigative
-Journalism Foundation 2024).
-
-The analysis was conducted using the statistical programming language R
-(R Core Team 2024) and several packages for data manipulation and
-visualization, including `dplyr` (Wickham et al. 2023), `ggplot2`
-(Wickham 2016), `readr` (Wickham et al. 2023), and `tidyr` (Wickham et
-al. 2023). \[UPDATE REFERENCES\]
-
-Political donations in Canada are subject to legal disclosure
-requirements, promoting transparency and accountability within the
-democratic process. Both federal and provincial election finance laws
-mandate that political parties, candidates, and associated entities
-report donations exceeding specific thresholds. These disclosures vary
-by jurisdiction in terms of frequency and reporting formats. For
-instance, Ontario requires annual returns for general and by-elections,
-leadership races, and donations to parties, candidates, and constituency
-associations, all available in downloadable spreadsheet formats from
-2007 onward. This regulatory framework ensures that significant
-contributions are systematically documented and publicly accessible,
-providing a reliable foundation for this analysis.
+Political donations in Canada are subject to legal disclosure requirements, promoting transparency and accountability within the democratic process.
+Both federal and provincial election finance laws mandate that political parties, candidates, and associated entities report donations exceeding specific thresholds.
+These disclosures vary by jurisdiction in terms of frequency and reporting formats.
+For instance, Ontario requires annual returns for general and by-elections, leadership races, and donations to parties, candidates, and constituency associations, all available in downloadable spreadsheet formats from 2007 onward.
+This regulatory framework ensures that significant contributions are systematically documented and publicly accessible, providing a reliable foundation for this analysis.
 
@@
 
-The dataset used in this analysis originates from the Investigative
-Journalism Foundation (IJF), which compiles political donation records
-from Canadian federal elections and the province of Ontario, covering
-the period from 2013 to 2024. The IJF collects this data by monitoring
-official elections agency websites at both the federal level and within
-each province and territory. Donations initially reported in various
-formats, including downloadable spreadsheets and PDFs, are converted
-into a standardized CSV format using Optical Character Recognition (OCR)
-tools such as Adobe Export PDF and ABBYY FineReader. Following
-conversion, manual verification is performed to correct any OCR-induced
-errors, ensuring the accuracy and consistency of the dataset.
+The dataset originates from @ijf_political_donors compilation of political donation records, which includes donations made to registered political parties, party leadership contestants, riding associations, and individual candidates.
+Donations are tracked across different types of electoral events, including general elections, by-elections, and leadership races.
+The IJF collects this data by monitoring official elections agency websites at both the federal level and within each province and territory.
 
-The target population for this analysis includes all political donations
-made to registered political parties, party leadership contestants,
-riding associations (also known as electoral district or constituency
-associations), and individual riding candidates during the specified
-time frame. Donations are tracked across different types of electoral
-events, including general elections, by-elections, and leadership races.
-This comprehensive coverage allows for an in-depth examination of
-donation patterns in various political contexts. The data collection
-methodology ensures that all significant financial contributions are
-captured, providing a robust foundation for analyzing the relationship
-between political power and financial support.
+The target population for this analysis includes all political donations made to registered political parties, party leadership contestants, riding associations (also known as electoral district or constituency associations), and individual riding candidates during the specified time frame.
+This comprehensive coverage allows for an in-depth examination of donation patterns in various political contexts.
+The data collection methodology ensures that all significant financial contributions are captured, providing a robust foundation for analyzing the relationship between political power and financial support.
 
-Data aggregation involves summing the total donations received by each
-political party annually, resulting in a dataset where each row
-represents the total donations for a specific party in a given year. Key
-variables include Political Party (categorizing major and minor
-parties), Donation Year (ranging from 2013 to 2024), and Amount Donated
-(the monetary value of each donation). Additional binary variables such
-as Recipient_in_Power (indicating whether the party was in power during
-the donation year), Party_Size (distinguishing major from minor parties
-based on legislative representation), and Election_Year (indicating
-whether the donation was made during an election year) are constructed
-to facilitate regression analyses.
+Data aggregation involves summing the total donations received by each political party annually, resulting in a dataset where each row represents the total donations for a specific party in a given year.
+Key variables include Political Party (categorizing major and minor parties), Donation Year (ranging from 2013 to 2024), and Amount Donated (the monetary value of each donation).
+Additional binary variables such as Recipient_in_Power (indicating whether the party was in power during the donation year), Party_Size (distinguishing major from minor parties based on legislative representation), and Election_Year (indicating whether the donation was made during an election year) are constructed to facilitate regression analyses.
 
-The data is organized into a structured format with each row detailing a
-specific donation event, including the party receiving the donation, the
-year it was made, the amount donated, and the constructed binary
-indicators. This organization allows for efficient manipulation and
-analysis using statistical programming tools like R. By aggregating
-donations at the party-year level and incorporating relevant variables,
-the dataset effectively translates real-world donation phenomena into
-analyzable data points, enabling the investigation of how political
-power status influences financial support in Canadian elections.
+The data is organized into a structured format with each row detailing a specific donation event, including the party receiving the donation, the year it was made, the amount donated, and the constructed binary indicators.
+By aggregating donations at the party-year level and incorporating relevant variables, the dataset effectively translates real-world donation phenomena into analyzable data points, enabling the investigation of how political power status influences financial support in Canadian elections.
 
@@
 
-The raw donation records underwent several preprocessing steps to ensure
-the dataset's integrity and suitability for analysis. Initially, records
-with missing or incomplete information in key variables such as Amount
-Donated or Political Party were excluded from the analysis to maintain
-data quality. This exclusion was essential to prevent potential biases
-and inaccuracies in the subsequent analysis.
-
-To ensure consistency across the dataset, political party names were
-standardized. Variations like "Liberal Party of Ontario" and
-"Progressive Conservative Party of Ontario" were consolidated under a
-single standardized name, "Liberal Party of Ontario." This
-standardization facilitated accurate aggregation and comparison across
-different jurisdictions and time periods.
-
-Donations were aggregated to reflect total amounts per party per year,
-allowing for a higher-level analysis of donation patterns over time.
-Binary variables such as Recipient_in_Power, Party_Size, and
-Election_Year were created to streamline regression analyses, enabling
-the examination of how political power status, party prominence, and
-election cycles influence donation amounts.
-
-Additional details on the data cleaning process, including specific
-handling of anomalies, are provided in Appendix A: Data Cleaning Notes.
+The raw donation records were carefully preprocessed to ensure data integrity before analysis.
+First, the data was imported using the read_csv function from the readr package.
+Records with missing or incomplete information in key variables were excluded to maintain quality.
+Political party names were standardized to address inconsistencies, such as consolidating variations like "Liberal Party of Ontario" and "Ontario Liberal Party" under a single name.
+Donations were then aggregated at the party-year level to analyze total annual contributions received by each party.
+Binary variables, including Recipient_in_Power, Party_Size, and Election_Year, were created to support regression analysis.
+To normalize the distribution and handle zero values, a new variable, Log_Total_Donations, was generated by applying a logarithmic transformation to Total_Donations + 1.
+Finally, the cleaned dataset was cross-validated against the original records to ensure accuracy.
+Additional details about the data cleaning process can be found in Appendix \ref{sec-appendix}.
 
@@
 
-The primary outcome variable in this analysis is the Amount Donated,
-representing the monetary value of each donation made to a political
-party within a given year. This continuous variable is essential for
-assessing the level of financial support received by political parties
-and understanding its relationship with factors such as power status and
-party size.
+The primary outcome variable is the Total Amount Donated, representing the total monetary value of donations received by each political party in a given year.
+This continuous variable is essential for assessing the level of financial support and understanding its relationship with factors such as power status and party size.
 
-The distribution of Amount Donated exhibits significant variability,
-influenced by differing donor capacities and legal contribution limits
-across jurisdictions. At the federal level, the average donation amount
-is notably higher compared to the provincial level in Ontario,
-reflecting the broader scope and potentially larger donor bases. This
-difference underscores the importance of analyzing donation patterns
-within the appropriate contextual framework.
-
-Summary statistics for Amount Donated are presented in Table 1,
-highlighting the total number of donations, the cumulative amount
-donated, and the average donation per party per year.
+The distribution of total donation amounts varies between the federal and provincial levels.
+At the federal level, the average total donations are higher, reflecting broader donor bases and different legal contribution limits.
 

### 2024-12-03 scripts/03-clean_data.R

@@ -33,7 +33,7 @@ donations_filtered <- donations_filtered %>%
       region == "Ontario" & political_party %in% c("Ontario Liberal Party", "Liberal Party") ~ "Liberal Party of Ontario",
       region == "Ontario" & political_party == "New Democratic Party" ~ "New Democratic Party of Ontario",
       region == "Ontario" & political_party == "Bloc Québécois" ~ "Bloc Quebecois",
-      region == "Federal" & political_party == "Bloc Québécois" ~ "Bloc Quebecois",  # Example for Federal region
+      region == "Federal" & political_party == "Bloc Québécois" ~ "Bloc Quebecois",
       TRUE ~ political_party
     )
   )
@@ -51,21 +51,53 @@ donations_filtered <- donations_filtered %>%
     )
   )
 
-# Federal Government Parties in Power from 2011 to 2024
+#### Define Government Parties in Power ####
+
+# Federal Government Parties in Power from 2006 to 2024
 federal_government <- data.frame(
-  year = 2011:2024,
+  year = 2006:2024,
   party_in_power = c(
-    rep("Conservative Party of Canada", 4),  # 2011-2014
-    rep("Liberal Party of Canada", 10)       # 2015-2024
+    # 2006-2007: Conservative
+    rep("Conservative Party of Canada", 2),
+    
+    # 2008-2010: Conservative
+    rep("Conservative Party of Canada", 3),
+    
+    # 2011-2014: Conservative
+    rep("Conservative Party of Canada", 4),
+    
+    # 2015-2018: Liberal
+    rep("Liberal Party of Canada", 4),
+    
+    # 2019-2020: Liberal
+    rep("Liberal Party of Canada", 2),
+    
+    # 2021-2024: Liberal
+    rep("Liberal Party of Canada", 4)
   )
 )
 
-# Ontario Government Parties in Power from 2011 to 2024
+# Ontario Government Parties in Power from 2003 to 2024
 ontario_government <- data.frame(
-  year = 2011:2024,
+  year = 2003:2024,
   party_in_power = c(
-    rep("Liberal Party of Ontario", 7),      # 2011-2017
-    rep("Progressive Conservative Party of Ontario", 7)  # 2018-2024
+    # 2003-2006: Liberal
+    rep("Liberal Party of Ontario", 4),
+    
+    # 2007-2010: Liberal
+    rep("Liberal Party of Ontario", 4),
+    
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R
extended data source

@@ -15,7 +15,7 @@ library(readr)
 library(tidyr)   
 
 #### Read data ####
-analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
 # Separate data into Ontario and Federal datasets
 ontario_data <- analysis_data %>% filter(region == "Ontario")
@@ -34,7 +34,7 @@ summary_stats_ontario <- data.frame(
   Metric = c("Total Donations", "Total Amount Donated", "Average Donation Amount"),
   Value = c(total_donations_ontario, total_amount_ontario, avg_donation_ontario)
 )
-write_csv(summary_stats_ontario, "../data/02-analysis_data/02.1-exploration_data/Ontario/summary_statistics_ontario.csv")
+write_csv(summary_stats_ontario, "data/02-analysis_data/02.1-exploration_data/Ontario/summary_statistics_ontario.csv")
 
 #### Donations by Year ####
 donations_by_year_ontario <- ontario_data %>%
@@ -47,7 +47,7 @@ donations_by_year_ontario <- ontario_data %>%
 
 # Display Donations by Year
 print(donations_by_year_ontario)
-write_csv(donations_by_year_ontario, "../data/02-analysis_data/02.1-exploration_data/Ontario/donations_by_year_ontario.csv")
+write_csv(donations_by_year_ontario, "data/02-analysis_data/02.1-exploration_data/Ontario/donations_by_year_ontario.csv")
 
 ## Donations by Political Party ##
 donations_by_party_ontario <- ontario_data %>%
@@ -60,7 +60,7 @@ donations_by_party_ontario <- ontario_data %>%
 
 # Display Donations by Political Party
 print(donations_by_party_ontario)
-write_csv(donations_by_party_ontario, "../data/02-analysis_data/02.1-exploration_data/Ontario/donations_by_political_party_ontario.csv")
+write_csv(donations_by_party_ontario, "data/02-analysis_data/02.1-exploration_data/Ontario/donations_by_political_party_ontario.csv")
 
 ## Donations Based on Recipient Power Status ##
 donations_in_power_ontario <- ontario_data %>%
@@ -73,7 +73,7 @@ donations_in_power_ontario <- ontario_data %>%
 
 # Display Donations Based on Recipient Power Status
 print(donations_in_power_ontario)
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -14,8 +14,11 @@ library(rstanarm)
 analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
 
+#### Model for Ontario ####
+
 # Filter data for Ontario
-ontario_data <- analysis_data %>% filter(region == "Ontario")
+ontario_data <- analysis_data %>% 
+  filter(region == "Ontario")
 
 # Aggregate total donations per party per year in Ontario
 total_donations_ontario <- ontario_data %>%
@@ -28,17 +31,19 @@ total_donations_ontario <- ontario_data %>%
   ) %>%
   ungroup()
 
-# Run Regression for Ontario with Controls
-
+# Log-transform the dependent variable to stabilize variance
 total_donations_ontario <- total_donations_ontario %>%
   mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
 
+# Run Regression for Ontario with Controls
 model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
 summary(model_ontario)
 
+#### Model for Federal ####
 
 # Filter data for Federal
-federal_data <- donations_filtered %>% filter(region == "Federal")
+federal_data <- analysis_data %>% 
+  filter(region == "Federal")
 
 # Aggregate total donations per party per year at Federal level
 total_donations_federal <- federal_data %>%
@@ -51,19 +56,15 @@ total_donations_federal <- federal_data %>%
   ) %>%
   ungroup()
 
-# Run Regression for Federal with Controls
-# For Federal
+# Log-transform the dependent variable to stabilize variance
 total_donations_federal <- total_donations_federal %>%
-  mutate(Log_Total_Donations = log(Total_Donations + 1))
+  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
 
+# Run Regression for Federal with Controls
 model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)
 summary(model_federal)
 
-
-
-
-
-### Conservative and Liberal 
+#### Conservative and Liberal Analysis ####
 
 # Filter for Conservative and Liberal parties
 cl_data <- analysis_data %>%
@@ -98,11 +99,10 @@ total_donations_cl <- cl_data %>%
 
 # Log-transform the dependent variable to stabilize variance
[…]

### 2024-12-04 paper/paper.qmd
update model

@@
     recipient party was in power during the donation year (1) or not
-    (0). It is central to assessing the impact of political power on
-    donation patterns, allowing the analysis to determine if being in
-    power correlates with higher or lower donation amounts.
+    (0).
 
--   Party_Size: Another binary variable categorizes parties as major (1)
+-   Party_Size: Another binary variable which categorizes parties as major (1)
     or minor (0) based on their representation in the legislature. Major
@@
     reducing the impact of highly variable donation figures.
-
--   Interaction Terms: Interaction terms such as Recipient_in_Power \*
-    Party_Size were included in the models to explore how the effect of
-    being in power varies between major and minor parties. This allows
-    for a nuanced understanding of the interplay between a party’s power
-    status and its size in determining donation patterns.
-
-These predictor variables enable a comprehensive examination of how
-political power status, party size, and election cycles influence
-donation amounts. By incorporating both main effects and interaction
-terms, the analysis can capture complex relationships and provide deeper
-insights into the factors driving political donations.
 
@@
   ) %>%
-  arrange(desc(total_amount)) # Arrange for better visual hierarchy
+  arrange(desc(total_amount))
 
@@
   theme_minimal() +
-  theme(legend.position = "none") # Hide legend to avoid redundancy
+  theme(legend.position = "none")
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+# Load required libraries
+library(dplyr)
+library(ggplot2)
+library(treemapify)
+
+# Summarize data for tree map
+donations_by_party_federal <- federal_data %>%
+  group_by(political_party) %>%
+  summarize(
+    total_amount = sum(amount, na.rm = TRUE)
+  ) %>%
+  arrange(desc(total_amount))
+
+# Create the tree map
+ggplot(donations_by_party_federal, aes(
+  area = total_amount, 
+  fill = political_party, 
+  label = paste(political_party, "\n$", scales::comma(total_amount))
+)) +
+  geom_treemap() +
+  geom_treemap_text(
+    color = "white", 
+    place = "center", 
+    grow = TRUE
+  ) +
+  labs(
+    title = "Total Donations by Political Party - Federal",
+    subtitle = "Size of rectangles represents total donation amounts",
+    caption = "Data source: Federal Political Donations"
+  ) +
+  theme_minimal() +
+  theme(legend.position = "none")
 
@@
 patterns within Canadian federal and Ontario provincial elections,
-multiple log regression models were developed. Specifically, separate
+multiple linear regression models were developed. Specifically, separate
 models were constructed for Ontario and Federal levels to account for
@@
 Conservative and Liberal parties to explore the interaction between
-power status and party size. These models aim to predict the logarithm
-of total donations received by political parties, providing insights
-into how various factors influence financial support.
+power status and party size.
 
@@
 
-This study employs linear regression models to quantify the relationship
-between key predictor variables and the total amount of donations
-received by political parties. The primary models include:
+To analyze the impact of a political party's power status on donation patterns, we employed linear regression models for both Ontario and federal levels:
 
-1.  Ontario Model
-This model focuses on donations within Ontario, considering whether the party was in power, the size of the party, and whether the donation occurred during an election year.
 ```{=tex}
 \[
-\log(\text{Total Donations}) = \beta_0 + \beta_1 \cdot \text{Recipient in Power} + \beta_2 \cdot \text{Party Size} + \beta_3 \cdot \text{Election Year} + \epsilon
+\log(\text{Total Donations}_i) = \beta_0 + \beta_1 \cdot \text{Recipient\_in\_Power}_i + \beta_2 \cdot \text{Party\_Size}_i + \beta_3 \cdot \text{Election\_Year}_i + \epsilon_i
 \]
+
+Where:
+
+- \(\log(\text{Total Donations}_i)\): The logarithm of total donations received by party \(i\) in a given year.\\
+- \(\text{Recipient\_in\_Power}_i\): A binary variable indicating if party \(i\) was in power.\\
+- \(\text{Party\_Size}_i\): A binary variable indicating if party \(i\) is a major party.\\
+- \(\text{Election\_Year}_i\): A binary variable indicating if it was an election year.\\
+- \(\epsilon_i\): The error term.
 ```
 
-2.  Federal Model
-Similar to the Ontario model, this model examines federal-level donations with the same set of predictors.
-```{=tex}
-\[
-\log(\text{Total Donations}) = \beta_0 + \beta_1 \cdot \text{Recipient in Power} + \beta_2 \cdot \text{Party Size} + \beta_3 \cdot \text{Election Year} + \epsilon
-\]
-```
-
-3.  Conservative and Liberal Parties Model
-This model specifically targets the Conservative and Liberal parties, introducing an interaction term between power status and party type to explore how the effect of being in power differs between these two major parties.
+For the Conservative and Liberal parties, we included an interaction term:
 ```{=tex}
@@
 
+Where:
+- \(\text{Party}_i\): A binary variable indicating 1 for Conservative, 0 for Liberal.\\
+
 ```
@@
 ### Model justification
-Linear regression was chosen for this analysis due to its simplicity, interpretability, and suitability for modeling continuous, normally distributed outcome variables.
-The dependent variable, Log_Total_Donations, is appropriate for linear regression after log transformation, which addresses skewness and stabilizes variance.
-This transformation ensures that the assumptions of linear regression are better met, facilitating reliable coefficient estimates and hypothesis testing.
+Linear regression was chosen for its simplicity and interpretability.
+The log transformation of the dependent variable stabilizes variance and normalizes the distribution, making it suitable for linear modeling.
+The models account for key factors influencing donation amounts, allowing us to isolate the effect of a party being in power.
 
-Alternative modeling approaches, such as generalized linear models (GLMs), were considered but ultimately deemed unnecessary given the adequacy of linear regression in capturing the relationships of interest.
-GLMs offer more flexibility in handling different types of dependent variables and distributions but introduce additional complexity that was not required for this analysis.
-Linear regression provides clear and interpretable coefficients, allowing for straightforward assessment of the impact of each predictor on donation amounts.
-
-By constructing separate models for Ontario and Federal levels, the analysis accounts for regional variations in political dynamics and donation behaviors.
-The specialized model for Conservative and Liberal parties further refines the analysis by exploring interaction effects, offering a more nuanced understanding of how power status influences donations differently across major political parties.
-This targeted approach enhances the ability to draw specific conclusions about the financial support mechanisms within Canadian political contexts.
+Alternative models, such as generalized linear models, were considered but deemed unnecessary due to the adequacy of linear regression for this analysis.
 
 ### Assumptions and Limitations
-The linear regression models operate under several key assumptions:
+The models assume:
 
-Linearity: It is assumed that there is a linear relationship between each predictor variable and the dependent variable (Log_Total_Donations).
-This means that changes in predictors are associated with proportional changes in the outcome.
-
-Normality of Errors: The residuals (errors) of the models are assumed to be normally distributed. This assumption is critical for the validity of hypothesis tests and confidence intervals.
-
-Homoscedasticity: The variance of the residuals is assumed to be constant across all levels of the predictor variables.
-Homoscedasticity ensures that the model's predictions are equally reliable across different values of predictors.
-
-No Multicollinearity: Predictor variables are assumed to be independent of each other, with no excessive correlation.
-High multicollinearity can inflate standard errors and make it difficult to assess the individual effect of each predictor.
-
-Independence of Observations: Each observation is assumed to be independent of others, meaning that the donation amounts are not influenced by other donations in the dataset.
-
-Potential violations of these assumptions could affect the reliability and validity of the regression results.
-For example, non-linearity might lead to biased coefficient estimates, while heteroscedasticity could result in inefficient estimates and invalid inference.
-Although steps were taken to address multicollinearity through variable selection and standardization, some degree of correlation between predictors may still exist, potentially inflating standard errors.
-
-Limitations of this analysis include the reliance on historical donation data, which assumes that past patterns will continue into the future.
-Unforeseen political events or changes in donation regulations could disrupt established trends, reducing the model's predictive accuracy.
-Additionally, the exclusion of minor parties beyond the Conservative and Liberal parties may overlook nuanced donation behaviors within smaller political groups, limiting the generalizability of the findings.
+Linearity: The relationship between predictors and the log of total donations is linear.
+Normality of Errors: The residuals are normally distributed.
+Homoscedasticity: Constant variance of residuals.
+No Multicollinearity: Predictors are not highly correlated.
+Independence of Observations: Each observation is independent.
+Potential limitations include unobserved variables affecting donation amounts and the assumption that past trends predict future behavior.
 

### 2024-12-04 paper/paper.qmd
Update model tests

@@
 ### Assumptions and Limitations
-The models assume:
+The models assume a linear relationship between the predictors and the logarithm of total donations, ensuring that the effects of the predictors are additive and proportional.
+Additionally, they assume that the residuals are normally distributed (normality of errors) and exhibit constant variance (homoscedasticity).
+The models also rely on the absence of multicollinearity, meaning that the predictors are not highly correlated, and they assume independence of observations, where each data point is independent of the others.
+Potential limitations of these models include the possibility of unobserved variables influencing donation amounts and the inherent assumption that past trends can reliably predict future behavior.
 
-Linearity: The relationship between predictors and the log of total donations is linear.
-Normality of Errors: The residuals are normally distributed.
-Homoscedasticity: Constant variance of residuals.
-No Multicollinearity: Predictors are not highly correlated.
-Independence of Observations: Each observation is independent.
-Potential limitations include unobserved variables affecting donation amounts and the assumption that past trends predict future behavior.
-
-### Model Validation 
+### Model Validation ******
 Model validation was conducted to assess the predictive accuracy and generalizability of the regression models.
@@
 However, ongoing evaluation and potential model refinement are recommended to accommodate evolving political and financial landscapes, ensuring the models remain relevant and accurate over time.
-
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-#| label: tbl-model-summary
-#| tbl-cap: "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models"
-#| echo: false
-#| message: false
-
-library(broom)
-library(knitr)
-library(kableExtra)
-
-ontario_data <- analysis_data %>% filter(region == "Ontario")
-
-# Aggregate total donations per party per year in Ontario
-total_donations_ontario <- ontario_data %>%
-  group_by(political_party, donation_year) %>%
-  summarize(
-    Total_Donations = sum(amount, na.rm = TRUE),
-    In_Power = first(recipient_in_power),
-    Party_Size = first(Party_Size),
-    Election_Year = first(Election_Year)
-  ) %>%
-  ungroup()
-
-# Run Regression for Ontario with Controls
-
-total_donations_ontario <- total_donations_ontario %>%
-  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
-
-model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
-
-federal_data <- analysis_data %>% filter(region == "Federal")
-
-# Aggregate total donations per party per year at Federal level
-total_donations_federal <- federal_data %>%
-  group_by(political_party, donation_year) %>%
-  summarize(
-    Total_Donations = sum(amount, na.rm = TRUE),
-    In_Power = first(recipient_in_power),
-    Party_Size = first(Party_Size),
-    Election_Year = first(Election_Year)
-  ) %>%
-  ungroup()
-
-# Run Regression for Federal with Controls
-# For Federal
-total_donations_federal <- total_donations_federal %>%
-  mutate(Log_Total_Donations = log(Total_Donations + 1))
-
-model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)
-
-cl_data <- analysis_data %>%
-  filter(political_party %in% c(
-    "Conservative Party of Canada",
-    "Liberal Party of Canada",
-    "Progressive Conservative Party of Ontario",
-    "Liberal Party of Ontario"
-  ))
-
-# Create variables
-cl_data <- cl_data %>%
-  mutate(
-    In_Power = recipient_in_power,  # Already binary (1/0)
-    Party = ifelse(
-      political_party %in% c("Conservative Party of Canada", "Progressive Conservative Party of Ontario"),
-      1,  # Conservative
-      0   # Liberal
-    )
-  )
-
-# Aggregate total donations per party per year
-total_donations_cl <- cl_data %>%
-  group_by(political_party, donation_year) %>%
-  summarize(
-    Total_Donations = sum(amount, na.rm = TRUE),
-    In_Power = first(In_Power),
-    Party = first(Party),
-    Election_Year = first(Election_Year)
-  ) %>%
-  ungroup()
-
-# Log-transform the dependent variable to stabilize variance
-total_donations_cl <- total_donations_cl %>%
-  mutate(Log_Total_Donations = log(Total_Donations + 1))
-
-# Run the regression with interaction term
-model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)
-
-# Tidy model summaries
-tidy_ontario <- tidy(model_ontario) %>%
-  mutate(Model = "Ontario") %>%
-  select(Model, term, estimate, std.error, p.value)
-
-tidy_federal <- tidy(model_federal) %>%
-  mutate(Model = "Federal") %>%
-  select(Model, term, estimate, std.error, p.value)
-
-tidy_cl <- tidy(model_cl) %>%
-  mutate(Model = "Conservative/Liberal") %>%
-  select(Model, term, estimate, std.error, p.value)
-
-# Combine tidy summaries
-tidy_all <- bind_rows(tidy_ontario, tidy_federal, tidy_cl)
-
-# Round numeric values to 3 decimal places
-tidy_all <- tidy_all %>%
-  mutate(
-    estimate = round(estimate, 3),
-    std.error = round(std.error, 3),
-    p.value = round(p.value, 3)
-  )
-
-# Create a formatted table
-kable(tidy_all, 
-      caption = "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models",
-      col.names = c("Model", "Term", "Estimate", "Std. Error", "p-value"),
-      align = "lcccc") %>%
-  kable_styling(full_width = FALSE, position = "left") %>%
-  add_header_above(c(" " = 1, "Regression Coefficients" = 4)) %>%
-  column_spec(1, bold = TRUE) %>%
-  column_spec(5, color = "red")
-
-```
 

### 2024-12-04 scripts/07-test_model.R

@@ -1,0 +1,254 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Maria Mangru
+# Date: 25 November 2024
+# Contact: maria.mangru@mail.utoronto.ca
+# License: MIT
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(caret)
+library(car)
+library(broom)
+library(ggplot2)
+
+#### Read data ####
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+
+# Ensure 'amount' is numeric
+analysis_data <- analysis_data %>%
+  mutate(amount = as.numeric(amount))
+
+#### Model for Ontario ####
+
+# Filter data for Ontario
+ontario_data <- analysis_data %>% 
+  filter(region == "Ontario")
+
+# Aggregate total donations per party per year in Ontario
+total_donations_ontario <- ontario_data %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(amount, na.rm = TRUE),
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Log-transform the dependent variable to stabilize variance
+total_donations_ontario <- total_donations_ontario %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))
+
+# Split data into training and test sets
+set.seed(123)
+train_index_ont <- createDataPartition(total_donations_ontario$Log_Total_Donations, p = 0.8, list = FALSE)
+train_data_ont <- total_donations_ontario[train_index_ont, ]
+test_data_ont <- total_donations_ontario[-train_index_ont, ]
+
+# Run Regression for Ontario with Controls on training data
+model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = train_data_ont)
+summary(model_ontario)
+
+# Predict on test data
+test_data_ont$Predicted_Log_Total_Donations <- predict(model_ontario, newdata = test_data_ont)
+
+# Calculate RMSE
+rmse_ontario <- sqrt(mean((test_data_ont$Log_Total_Donations - test_data_ont$Predicted_Log_Total_Donations)^2))
+print(paste("RMSE for Ontario model:", round(rmse_ontario, 4)))
+
+# Model diagnostics for Ontario model
+par(mfrow = c(2,2))
+plot(model_ontario)
+par(mfrow = c(1,1))
+
[…]

### 2024-12-04 paper/paper.qmd
update appendix

@@
 ---
-title: "Impact of Federal Party Leadership Changes on Donation Patterns"
+title: "Impact of Party Leadership on Donation Patterns"
 subtitle: "What Data from 2013 - 2024 Reveal About Party Funding Trends"
@@
 
-Table 1 presents the summary of key model estimates for the Ontario, Federal, and Conservative/Liberal models.
+Table @tbl-modelsummary presents the summary of key model estimates for the Ontario, Federal, and Conservative/Liberal models.
 These models predict the logarithm of total donations received by political parties based on whether the party was in power, the size of the party, and whether the donation occurred during an election year.
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
+#| label: tbl-modelsummary
 #| tbl-cap: "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models"
@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
+#| label: fig-estimates
 #| label: fig-model-estimates
@@
 ##The Influence of Party Size on Donation Amounts {#sec-first-point}
+At the heart of politics is finance, as revealed by Traag's paper on the complex contagion of campaign donations @traag2016complex.
+Their study demonstrates that a majority of campaign contributions in the US originate from a few wealthy individuals through large social networks.
+With the constant battle between government and corporations, the findings in this study underscore the significant influence of financial power in shaping political dynamics.\\
+
+The landscape in Canada for political financing is growing each year.
+Results from studies demonstrate that individual contributors may prefer to support challengers for several reasons.
+An optimistic but naive explanation may be that individuals hope to foster competitive elections, given that donations depend not on political leaning but on who is currently in power.
+However, a more realistic explanation for this phenomenon is that some individuals want to make policy changes.
+It is no secret that politicians, particularly those out of power, often solicit funding and political support in exchange for advancing the interests of their stakeholders @mollona2022ethics.
+These two dynamics may be working in tandem to create an effect where parties eager to be elected rely more on funding from external sources, which can switch each election cycle.
+
+Moreover, these trends of political donations fluctuating based on the party in power are not unique to Canada.
+The study "Financing UK Democracy: A Stocktake of 20 Years of Political Donations" highlights a surge in private political donations, more specifically from superdonors, widening the resource gap between large parties @draca2022financing.
+Similar to Canada, the UK also sees large donations from a wealthy few influencing their elections each year.
+The parallel between Canada and the UK underscores a broader democratic issue of financial power being used to influence political outcomes.
+Though this concept is not novel, this study contributes to the already growing spotlight on the persistent lack of proper reform in political systems.
+
+This study provides valuable insights into donation behaviours across Canada but it its limitations.
+First, the analysis is applied to only federal and provincial donation records, overlooking regional variations found in other provinces and territories.
+The geographical limitations of this study hinder its ability to generalize across the entire country.
+Moreover, the study only spans from 2006 to 2024 which economic or political events could have influence over the donation patterns that the models do not account for.
 
@@
 ## Appendix A: Data Cleaning Notes
+The dataset used in this analysis underwent a rigorous data cleaning process to ensure accuracy, consistency, and reliability.
+The raw donation records were sourced from the Investigative Journalism Foundation (IJF), encompassing political donations at both the federal and Ontario provincial levels from 2013 to 2024.
+The data cleaning process involved several critical steps, detailed below.
 
-The data cleaning process was meticulously carried out to ensure the
-dataset's reliability and consistency. The key steps included:
+Initially, the dataset was examined for missing or incomplete values in key variables such as Amount Donated, Political Party, Donation Year, and Region.
+Records with missing values in these essential fields were excluded from the analysis to maintain data integrity and prevent biases that could arise from incomplete data entries.
 
-1.  **Importing Raw Data:**
-    -   The raw donation data was imported using the `read_csv` function
-        from the `tidyverse` package.
-    -   Data from PDF sources were converted to CSV using OCR tools such
-        as Adobe Export PDF and ABBYY FineReader, followed by manual
-        verification to correct OCR-induced errors.
-2.  **Handling Missing Values:**
-    -   Missing values in critical variables (`amount_donated`,
-        `recipient_in_power`, `party_size`, `election_year`) were
-        addressed by excluding incomplete records to maintain data
-        integrity.
-3.  **Standardizing Party Names:**
-    -   Political party names were standardized across all
-        jurisdictions. For example, variations like “Progressive
-        Conservative,” “Progressive Conservative Party of Ontario,” and
-        “The Progressive Conservative Association of Ontario” were
-        consolidated into “Progressive Conservative Association of
-        Ontario.”
-4.  **Removing Outliers:**
-    -   Donation amounts exceeding three standard deviations from the
-        mean were identified and removed to prevent skewing the
-        analysis.
-5.  **Aggregating Donations:**
-    -   Donations were aggregated at the party-year level, resulting in
-        a dataset where each row represents the total donations received
-        by a party in a given year.
-6.  **Creating Binary Variables:**
-    -   Binary indicators were created for `recipient_in_power`,
-        `party_size`, and `election_year` to facilitate regression
-        analysis.
-7.  **Log Transformation:**
-    -   The `log_total_donations` variable was created by taking the
-        logarithm of `Total_Donations + 1` to normalize the distribution
-        and handle zero values.
-8.  **Final Verification:**
-    -   The cleaned dataset was cross-validated against original records
-        to ensure accuracy and consistency.
+Political party names often appeared in various formats due to inconsistencies in reporting.
+To address this, party names were standardized across all records.
+For instance, variations such as "Liberal Party of Ontario" and "Ontario Liberal Party" were consolidated under the standardized name "Ontario Liberal Party." This standardization ensured accurate attribution of donations to the correct parties.
+
+Donor information, including names and locations, was also standardized where possible.
+Typographical errors were corrected, and formatting was made consistent (e.g., "First Name Last Name" instead of "Last Name, First Name").
+However, original donor names were retained in cases where standardization could introduce inaccuracies, particularly when dealing with individuals with similar names.
+
+The dataset included both monetary (amount_monetary) and non-monetary (amount_non_monetary) contributions.
+A new variable, Amount Donated, was created by summing these two fields for each record.
+This provided a comprehensive measure of the total value of donations received by each party.
+All monetary values were converted to a consistent currency format, and any discrepancies in currency symbols or formats were corrected to ensure accurate aggregation and comparison of donation amounts.
+
+To facilitate analysis at the party-year level, individual donation records were aggregated.
+Donations were summed for each political party within each year, resulting in a dataset where each entry represented the total donations received by a party in a specific year.
+This aggregation allowed for the examination of trends over time and assessment of the impact of factors such as power status and election years on donation amounts.
+
+Several binary variables were introduced to capture key aspects of the data:
+
+Recipient_in_Power: Indicates whether the recipient party was in power during the donation year (1 for in power, 0 for not in power).
+Party_Size: Categorizes parties as major (1) or minor (0) based on legislative representation and historical influence.
+Election_Year: Indicates whether the donation was made during an election year (1 for election year, 0 for non-election year).
+These variables were crucial for the regression analysis, allowing for the assessment of the influence of these factors on total donations.
+
+The distribution of total donation amounts was highly skewed, with a small number of large donations and many smaller ones.
+To address this skewness and stabilize variance, a logarithmic transformation was applied to the Total Donations variable, creating Log_Total_Donations.
+This transformation also helped in handling zero values by adding a small constant (e.g., 1) before taking the logarithm.
+
+To ensure the accuracy of the cleaned dataset, cross-validation was performed against original records and official financial statements where available.
+This step involved spot-checking aggregated totals and individual records to confirm that the cleaning and aggregation processes did not introduce errors.
+
+Some challenges were encountered due to inconsistencies in the raw data, such as typographical errors, inconsistent reporting formats, and missing donor types.
+While efforts were made to standardize and correct these issues, some limitations remain due to the quality of the original data sources.
+In cases where donor type information was missing and legal restrictions indicated that only individuals could donate (e.g., Ontario after 2017), the donor type was inferred as "Individual." While this assumption is reasonable based on the legal context, it may not capture all nuances in the data.
 
@@
 
-The Investigative Journalism Foundation (IJF) employs a comprehensive approach to gather data on political donations in Ontario and at the federal level in Canada.
-Their methodology involves the following key steps:
+The Investigative Journalism Foundation (IJF) employs a comprehensive approach to collect and compile political donation data across Canada.
+Understanding the IJF's data collection methodology is essential for evaluating the reliability and scope of the data used in this analysis.
 
-1.
-**Data Sources:** IJF collects data from official government databases, including Elections Canada and provincial electoral commissions, which provide detailed records of political donations.
-Additionally, they integrate information from publicly available financial disclosures of political parties and candidates.
+Data Sources and Acquisition
 
-2.
-**Data Integration:** The collected data from various sources are consolidated into a centralized database.
-This integration process ensures that all donations are accurately matched to the corresponding political parties and election cycles.
+Every day, the IJF checks election agency websites at the federal level and in each province and territory for new political donations data.
+Historical data is obtained from a variety of government sources, including provincial archives, legislative libraries, and elections agencies, covering the period from 1993 to the present day, though this time range varies by jurisdiction.
 
-3.
-**Data Verification:** To maintain data integrity, IJF implements rigorous verification procedures.
-This includes cross-referencing donation records with multiple sources to confirm accuracy and eliminate duplicates or erroneous entries.
+Political parties and candidates are legally required to submit records of donations, which are maintained and made publicly accessible by election agencies.
+The frequency and format of these disclosures differ among jurisdictions; some require quarterly returns, while others mandate annual reports.
+The IJF systematically gathers these records to create a unified database of political donations.
 
-4.
-**Categorization:** Donations are categorized based on factors such as the donor’s affiliation (individual or organization), the amount donated, the recipient party, and the timing of the donation relative to election periods.
-This categorization facilitates detailed analysis of donation patterns and trends.
+Scope of Data Collected
 
-5.
-**Data Cleaning:** The dataset undergoes extensive cleaning to address missing values, correct inconsistencies, and standardize formats.
-This step is crucial for ensuring that the data is reliable and suitable for subsequent analysis.
+The IJF's dataset encompasses donations to various political entities, including:
 
-6.
-**Ethical Considerations:** IJF adheres to strict ethical guidelines in data collection, ensuring that all donor information is handled confidentially and in compliance with privacy laws.
-Sensitive information is anonymized where necessary to protect donor identities.
+Parties: Registered political parties at the federal and provincial levels.
+Candidates: Individual candidates running for office.
+Riding Associations: Also known as electoral district or constituency associations.
+Leadership and Nomination Contestants: Individuals participating in party leadership races or seeking nomination as candidates.
+The dataset includes records from all 13 provinces and territories, as well as city-specific donations from municipalities like Victoria and Vancouver.
 
-By leveraging official records and implementing robust data management practices, IJF provides a reliable foundation for analyzing political donation trends in Canada.
+Donation Laws and Contribution Limits
+
+Contribution limits and the types of entities permitted to donate vary across jurisdictions.
+For instance, at the federal level, only individuals who are Canadian citizens or permanent residents can donate, with maximum annual contributions set at specified amounts.
+In contrast, some provinces allow donations from corporations, unions, and other organizations and may have different or no contribution limits.
+
+Data Formats and Conversion
+
+The source data is available in multiple formats, including downloadable spreadsheets, PDFs, and HTML files.
+For jurisdictions where data is only available in PDF format, the IJF employs Optical Character Recognition (OCR) technology to convert the documents into machine-readable CSV files.
+OCR tools such as Adobe Export PDF, Cometdocs, and ABBYY FineReader are used for this purpose.
+
+Data Cleaning and Standardization
+
+Extensive manual cleaning is performed to correct errors arising from OCR conversion and to standardize the data. Key steps in the data cleaning process include:
+
+Correcting OCR Errors: Manual verification and correction of misread characters (e.g., misinterpreted currency symbols or letters).
+Standardizing Donor and Party Names: Ensuring consistency in naming conventions for donors and political parties (e.g., standardizing "Progressive Conservative Party of Ontario" across all records).
+Formatting Dates and Amounts: Standardizing date formats to YYYY-MM-DD and ensuring monetary values are consistently formatted.
+The IJF also amalgamates similar categories in variables such as donor types to improve data legibility (e.g., merging "Business" and "Corporation" into a single "Corporation" category).
+In cases where donor type information is missing and laws specify that only individuals can donate (e.g., Ontario after 2017), the IJF adds "Individual" to the donor type column.
+
+Limitations and Challenges
+
+While the IJF endeavors to provide accurate and comprehensive data, several limitations exist:
+
+Typos and Reporting Errors: Original records submitted by political entities may contain typographical errors or inaccuracies, which are reflected in the dataset.
+Incomplete or Inconsistent Data: Variations in reporting requirements and practices across jurisdictions can lead to inconsistencies in the data.
+OCR Limitations: Despite manual verification, OCR technology may not perfectly capture handwritten or poorly scanned documents, potentially introducing errors.
+Conclusion
+
+The IJF's meticulous data collection and cleaning processes result in a robust dataset suitable for analyzing political donation patterns in Canada.
+Understanding the methodologies employed allows for a better assessment of the data's strengths and limitations, ensuring analyses are grounded in a thorough appreciation of the underlying data quality.
 

### 2024-12-04 paper/paper.qmd
update appendix

@@
 
+#### Read data ####
+analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")
+
+#### Model for Ontario ####
+
+# Filter data for Ontario
+ontario_data <- analysis_data %>% 
+  filter(region == "Ontario")
+
+# Aggregate total donations per party per year in Ontario
+total_donations_ontario <- ontario_data %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(amount, na.rm = TRUE),
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Log-transform the dependent variable to stabilize variance
+total_donations_ontario <- total_donations_ontario %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
+
+# Run Regression for Ontario with Controls
+model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
+
+#### Model for Federal ####
+
+# Filter data for Federal
+federal_data <- analysis_data %>% 
+  filter(region == "Federal")
+
+# Aggregate total donations per party per year at Federal level
+total_donations_federal <- federal_data %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(amount, na.rm = TRUE),
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Log-transform the dependent variable to stabilize variance
+total_donations_federal <- total_donations_federal %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
+
+# Run Regression for Federal with Controls
+model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)
+
+# Filter for Conservative and Liberal parties
+cl_data <- analysis_data %>%
+  filter(political_party %in% c(
+    "Conservative Party of Canada",
+    "Liberal Party of Canada",
+    "Progressive Conservative Party of Ontario",
+    "Liberal Party of Ontario"
+  ))
+
+# Create variables
+cl_data <- cl_data %>%
+  mutate(
+    In_Power = recipient_in_power,  # Already binary (1/0)
+    Party = ifelse(
+      political_party %in% c("Conservative Party of Canada", "Progressive Conservative Party of Ontario"),
+      1,  # Conservative
+      0   # Liberal
+    )
+  )
+
+# Aggregate total donations per party per year
+total_donations_cl <- cl_data %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(amount, na.rm = TRUE),
+    In_Power = first(In_Power),
+    Party = first(Party),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Log-transform the dependent variable to stabilize variance
+total_donations_cl <- total_donations_cl %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
+
+# Run the regression with interaction term
+model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)
+
 # Define the list of models
@@
 #| label: fig-estimates
-#| label: fig-model-estimates
 #| fig-cap: "Coefficient Estimates for Ontario, Federal, and Conservative/Liberal Models"

### 2024-12-04 paper/paper.qmd
update idealized methodology

@@
 
-Political donations are an important part of democratic processes, providing essential financial support for political campaigns and influencing electoral competition.
-In Canada, both federal and provincial elections rely on contributions from individuals and organizations.
-This can shape the political landscape and reflect public support for different parties.
-Understanding the factors that drive donation patterns is crucial for ensuring transparency and fairness in the democratic process and for minimizing the risk of undue influence on policy-making.
-
-A key question in political finance is whether a party's position in power affects the amount of financial support it receives.
-Studies in other countries, such as the United States, have shown that incumbents often benefit from more donations, particularly from interest groups seeking to maintain influence (@fouirnaies2014financial).
-However, there has been less focus on whether this holds true in Canada, where the political system and donation regulations differ significantly.
-This paper aims to fill that gap by analyzing how a party's power status influences donor behaviour in federal and Ontario provincial elections.
-
-The primary estimand of this study is the effect of a political party's power status on the total amount of financial donations it receives from individual contributors.
-By comparing the financial support received by parties in power versus those in opposition, this research aims to understand whether being in power affects a party's fundraising success.
-
-Using donation data from 2013 to 2024, this study examines total donations by year and party, accounting for factors such as election timing and party size (major party or not).
-The analysis employs linear regression models to quantify the relationship between power status and donation amounts, providing insights into donor motivations and the strategic considerations of political parties.
-
-The results indicate that opposition parties tend to receive higher total donations than those in power, suggesting that donors often prioritize supporting challengers, possibly to influence political change.
-This information is valuable not only for political parties developing fundraising strategies but also for policymakers and researchers seeking to ensure fairness and accountability in political financing.
-
-The paper is organized as follows: Section \ref{sec-data} discusses the data sources and methodology, including measurement and data cleaning processes.
-Section \ref{sec-model} presents the regression models used in the analysis, along with justification and validation.
-Section \ref{sec-results} outlines the results, and Section \ref{sec-discussion} discusses the implications, limitations, and suggestions for future research.
-Section \ref{sec-appendix} offers further detailed insights into the data, modeling approach, and methodology.
+Political donations are an important part of democratic processes,
+providing essential financial support for political campaigns and
+influencing electoral competition. In Canada, both federal and
+provincial elections rely on contributions from individuals and
+organizations. This can shape the political landscape and reflect public
+support for different parties. Understanding the factors that drive
+donation patterns is crucial for ensuring transparency and fairness in
+the democratic process and for minimizing the risk of undue influence on
+policy-making.
+
+A key question in political finance is whether a party's position in
+power affects the amount of financial support it receives. Studies in
+other countries, such as the United States, have shown that incumbents
+often benefit from more donations, particularly from interest groups
+seeking to maintain influence (@fouirnaies2014financial). However, there
+has been less focus on whether this holds true in Canada, where the
+political system and donation regulations differ significantly. This
+paper aims to fill that gap by analyzing how a party's power status
+influences donor behaviour in federal and Ontario provincial elections.
+
+The primary estimand of this study is the effect of a political party's
+power status on the total amount of financial donations it receives from
+individual contributors. By comparing the financial support received by
+parties in power versus those in opposition, this research aims to
+understand whether being in power affects a party's fundraising success.
+
+Using donation data from 2013 to 2024, this study examines total
+donations by year and party, accounting for factors such as election
+timing and party size (major party or not). The analysis employs linear
+regression models to quantify the relationship between power status and
+donation amounts, providing insights into donor motivations and the
+strategic considerations of political parties.
+
+The results indicate that opposition parties tend to receive higher
+total donations than those in power, suggesting that donors often
+prioritize supporting challengers, possibly to influence political
+change. This information is valuable not only for political parties
+developing fundraising strategies but also for policymakers and
+researchers seeking to ensure fairness and accountability in political
+financing.
+
+The paper is organized as follows: Section \ref{sec-data} discusses the
+data sources and methodology, including measurement and data cleaning
+processes. Section \ref{sec-model} presents the regression models used
+in the analysis, along with justification and validation. Section
+\ref{sec-results} outlines the results, and Section \ref{sec-discussion}
+discusses the implications, limitations, and suggestions for future
+research. Section \ref{sec-appendix} offers further detailed insights
+into the data, modeling approach, and methodology.
 
@@
 ## Overview
-The dataset used in this analysis comprises political donation records from 2013 to 2024, covering both the federal level and the province of Ontario in Canada.
-These records were sourced from the @ijf_political_donors Political Donors Dataset, which systematically collects and maintains comprehensive data on political donations across Canadian jurisdictions.
-The IJF updates this dataset daily by monitoring election agency websites at the federal level and within each province and territory, ensuring the data remains current and reliable.
-The analysis was conducted using both R (@citeR) and Python (@python), leveraging a range of libraries for data manipulation and visualization.
-In R, key packages included dplyr (@dplyr), lubridate (@lubridate), tidyverse (@tidyverse), ggplot2 (@ggplot), readr (@readr), tibble (@tibble), and rstanarm (@rstanarm) for statistical modeling.
-In Python, pandas (@pandas), requests (@requests), and matplotlib (@matplotlib) were utilized for additional data processing and visualization tasks
-
-Political donations in Canada are subject to legal disclosure requirements, promoting transparency and accountability within the democratic process.
-Both federal and provincial election finance laws mandate that political parties, candidates, and associated entities report donations exceeding specific thresholds.
-These disclosures vary by jurisdiction in terms of frequency and reporting formats.
-For instance, Ontario requires annual returns for general and by-elections, leadership races, and donations to parties, candidates, and constituency associations, all available in downloadable spreadsheet formats from 2007 onward.
-This regulatory framework ensures that significant contributions are systematically documented and publicly accessible, providing a reliable foundation for this analysis.
+
+The dataset used in this analysis comprises political donation records
+from 2013 to 2024, covering both the federal level and the province of
+Ontario in Canada. These records were sourced from the
+@ijf_political_donors Political Donors Dataset, which systematically
+collects and maintains comprehensive data on political donations across
+Canadian jurisdictions. The IJF updates this dataset daily by monitoring
+election agency websites at the federal level and within each province
+and territory, ensuring the data remains current and reliable. The
+analysis was conducted using both R (@citeR) and Python (@python),
+leveraging a range of libraries for data manipulation and visualization.
+In R, key packages included dplyr (@dplyr), lubridate (@lubridate),
+tidyverse (@tidyverse), ggplot2 (@ggplot), readr (@readr), tibble
+(@tibble), and rstanarm (@rstanarm) for statistical modeling. In Python,
+pandas (@pandas), requests (@requests), and matplotlib (@matplotlib)
+were utilized for additional data processing and visualization tasks
+
+Political donations in Canada are subject to legal disclosure
+requirements, promoting transparency and accountability within the
+democratic process. Both federal and provincial election finance laws
+mandate that political parties, candidates, and associated entities
+report donations exceeding specific thresholds. These disclosures vary
+by jurisdiction in terms of frequency and reporting formats. For
+instance, Ontario requires annual returns for general and by-elections,
+leadership races, and donations to parties, candidates, and constituency
+associations, all available in downloadable spreadsheet formats from
+2007 onward. This regulatory framework ensures that significant
+contributions are systematically documented and publicly accessible,
+providing a reliable foundation for this analysis.
 
@@
 
-The dataset originates from @ijf_political_donors compilation of political donation records, which includes donations made to registered political parties, party leadership contestants, riding associations, and individual candidates.
-Donations are tracked across different types of electoral events, including general elections, by-elections, and leadership races.
-The IJF collects this data by monitoring official elections agency websites at both the federal level and within each province and territory.
-
-The target population for this analysis includes all political donations made to registered political parties, party leadership contestants, riding associations (also known as electoral district or constituency associations), and individual riding candidates during the specified time frame.
-This comprehensive coverage allows for an in-depth examination of donation patterns in various political contexts.
-The data collection methodology ensures that all significant financial contributions are captured, providing a robust foundation for analyzing the relationship between political power and financial support.
-
-Data aggregation involves summing the total donations received by each political party annually, resulting in a dataset where each row represents the total donations for a specific party in a given year.
-Key variables include Political Party (categorizing major and minor parties), Donation Year (ranging from 2013 to 2024), and Amount Donated (the monetary value of each donation).
-Additional binary variables such as Recipient_in_Power (indicating whether the party was in power during the donation year), Party_Size (distinguishing major from minor parties based on legislative representation), and Election_Year (indicating whether the donation was made during an election year) are constructed to facilitate regression analyses.
-
-The data is organized into a structured format with each row detailing a specific donation event, including the party receiving the donation, the year it was made, the amount donated, and the constructed binary indicators.
-By aggregating donations at the party-year level and incorporating relevant variables, the dataset effectively translates real-world donation phenomena into analyzable data points, enabling the investigation of how political power status influences financial support in Canadian elections.
+The dataset originates from @ijf_political_donors compilation of
+political donation records, which includes donations made to registered
+political parties, party leadership contestants, riding associations,
+and individual candidates. Donations are tracked across different types
+of electoral events, including general elections, by-elections, and
+leadership races. The IJF collects this data by monitoring official
+elections agency websites at both the federal level and within each
+province and territory.
+
+The target population for this analysis includes all political donations
+made to registered political parties, party leadership contestants,
+riding associations (also known as electoral district or constituency
+associations), and individual riding candidates during the specified
+time frame. This comprehensive coverage allows for an in-depth
+examination of donation patterns in various political contexts. The data
+collection methodology ensures that all significant financial
+contributions are captured, providing a robust foundation for analyzing
+the relationship between political power and financial support.
+
+Data aggregation involves summing the total donations received by each
+political party annually, resulting in a dataset where each row
+represents the total donations for a specific party in a given year. Key
+variables include Political Party (categorizing major and minor
+parties), Donation Year (ranging from 2013 to 2024), and Amount Donated
+(the monetary value of each donation). Additional binary variables such
+as Recipient_in_Power (indicating whether the party was in power during
+the donation year), Party_Size (distinguishing major from minor parties
+based on legislative representation), and Election_Year (indicating
+whether the donation was made during an election year) are constructed
+to facilitate regression analyses.
+
+The data is organized into a structured format with each row detailing a
+specific donation event, including the party receiving the donation, the
+year it was made, the amount donated, and the constructed binary
+indicators. By aggregating donations at the party-year level and
+incorporating relevant variables, the dataset effectively translates
+real-world donation phenomena into analyzable data points, enabling the
+investigation of how political power status influences financial support
+in Canadian elections.
 
@@
 
-The raw donation records were carefully preprocessed to ensure data integrity before analysis.
-First, the data was imported using the read_csv function from the readr package.
-Records with missing or incomplete information in key variables were excluded to maintain quality.
-Political party names were standardized to address inconsistencies, such as consolidating variations like "Liberal Party of Ontario" and "Ontario Liberal Party" under a single name.
-Donations were then aggregated at the party-year level to analyze total annual contributions received by each party.
-Binary variables, including Recipient_in_Power, Party_Size, and Election_Year, were created to support regression analysis.
-To normalize the distribution and handle zero values, a new variable, Log_Total_Donations, was generated by applying a logarithmic transformation to Total_Donations + 1.
-Finally, the cleaned dataset was cross-validated against the original records to ensure accuracy.
-Additional details about the data cleaning process can be found in Appendix \ref{sec-appendix}.
+The raw donation records were carefully preprocessed to ensure data
+integrity before analysis. First, the data was imported using the
+read_csv function from the readr package. Records with missing or
+incomplete information in key variables were excluded to maintain
+quality. Political party names were standardized to address
+inconsistencies, such as consolidating variations like "Liberal Party of
+Ontario" and "Ontario Liberal Party" under a single name. Donations were
+then aggregated at the party-year level to analyze total annual
+contributions received by each party. Binary variables, including
+Recipient_in_Power, Party_Size, and Election_Year, were created to
+support regression analysis. To normalize the distribution and handle
+zero values, a new variable, Log_Total_Donations, was generated by
+applying a logarithmic transformation to Total_Donations + 1. Finally,
+the cleaned dataset was cross-validated against the original records to
+ensure accuracy. Additional details about the data cleaning process can
+be found in Appendix \ref{sec-appendix}.
 
@@
 
-The primary outcome variable is the Total Amount Donated, representing the total monetary value of donations received by each political party in a given year.
-This continuous variable is essential for assessing the level of financial support and understanding its relationship with factors such as power status and party size.
-
-The distribution of total donation amounts varies between the federal and provincial levels.
-At the federal level, the average total donations are higher, reflecting broader donor bases and different legal contribution limits.
+The primary outcome variable is the Total Amount Donated, representing
+the total monetary value of donations received by each political party
+in a given year. This continuous variable is essential for assessing the
+level of financial support and understanding its relationship with
+factors such as power status and party size.
+
+The distribution of total donation amounts varies between the federal
+and provincial levels. At the federal level, the average total donations
+are higher, reflecting broader donor bases and different legal
+contribution limits.
 
@@
 
--   Party_Size: Another binary variable which categorizes parties as major (1)
-    or minor (0) based on their representation in the legislature. Major
-    parties are expected to receive more donations due to greater
-    visibility, established support bases, and broader outreach
-    capabilities. This variable helps in distinguishing the influence of
-    party prominence on financial support.
+-   Party_Size: Another binary variable which categorizes parties as
+    major (1) or minor (0) based on their representation in the
+    legislature. Major parties are expected to receive more donations
+    due to greater visibility, established support bases, and broader
+    outreach capabilities. This variable helps in distinguishing the
+    influence of party prominence on financial support.
 
@@
 
-To analyze the impact of a political party's power status on donation patterns, we employed linear regression models for both Ontario and federal levels:
+To analyze the impact of a political party's power status on donation
+patterns, we employed linear regression models for both Ontario and
+federal levels:
 
@@
 ```
-
-For the Conservative and Liberal parties, we included an interaction term:
+For the Conservative and Liberal parties, we included an interaction
+term:
+
 ```{=tex}
@@
 - \(\text{Party}_i\): A binary variable indicating 1 for Conservative, 0 for Liberal.\\
-
 ```
-
-Each model uses the logarithm of total donations (Log_Total_Donations) as the dependent variable to stabilize variance and normalize the distribution of donation amounts.
-Predictor variables include binary indicators for whether the party was in power (In_Power), the size of the party (Party_Size), and whether the donation was made during an election year (Election_Year).
-The specialized model for Conservative and Liberal parties includes an additional binary variable (Party) and an interaction term (In_Power * Party) to capture the combined effect of power status and party type on donation amounts.
+Each model uses the logarithm of total donations (Log_Total_Donations)
+as the dependent variable to stabilize variance and normalize the
+distribution of donation amounts. Predictor variables include binary
+indicators for whether the party was in power (In_Power), the size of
+the party (Party_Size), and whether the donation was made during an
+election year (Election_Year). The specialized model for Conservative
+and Liberal parties includes an additional binary variable (Party) and
+an interaction term (In_Power \* Party) to capture the combined effect
+of power status and party type on donation amounts.
 
 ### Model justification
+
 Linear regression was chosen for its simplicity and interpretability.
-The log transformation of the dependent variable stabilizes variance and normalizes the distribution, making it suitable for linear modeling.
-The models account for key factors influencing donation amounts, allowing us to isolate the effect of a party being in power.
-
-Alternative models, such as generalized linear models, were considered but deemed unnecessary due to the adequacy of linear regression for this analysis.
+The log transformation of the dependent variable stabilizes variance and
+normalizes the distribution, making it suitable for linear modeling. The
+models account for key factors influencing donation amounts, allowing us
+to isolate the effect of a party being in power.
+
+Alternative models, such as generalized linear models, were considered
+but deemed unnecessary due to the adequacy of linear regression for this
+analysis.
 
 ### Assumptions and Limitations
-The models assume a linear relationship between the predictors and the logarithm of total donations, ensuring that the effects of the predictors are additive and proportional.
-Additionally, they assume that the residuals are normally distributed (normality of errors) and exhibit constant variance (homoscedasticity).
-The models also rely on the absence of multicollinearity, meaning that the predictors are not highly correlated, and they assume independence of observations, where each data point is independent of the others.
-Potential limitations of these models include the possibility of unobserved variables influencing donation amounts and the inherent assumption that past trends can reliably predict future behavior.
-
-### Model Validation ******
-Model validation was conducted to assess the predictive accuracy and generalizability of the regression models.
-This involved evaluating key metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square Error (RMSE) to determine how well the models explain the variance in Log_Total_Donations and their predictive performance.
-
-For the Ontario Model, the regression analysis yielded an R² of 0.645 and an Adjusted R² of 0.635, indicating that approximately 64.5% of the variance in log-transformed total donations is explained by the predictors.
-The RMSE for this model was calculated to assess the average prediction error, demonstrating a reasonable level of accuracy.
-
-The Federal Model showed an R² of 0.515 and an Adjusted R² of 0.503, suggesting that 51.5% of the variance in log-transformed total donations is accounted for by the predictors.
-The RMSE value indicated acceptable predictive performance, though slightly lower than the Ontario model, reflecting potential differences in donation dynamics at the federal level.
-
-In the Conservative and Liberal Parties Model, the inclusion of the interaction term between In_Power and Party resulted in an R² of 0.630 and an Adjusted R² of 0.600.
-This model effectively captures the combined effect of power status and party type on donation amounts, with the interaction term providing significant insights into how being in power influences donations differently for Conservative and Liberal parties.
-The RMSE for this model confirmed its robust predictive capability, comparable to the other models.
-
-Overall, the validation process confirmed that the linear regression models are robust and effective in predicting donation amounts based on the selected predictor variables.
-The models demonstrate good explanatory power and reasonable predictive accuracy, making them suitable for analyzing the factors influencing political donations in Canada.
-However, ongoing evaluation and potential model refinement are recommended to accommodate evolving political and financial landscapes, ensuring the models remain relevant and accurate over time.
+
+The models assume a linear relationship between the predictors and the
+logarithm of total donations, ensuring that the effects of the
+predictors are additive and proportional. Additionally, they assume that
+the residuals are normally distributed (normality of errors) and exhibit
+constant variance (homoscedasticity). The models also rely on the
+absence of multicollinearity, meaning that the predictors are not highly
+correlated, and they assume independence of observations, where each
+data point is independent of the others. Potential limitations of these
+models include the possibility of unobserved variables influencing
+donation amounts and the inherent assumption that past trends can
+reliably predict future behavior.
+
+### Model Validation \*\*\*\*\*\*
+
+Model validation was conducted to assess the predictive accuracy and
+generalizability of the regression models. This involved evaluating key
+metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square
+Error (RMSE) to determine how well the models explain the variance in
+Log_Total_Donations and their predictive performance.
+
+For the Ontario Model, the regression analysis yielded an R² of 0.645
+and an Adjusted R² of 0.635, indicating that approximately 64.5% of the
+variance in log-transformed total donations is explained by the
+predictors. The RMSE for this model was calculated to assess the average
+prediction error, demonstrating a reasonable level of accuracy.
+
+The Federal Model showed an R² of 0.515 and an Adjusted R² of 0.503,
+suggesting that 51.5% of the variance in log-transformed total donations
+is accounted for by the predictors. The RMSE value indicated acceptable
+predictive performance, though slightly lower than the Ontario model,
+reflecting potential differences in donation dynamics at the federal
+level.
+
+In the Conservative and Liberal Parties Model, the inclusion of the
+interaction term between In_Power and Party resulted in an R² of 0.630
+and an Adjusted R² of 0.600. This model effectively captures the
+combined effect of power status and party type on donation amounts, with
+the interaction term providing significant insights into how being in
+power influences donations differently for Conservative and Liberal
+parties. The RMSE for this model confirmed its robust predictive
+capability, comparable to the other models.
+
+Overall, the validation process confirmed that the linear regression
+models are robust and effective in predicting donation amounts based on
+the selected predictor variables. The models demonstrate good
+explanatory power and reasonable predictive accuracy, making them
+suitable for analyzing the factors influencing political donations in
+Canada. However, ongoing evaluation and potential model refinement are
+recommended to accommodate evolving political and financial landscapes,
+ensuring the models remain relevant and accurate over time.
 
@@
 
-Table @tbl-modelsummary presents the summary of key model estimates for the Ontario, Federal, and Conservative/Liberal models.
-These models predict the logarithm of total donations received by political parties based on whether the party was in power, the size of the party, and whether the donation occurred during an election year.
-The Conservative/Liberal model includes an interaction term to explore how the effect of being in power varies between these two major parties.
+Table @tbl-modelsummary presents the summary of key model estimates for
+the Ontario, Federal, and Conservative/Liberal models. These models
+predict the logarithm of total donations received by political parties
+based on whether the party was in power, the size of the party, and
+whether the donation occurred during an election year. The
+Conservative/Liberal model includes an interaction term to explore how
+the effect of being in power varies between these two major parties.
 
@@
 
-The regression analysis reveals that Party Size is a significant predictor of total donations across both the Ontario and Federal models, indicating that larger parties receive more financial support.
-Specifically, in the Ontario model, Party Size has a coefficient of 2.302 (p < 0.001), while in the Federal model, it is 3.524 (p < 0.001).
-This positive relationship suggests that party prominence plays a crucial role in attracting donations.
-
-Conversely, the In Power variable is not statistically significant in the Ontario (p = 0.322) and Federal (p = 0.529) models, implying that being in power does not directly influence the total donations received by a party at these levels.
-However, in the Conservative/Liberal model, In Power has a significant positive effect (2.513, p < 0.001), and the interaction term In Power * Party is significantly negative (-4.865, p < 0.001).
-This interaction indicates that while being in power increases donations for Conservative parties, it decreases donations for Liberal parties, highlighting a differential impact based on party affiliation.
-
-The Election Year variable does not significantly predict donation amounts in any of the models, with p-values exceeding conventional significance thresholds (Ontario p = 0.616, Federal p = 0.975, Conservative/Liberal p = 0.325).
-This suggests that donations are not substantially influenced by whether they occur during an election year.
-
-Overall, the models explain a moderate portion of the variance in total donations, with R² values ranging from 0.3046 (Ontario) to 0.5825 (Federal), and the Conservative/Liberal model achieving an R² of 0.4481.
-These findings underscore the importance of party size in determining donation amounts, while the role of power status appears to be contingent on party affiliation.
+The regression analysis reveals that Party Size is a significant
+predictor of total donations across both the Ontario and Federal models,
+indicating that larger parties receive more financial support.
+Specifically, in the Ontario model, Party Size has a coefficient of
+2.302 (p \< 0.001), while in the Federal model, it is 3.524 (p \<
+0.001). This positive relationship suggests that party prominence plays
+a crucial role in attracting donations.
+
+Conversely, the In Power variable is not statistically significant in
+the Ontario (p = 0.322) and Federal (p = 0.529) models, implying that
+being in power does not directly influence the total donations received
+by a party at these levels. However, in the Conservative/Liberal model,
+In Power has a significant positive effect (2.513, p \< 0.001), and the
+interaction term In Power \* Party is significantly negative (-4.865, p
+\< 0.001). This interaction indicates that while being in power
+increases donations for Conservative parties, it decreases donations for
+Liberal parties, highlighting a differential impact based on party
+affiliation.
+
+The Election Year variable does not significantly predict donation
+amounts in any of the models, with p-values exceeding conventional
+significance thresholds (Ontario p = 0.616, Federal p = 0.975,
+Conservative/Liberal p = 0.325). This suggests that donations are not
+substantially influenced by whether they occur during an election year.
+
+Overall, the models explain a moderate portion of the variance in total
+donations, with R² values ranging from 0.3046 (Ontario) to 0.5825
+(Federal), and the Conservative/Liberal model achieving an R² of 0.4481.
+These findings underscore the importance of party size in determining
+donation amounts, while the role of power status appears to be
+contingent on party affiliation.
 
@@
 
-##The Influence of Party Size on Donation Amounts {#sec-first-point}
-At the heart of politics is finance, as revealed by Traag's paper on the complex contagion of campaign donations @traag2016complex.
-Their study demonstrates that a majority of campaign contributions in the US originate from a few wealthy individuals through large social networks.
-With the constant battle between government and corporations, the findings in this study underscore the significant influence of financial power in shaping political dynamics.\\
+## The Influence of Party Size on Donation Amounts {#sec-first-point}
+
+At the heart of politics is finance, as revealed by Traag's paper on the
+complex contagion of campaign donations @traag2016complex. Their study
+demonstrates that a majority of campaign contributions in the US
+originate from a few wealthy individuals through large social networks.
+With the constant battle between government and corporations, the
+findings in this study underscore the significant influence of financial
+power in shaping political dynamics.
 
 The landscape in Canada for political financing is growing each year.
-Results from studies demonstrate that individual contributors may prefer to support challengers for several reasons.
-An optimistic but naive explanation may be that individuals hope to foster competitive elections, given that donations depend not on political leaning but on who is currently in power.
-However, a more realistic explanation for this phenomenon is that some individuals want to make policy changes.
-It is no secret that politicians, particularly those out of power, often solicit funding and political support in exchange for advancing the interests of their stakeholders @mollona2022ethics.
-These two dynamics may be working in tandem to create an effect where parties eager to be elected rely more on funding from external sources, which can switch each election cycle.
-
-Moreover, these trends of political donations fluctuating based on the party in power are not unique to Canada.
-The study "Financing UK Democracy: A Stocktake of 20 Years of Political Donations" highlights a surge in private political donations, more specifically from superdonors, widening the resource gap between large parties @draca2022financing.
-Similar to Canada, the UK also sees large donations from a wealthy few influencing their elections each year.
-The parallel between Canada and the UK underscores a broader democratic issue of financial power being used to influence political outcomes.
-Though this concept is not novel, this study contributes to the already growing spotlight on the persistent lack of proper reform in political systems.
-
-This study provides valuable insights into donation behaviours across Canada but it its limitations.
-First, the analysis is applied to only federal and provincial donation records, overlooking regional variations found in other provinces and territories.
-The geographical limitations of this study hinder its ability to generalize across the entire country.
-Moreover, the study only spans from 2006 to 2024 which economic or political events could have influence over the donation patterns that the models do not account for.
+Results from studies demonstrate that individual contributors may prefer
+to support challengers for several reasons. An optimistic but naive
+explanation may be that individuals hope to foster competitive
+elections, given that donations depend not on political leaning but on
+who is currently in power. However, a more realistic explanation for
+this phenomenon is that some individuals want to make policy changes. It
+is no secret that politicians, particularly those out of power, often
+solicit funding and political support in exchange for advancing the
+interests of their stakeholders @mollona2022ethics. These two dynamics
+may be working in tandem to create an effect where parties eager to be
+elected rely more on funding from external sources, which can switch
+each election cycle.
+
+Moreover, these trends of political donations fluctuating based on the
+party in power are not unique to Canada. The study "Financing UK
+Democracy: A Stocktake of 20 Years of Political Donations" highlights a
+surge in private political donations, more specifically from
+superdonors, widening the resource gap between large parties
+@draca2022financing. Similar to Canada, the UK also sees large donations
+from a wealthy few influencing their elections each year. The parallel
+between Canada and the UK underscores a broader democratic issue of
+financial power being used to influence political outcomes. Though this
+concept is not novel, this study contributes to the already growing
+spotlight on the persistent lack of proper reform in political systems.
+
+This study provides valuable insights into donation behaviours across
+Canada but it its limitations. First, the analysis is applied to only
+federal and provincial donation records, overlooking regional variations
+found in other provinces and territories. The geographical limitations
+of this study hinder its ability to generalize across the entire
+country. Moreover, the study only spans from 2006 to 2024 which economic
+or political events could have influence over the donation patterns that
+the models do not account for.
 
@@
 ## Appendix A: Data Cleaning Notes
-The dataset used in this analysis underwent a rigorous data cleaning process to ensure accuracy, consistency, and reliability.
-The raw donation records were sourced from the Investigative Journalism Foundation (IJF), encompassing political donations at both the federal and Ontario provincial levels from 2013 to 2024.
-The data cleaning process involved several critical steps, detailed below.
-
-Initially, the dataset was examined for missing or incomplete values in key variables such as Amount Donated, Political Party, Donation Year, and Region.
-Records with missing values in these essential fields were excluded from the analysis to maintain data integrity and prevent biases that could arise from incomplete data entries.
-
-Political party names often appeared in various formats due to inconsistencies in reporting.
-To address this, party names were standardized across all records.
-For instance, variations such as "Liberal Party of Ontario" and "Ontario Liberal Party" were consolidated under the standardized name "Ontario Liberal Party." This standardization ensured accurate attribution of donations to the correct parties.
-
-Donor information, including names and locations, was also standardized where possible.
-Typographical errors were corrected, and formatting was made consistent (e.g., "First Name Last Name" instead of "Last Name, First Name").
-However, original donor names were retained in cases where standardization could introduce inaccuracies, particularly when dealing with individuals with similar names.
-
-The dataset included both monetary (amount_monetary) and non-monetary (amount_non_monetary) contributions.
-A new variable, Amount Donated, was created by summing these two fields for each record.
-This provided a comprehensive measure of the total value of donations received by each party.
-All monetary values were converted to a consistent currency format, and any discrepancies in currency symbols or formats were corrected to ensure accurate aggregation and comparison of donation amounts.
-
-To facilitate analysis at the party-year level, individual donation records were aggregated.
-Donations were summed for each political party within each year, resulting in a dataset where each entry represented the total donations received by a party in a specific year.
-This aggregation allowed for the examination of trends over time and assessment of the impact of factors such as power status and election years on donation amounts.
-
-Several binary variables were introduced to capture key aspects of the data:
-
-Recipient_in_Power: Indicates whether the recipient party was in power during the donation year (1 for in power, 0 for not in power).
-Party_Size: Categorizes parties as major (1) or minor (0) based on legislative representation and historical influence.
-Election_Year: Indicates whether the donation was made during an election year (1 for election year, 0 for non-election year).
-These variables were crucial for the regression analysis, allowing for the assessment of the influence of these factors on total donations.
-
-The distribution of total donation amounts was highly skewed, with a small number of large donations and many smaller ones.
-To address this skewness and stabilize variance, a logarithmic transformation was applied to the Total Donations variable, creating Log_Total_Donations.
-This transformation also helped in handling zero values by adding a small constant (e.g., 1) before taking the logarithm.
-
-To ensure the accuracy of the cleaned dataset, cross-validation was performed against original records and official financial statements where available.
-This step involved spot-checking aggregated totals and individual records to confirm that the cleaning and aggregation processes did not introduce errors.
-
-Some challenges were encountered due to inconsistencies in the raw data, such as typographical errors, inconsistent reporting formats, and missing donor types.
-While efforts were made to standardize and correct these issues, some limitations remain due to the quality of the original data sources.
-In cases where donor type information was missing and legal restrictions indicated that only individuals could donate (e.g., Ontario after 2017), the donor type was inferred as "Individual." While this assumption is reasonable based on the legal context, it may not capture all nuances in the data.
+
+The dataset used in this analysis underwent a rigorous data cleaning
+process to ensure accuracy, consistency, and reliability. The raw
+donation records were sourced from the Investigative Journalism
+Foundation (IJF), encompassing political donations at both the federal
+and Ontario provincial levels from 2013 to 2024. The data cleaning
+process involved several critical steps, detailed below.
+
+Initially, the dataset was examined for missing or incomplete values in
+key variables such as Amount Donated, Political Party, Donation Year,
+and Region. Records with missing values in these essential fields were
+excluded from the analysis to maintain data integrity and prevent biases
+that could arise from incomplete data entries.
+
+Political party names often appeared in various formats due to
+inconsistencies in reporting. To address this, party names were
+standardized across all records. For instance, variations such as
+"Liberal Party of Ontario" and "Ontario Liberal Party" were consolidated
+under the standardized name "Ontario Liberal Party." This
+standardization ensured accurate attribution of donations to the correct
+parties.
+
+Donor information, including names and locations, was also standardized
+where possible. Typographical errors were corrected, and formatting was
+made consistent (e.g., "First Name Last Name" instead of "Last Name,
+First Name"). However, original donor names were retained in cases where
+standardization could introduce inaccuracies, particularly when dealing
+with individuals with similar names.
+
+The dataset included both monetary (amount_monetary) and non-monetary
+(amount_non_monetary) contributions. A new variable, Amount Donated, was
+created by summing these two fields for each record. This provided a
+comprehensive measure of the total value of donations received by each
+party. All monetary values were converted to a consistent currency
+format, and any discrepancies in currency symbols or formats were
+corrected to ensure accurate aggregation and comparison of donation
+amounts.
+
+To facilitate analysis at the party-year level, individual donation
+records were aggregated. Donations were summed for each political party
+within each year, resulting in a dataset where each entry represented
+the total donations received by a party in a specific year. This
+aggregation allowed for the examination of trends over time and
+assessment of the impact of factors such as power status and election
+years on donation amounts.
+
+Several binary variables were introduced to capture key aspects of the
+data:
+
+Recipient_in_Power: Indicates whether the recipient party was in power
+during the donation year (1 for in power, 0 for not in power).
+Party_Size: Categorizes parties as major (1) or minor (0) based on
+legislative representation and historical influence. Election_Year:
+Indicates whether the donation was made during an election year (1 for
+election year, 0 for non-election year). These variables were crucial
+for the regression analysis, allowing for the assessment of the
+influence of these factors on total donations.
+
+The distribution of total donation amounts was highly skewed, with a
+small number of large donations and many smaller ones. To address this
+skewness and stabilize variance, a logarithmic transformation was
+applied to the Total Donations variable, creating Log_Total_Donations.
+This transformation also helped in handling zero values by adding a
+small constant (e.g., 1) before taking the logarithm.
+
+To ensure the accuracy of the cleaned dataset, cross-validation was
+performed against original records and official financial statements
+where available. This step involved spot-checking aggregated totals and
+individual records to confirm that the cleaning and aggregation
+processes did not introduce errors.
+
+Some challenges were encountered due to inconsistencies in the raw data,
+such as typographical errors, inconsistent reporting formats, and
+missing donor types. While efforts were made to standardize and correct
+these issues, some limitations remain due to the quality of the original
+data sources. In cases where donor type information was missing and
+legal restrictions indicated that only individuals could donate (e.g.,
+Ontario after 2017), the donor type was inferred as "Individual." While
+this assumption is reasonable based on the legal context, it may not
+capture all nuances in the data.
 
@@
 
-The Investigative Journalism Foundation (IJF) employs a comprehensive approach to collect and compile political donation data across Canada.
-Understanding the IJF's data collection methodology is essential for evaluating the reliability and scope of the data used in this analysis.
-
-Data Sources and Acquisition
-
-Every day, the IJF checks election agency websites at the federal level and in each province and territory for new political donations data.
-Historical data is obtained from a variety of government sources, including provincial archives, legislative libraries, and elections agencies, covering the period from 1993 to the present day, though this time range varies by jurisdiction.
-
-Political parties and candidates are legally required to submit records of donations, which are maintained and made publicly accessible by election agencies.
-The frequency and format of these disclosures differ among jurisdictions; some require quarterly returns, while others mandate annual reports.
-The IJF systematically gathers these records to create a unified database of political donations.
-
-Scope of Data Collected
-
-The IJF's dataset encompasses donations to various political entities, including:
-
-Parties: Registered political parties at the federal and provincial levels.
-Candidates: Individual candidates running for office.
-Riding Associations: Also known as electoral district or constituency associations.
-Leadership and Nomination Contestants: Individuals participating in party leadership races or seeking nomination as candidates.
-The dataset includes records from all 13 provinces and territories, as well as city-specific donations from municipalities like Victoria and Vancouver.
-
-Donation Laws and Contribution Limits
-
-Contribution limits and the types of entities permitted to donate vary across jurisdictions.
-For instance, at the federal level, only individuals who are Canadian citizens or permanent residents can donate, with maximum annual contributions set at specified amounts.
-In contrast, some provinces allow donations from corporations, unions, and other organizations and may have different or no contribution limits.
-
-Data Formats and Conversion
-
-The source data is available in multiple formats, including downloadable spreadsheets, PDFs, and HTML files.
-For jurisdictions where data is only available in PDF format, the IJF employs Optical Character Recognition (OCR) technology to convert the documents into machine-readable CSV files.
-OCR tools such as Adobe Export PDF, Cometdocs, and ABBYY FineReader are used for this purpose.
-
-Data Cleaning and Standardization
-
-Extensive manual cleaning is performed to correct errors arising from OCR conversion and to standardize the data. Key steps in the data cleaning process include:
-
-Correcting OCR Errors: Manual verification and correction of misread characters (e.g., misinterpreted currency symbols or letters).
-Standardizing Donor and Party Names: Ensuring consistency in naming conventions for donors and political parties (e.g., standardizing "Progressive Conservative Party of Ontario" across all records).
-Formatting Dates and Amounts: Standardizing date formats to YYYY-MM-DD and ensuring monetary values are consistently formatted.
-The IJF also amalgamates similar categories in variables such as donor types to improve data legibility (e.g., merging "Business" and "Corporation" into a single "Corporation" category).
-In cases where donor type information is missing and laws specify that only individuals can donate (e.g., Ontario after 2017), the IJF adds "Individual" to the donor type column.
-
-Limitations and Challenges
-
-While the IJF endeavors to provide accurate and comprehensive data, several limitations exist:
-
-Typos and Reporting Errors: Original records submitted by political entities may contain typographical errors or inaccuracies, which are reflected in the dataset.
-Incomplete or Inconsistent Data: Variations in reporting requirements and practices across jurisdictions can lead to inconsistencies in the data.
-OCR Limitations: Despite manual verification, OCR technology may not perfectly capture handwritten or poorly scanned documents, potentially introducing errors.
-Conclusion
-
-The IJF's meticulous data collection and cleaning processes result in a robust dataset suitable for analyzing political donation patterns in Canada.
-Understanding the methodologies employed allows for a better assessment of the data's strengths and limitations, ensuring analyses are grounded in a thorough appreciation of the underlying data quality.
+The Investigative Journalism Foundation (IJF) employs a comprehensive
+approach to collect and compile political donation data across Canada.
+Understanding the IJF's data collection methodology is essential for
+evaluating the reliability and scope of the data used in this analysis.
+
+**Data Sources and Acquisition**
+
+Every day, the IJF checks election agency websites at the federal level
+and in each province and territory for new political donations data.
+Historical data is obtained from a variety of government sources,
+including provincial archives, legislative libraries, and elections
+agencies, covering the period from 1993 to the present day, though this
+time range varies by jurisdiction.
+
+Political parties and candidates are legally required to submit records
+of donations, which are maintained and made publicly accessible by
+election agencies. The frequency and format of these disclosures differ
+among jurisdictions; some require quarterly returns, while others
+mandate annual reports. The IJF systematically gathers these records to
+create a unified database of political donations.
+
+**Scope of Data Collected**
+
+The IJF's dataset encompasses donations to various political entities,
+including:
+
+-   Parties: Registered political parties at the federal and provincial
+    levels.
+
+-   Candidates: Individual candidates running for office.
+
+-   Riding Associations: Also known as electoral district or
+    constituency associations.
+
+-   Leadership and Nomination Contestants: Individuals participating in
+    party leadership races or seeking nomination as candidates.
+
+The dataset includes records from all 13 provinces and territories, as
+well as city-specific donations from municipalities like Victoria and
+Vancouver.
+
+**Donation Laws and Contribution Limits**
+
+Contribution limits and the types of entities permitted to donate vary
+across jurisdictions. For instance, at the federal level, only
+individuals who are Canadian citizens or permanent residents can donate,
+with maximum annual contributions set at specified amounts. In contrast,
+some provinces allow donations from corporations, unions, and other
+organizations and may have different or no contribution limits.
+
+**Data Formats and Conversion**
+
+The source data is available in multiple formats, including downloadable
+spreadsheets, PDFs, and HTML files. For jurisdictions where data is only
+available in PDF format, the IJF employs Optical Character Recognition
+(OCR) technology to convert the documents into machine-readable CSV
+files. OCR tools such as Adobe Export PDF, Cometdocs, and ABBYY
+FineReader are used for this purpose.
+
+**Data Cleaning and Standardization**
+
+Extensive manual cleaning is performed to correct errors arising from
+OCR conversion and to standardize the data. Key steps in the data
+cleaning process include:
+
+Correcting OCR Errors: Manual verification and correction of misread
+characters (e.g., misinterpreted currency symbols or letters).
+Standardizing Donor and Party Names: Ensuring consistency in naming
+conventions for donors and political parties (e.g., standardizing
+"Progressive Conservative Party of Ontario" across all records).
+Formatting Dates and Amounts: Standardizing date formats to YYYY-MM-DD
+and ensuring monetary values are consistently formatted. The IJF also
+amalgamates similar categories in variables such as donor types to
+improve data legibility (e.g., merging "Business" and "Corporation" into
+a single "Corporation" category). In cases where donor type information
+is missing and laws specify that only individuals can donate (e.g.,
+Ontario after 2017), the IJF adds "Individual" to the donor type column.
+
+**Limitations and Challenges**
+
+Original records submitted by political entities may contain
+typographical errors or inaccuracies, which are reflected in the
+dataset. Incomplete or Inconsistent Data: Variations in reporting
+requirements and practices across jurisdictions can lead to
+inconsistencies in the data. OCR Limitations: Despite manual
+verification, OCR technology may not perfectly capture handwritten or
+poorly scanned documents, potentially introducing errors. Conclusion
 
@@
 
-1. **Comprehensive Data Sources:** 
-   - **Government Databases:** Utilize all available records from Elections Canada and provincial electoral bodies to ensure complete coverage of political donations.
-   - **Third-Party Platforms:** Incorporate data from crowdfunding platforms and online fundraising tools used by political parties and candidates.
-   - **Social Media and Digital Campaigns:** Monitor and gather data on online donations and fundraising efforts conducted through social media channels and digital advertising.
-
-2. **Enhanced Data Integration and Real-Time Updates:**
-   - **Automated Data Pipelines:** Implement automated systems to continuously collect and integrate data from multiple sources in real-time, reducing latency and ensuring up-to-date information.
-   - **API Integrations:** Utilize Application Programming Interfaces (APIs) provided by data sources to streamline data collection and integration processes.
-
-3. **Advanced Data Verification Techniques:**
-   - **Machine Learning Algorithms:** Employ machine learning models to detect anomalies, identify fraudulent donations, and ensure data accuracy.
-   - **Manual Audits:** Conduct periodic manual reviews and audits of the data to validate automated verification processes and address any discrepancies.
-
-4. **Detailed Categorization and Enrichment:**
-   - **Donor Demographics:** Collect additional demographic information about donors, such as age, occupation, and geographic location, to enable more granular analysis.
-   - **Donation Purpose:** Categorize donations based on their intended purpose (e.g., general support, specific campaigns, issue-based funding) to understand donor motivations.
-
-5. **Robust Data Cleaning and Standardization:**
-   - **Automated Cleaning Tools:** Utilize advanced data cleaning software to automatically detect and correct errors, handle missing values, and standardize data formats.
-   - **Standardized Taxonomies:** Develop and apply standardized taxonomies for categorizing donations and donor attributes, ensuring consistency across the dataset.
-
-6. **Ethical and Transparent Practices:**
-   - **Data Privacy Compliance:** Strictly adhere to data privacy regulations, ensuring that all personal information is securely stored and processed.
-   - **Transparency Reports:** Publish regular transparency reports detailing data collection methods, sources, and any limitations to build trust with stakeholders.
-
-7. **Comprehensive Documentation and Accessibility:**
-   - **Detailed Documentation:** Maintain thorough documentation of all data collection, processing, and analysis procedures to facilitate reproducibility and accountability.
-   - **Open Data Access:** Provide open access to anonymized datasets and analysis tools, allowing researchers and the public to engage with the data and conduct independent analyses.
-
-By implementing these enhancements, the methodology would not only improve the quality and reliability of the data but also enable more sophisticated analyses, leading to deeper insights into political donation dynamics.
+The idealized methodology for analyzing how a political party's power
+status affects the total donations it receives in Canada involves
+conducting a comprehensive survey targeting individual political donors
+across the country. This methodology aims to gather detailed data on
+donor behaviours, motivations, and perceptions, providing deeper
+insights into the factors influencing political donations at both
+federal and Ontario provincial levels.
+
+**Target Population**
+
+The target population consists of individuals who have made financial
+contributions to Canadian federal or provincial political parties,
+candidates, riding associations, or leadership contestants between 2013
+and 2024. This includes donors from all provinces and territories,
+ensuring representation across different regions and political contexts.
+
+**Sampling Method**
+
+A stratified random sampling technique will be utilized to ensure
+proportional representation from various demographics and regions. The
+sample will be stratified based on:
+
+-   Region: All provinces and territories.
+
+-   Donor Type: Donors to parties in power versus opposition parties.
+
+-   Demographics: Age, gender, income level, education level, and
+    political affiliation.
+
+In Ontario and other populous provinces, oversampling will be conducted
+to capture more granular data and reduce the margin of error. This
+approach ensures that key voter segments are adequately represented,
+particularly those in regions with significant political activity.
+
+**Sample Size and Response Rate**
+
+To maintain a national margin of error of ±3% at a 95% confidence level,
+a total sample size of approximately 1,500 respondents is required. This
+sample size allows for meaningful analysis of subgroups within the data.
+Given an estimated response rate of 10%, outreach efforts will target
+approximately 15,000 potential respondents across Canada.
+
+**Recruitment Methods**
+
+Respondents will be recruited using a combination of methods to maximize
+reach and diversity:
+
+-   Online Panel Providers: Collaborate with established online panel
+    providers such as Ipsos, Leger, or Angus Reid to access a pool of
+    verified respondents who have consented to participate in surveys.
+
+-   Email Invitations: Where legally permissible and ethically
+    acceptable, utilize contact information from public donor records to
+    send email invitations directly to known political donors.
+
+-   Social Media Advertising: Use targeted advertisements on platforms
+    like Facebook, Twitter, and LinkedIn to reach potential respondents
+    interested in political activities.
+
+-   Telephone Outreach: Conduct telephone surveys using interactive
+    voice response (IVR) systems and live calls to reach donors who may
+    not be active online, particularly older individuals or those in
+    rural areas.
+
+To incentivize participation, all respondents will be offered a small
+monetary compensation (e.g., a \$10 electronic gift card) or entered
+into a national sweepstakes with a chance to win one of ten \$500 cash
+prizes.
+
+**Data Collection Methods**
+
+Data collection will be conducted via:
+
+-   Online Surveys utilizing platforms like Qualtrics or SurveyMonkey to
+    administer the survey electronically. The survey will be designed to
+    be mobile-friendly and accessible to individuals with disabilities.
+
+-   Telephone Surveys for respondents preferring or requiring this
+    method, trained interviewers will administer the survey, ensuring
+    consistency with the online questionnaire.
+
+The survey is designed to take approximately 10 to 15 minutes to
+complete, minimizing drop-off rates while collecting comprehensive data.
+
+**Data Validation Measures**
+
+Several measures will be implemented to ensure data quality and
+integrity:
+
+-   Completion time checks will be used monitor the time respondents
+    take to complete the survey to identify and exclude those who rush
+    through it (speeding).
+
+-   Implement reCAPTCHA technology to guard against bots and automated
+    responses.
+
+-   nclude attention-check questions to identify inattentive or
+    fraudulent responses.
+
+-   After data collection, apply statistical weighting to adjust for any
+    demographic discrepancies, ensuring the sample accurately reflects
+    the population of political donors in Canada.
+
+**Budget Allocation**
+
+The budget for this study is allocated as follows:
+
+-   Survey Development: \$8,000
+
+    -   This covers costs associated with designing the survey,
+        including question formulation, structuring for clarity and
+        flow, pilot testing, and ensuring compliance with ethical
+        standards. Additionally, funds account for bilingual translation
+        services (English and French).
+
+-   Sampling and Recruitment: \$15,000
+
+-   Details: Expenses related to accessing online panel respondents,
+    purchasing targeted advertising on social media, and collaborating
+    with panel providers. Funds also cover the services of statisticians
+    to ensure the sampling aligns with the population distribution.
+
+-   Respondent Incentives: \$20,000
+
+-   Details: Allocated for compensating participants with \$10
+    electronic gift cards or funding the national sweepstakes prizes.
+    Incentives are crucial for improving response rates and attracting a
+    diverse respondent pool.
+
+-   Data Collection: \$25,000
+
+-   Details: Operational costs for administering the survey, including
+    online platform fees, telephone survey expenses, interviewer
+    training, and call center operations. This ensures data collection
+    is efficient and reaches a broad audience.
+
+-   Data Analysis and Reporting: \$12,000
+
+-   Details: Funds dedicated to analyzing the collected data, including
+    hiring data analysts proficient in statistical software, conducting
+    comprehensive analyses, and preparing detailed reports and
+    visualizations to communicate findings effectively.
+
+-   Ethical Compliance and Data Security: \$5,000
+
+-   Details: Covers the costs of obtaining ethical approvals, ensuring
+    compliance with privacy legislation (e.g., Personal Information
+    Protection and Electronic Documents Act), and implementing robust
+    data security measures to protect respondent information.
+
+-   Total Estimated Budget: \$85,000
+
+Timeline
+
+-   Months 1-2: Survey design, ethical approval processes, and pilot
+    testing to refine the questionnaire.
+
+-   Months 3-4: Sampling frame development and recruitment of
+    respondents through various channels.
+
+-   Months 5-6: Data collection via online and telephone surveys.
+
+-   Month 7: Data cleaning, validation, and preliminary analysis.
+
+-   Months 8-9: Comprehensive data analysis and report preparation.
+
+-   Month 10: Dissemination of findings through academic publications
+    and presentations.
 
@@
 
-The survey should be meticulously designed to ensure high response rates, representativeness, and data quality. Below is an outline of an idealized survey for studying political donations:
-
-Survey Objectives
-- **Understand Donor Motivations:** Explore the reasons behind individuals and organizations choosing to donate to specific political parties.
-- **Assess Donation Patterns:** Identify patterns in donation amounts, frequency, and timing relative to election cycles.
-- **Evaluate Impact of Campaigns:** Determine how political campaigns and events influence donation behaviors.
-- **Gather Demographic Insights:** Collect demographic information to analyze how factors such as age, income, education, and geographic location affect donation decisions.
-
 Survey Structure
@@
 **Section 1: Demographics**
-- **State of Residence:** Dropdown list of all provinces and territories.
-- **Age:** Multiple-choice options (e.g., 18-24, 25-34, etc.).
-- **Gender:** Options including Male, Female, Non-binary/Other, Prefer not to say.
-- **Income Level:** Income brackets (e.g., Less than $25,000, $25,000-$49,999, etc.).
-- **Education Level:** Highest degree obtained (e.g., High school, Bachelor’s, Master’s, etc.).
-- **Occupation:** Open-ended or categorized options.
-- **Political Affiliation:** Conservative, Liberal, NDP, Other.
-
-**Section 2: Donation Behavior**
-- **Are you currently registered to donate to political parties?**
-  - Yes
-  - No
-- **How frequently do you donate to political parties?**
-  - Never
-  - Rarely (once a year or less)
-  - Occasionally (a few times a year)
-  - Regularly (monthly or more)
-- **Average Donation Amount:** Slider or multiple-choice options (e.g., Less than $50, $50-$199, etc.).
-- **Preferred Donation Channels:** Multiple selections (e.g., Online platforms, Mail-in donations, In-person events, etc.).
-
-**Section 3: Motivations for Donating**
-- **What motivates you to donate to a political party?** (Select all that apply)
-  - Support for specific policies or issues
-  - Desire to influence election outcomes
-  - Loyalty to the party or its leaders
-  - Social pressure or community influence
-  - Tax benefits
-  - Other (please specify)
-- **Which factors most influence your decision to donate?** (Rank in order of importance)
-  - Candidate's leadership qualities
-  - Party's stance on key issues
-  - Party's past performance
-  - Recommendations from peers or influencers
-  - Media coverage
-  - Other (please specify)
-
-**Section 4: Impact of Campaigns and Events**
-- **Have recent political events influenced your donation behavior?**
-  - Yes
-  - No
-- **If yes, please specify the events and how they influenced your donations.** (Open-ended)
-- **How do you perceive the effectiveness of political campaigns in encouraging donations?**
-  - Very ineffective
-  - Ineffective
-  - Neutral
-  - Effective
-  - Very effective
-
-**Section 5: Feedback and Suggestions**
-- **What could political parties do to encourage more donations from supporters?** (Open-ended)
-- **Any additional comments or suggestions regarding political donations?** (Open-ended)
+
+**State of Residence:** Dropdown list of all provinces and
+territories. - **Age:** Multiple-choice options (e.g., 18-24, 25-34,
+etc.). - **Gender:** Options including Male, Female, Non-binary/Other,
+Prefer not to say. - **Income Level:** Income brackets (e.g., Less than
+\$25,000, \$25,000-\$49,999, etc.). - **Education Level:** Highest
+degree obtained (e.g., High school, Bachelor’s, Master’s, etc.). -
+**Occupation:** Open-ended or categorized options. - **Political
+Affiliation:** Conservative, Liberal, NDP, Other.
+
+**Section 2: Donation Behavior** - **Are you currently registered to
+donate to political parties?** - Yes - No - **How frequently do you
+donate to political parties?** - Never - Rarely (once a year or less) -
+Occasionally (a few times a year) - Regularly (monthly or more) -
+**Average Donation Amount:** Slider or multiple-choice options (e.g.,
+Less than \$50, \$50-\$199, etc.). - **Preferred Donation Channels:**
+Multiple selections (e.g., Online platforms, Mail-in donations,
+In-person events, etc.).
+
+**Section 3: Motivations for Donating** - **What motivates you to donate
+to a political party?** (Select all that apply) - Support for specific
+policies or issues - Desire to influence election outcomes - Loyalty to
+the party or its leaders - Social pressure or community influence - Tax
+benefits - Other (please specify) - **Which factors most influence your
+decision to donate?** (Rank in order of importance) - Candidate's
+leadership qualities - Party's stance on key issues - Party's past
+performance - Recommendations from peers or influencers - Media
+coverage - Other (please specify)
+
+**Section 4: Impact of Campaigns and Events** - **Have recent political
+events influenced your donation behavior?** - Yes - No - **If yes,
+please specify the events and how they influenced your donations.**
+(Open-ended) - **How do you perceive the effectiveness of political
+campaigns in encouraging donations?** - Very ineffective - Ineffective -
+Neutral - Effective - Very effective
+
+**Section 5: Feedback and Suggestions** - **What could political parties
+do to encourage more donations from supporters?** (Open-ended) - **Any
+additional comments or suggestions regarding political donations?**
+(Open-ended)
 
@@
 
-1.
-**Sampling Method:** Utilize stratified random sampling to ensure representation across different provinces, income levels, education backgrounds, and political affiliations.
-Oversample underrepresented groups to enhance data diversity.
+1.  **Sampling Method:** Utilize stratified random sampling to ensure
+    representation across different provinces, income levels, education
+    backgrounds, and political affiliations. Oversample underrepresented
+    groups to enhance data diversity.
 
 2.  **Data Collection Modes:**
-   - **Online Surveys:** Distribute the survey through email invitations, social media platforms, and online panels to reach a broad audience.
-   - **Telephone Surveys:** Conduct phone interviews to include respondents who may have limited internet access or prefer verbal communication.
-   - **In-Person Interviews:** Organize face-to-face interviews in selected regions to capture detailed qualitative data.
-
-3. **Incentives:** Offer incentives such as entry into a sweepstakes, gift cards, or charitable donations to encourage participation and improve response rates.
+
+    -   **Online Surveys:** Distribute the survey through email
+        invitations, social media platforms, and online panels to reach
+        a broad audience.
+    -   **Telephone Surveys:** Conduct phone interviews to include
+        respondents who may have limited internet access or prefer
+        verbal communication.
+    -   **In-Person Interviews:** Organize face-to-face interviews in
+        selected regions to capture detailed qualitative data.
+
+3.  **Incentives:** Offer incentives such as entry into a sweepstakes,
+    gift cards, or charitable donations to encourage participation and
+    improve response rates.
 
 4.  **Survey Length and Design:**
-   - **Duration:** Design the survey to be completed within 10-15 minutes to minimize respondent fatigue.
-   - **Clarity:** Use clear and concise language, avoiding technical jargon to ensure respondents understand each question.
-   - **Question Types:** Employ a mix of multiple-choice, Likert scales, and open-ended questions to capture both quantitative and qualitative data.
+
+    -   **Duration:** Design the survey to be completed within 10-15
+        minutes to minimize respondent fatigue.
+    -   **Clarity:** Use clear and concise language, avoiding technical
+        jargon to ensure respondents understand each question.
+    -   **Question Types:** Employ a mix of multiple-choice, Likert
+        scales, and open-ended questions to capture both quantitative
+        and qualitative data.
 
 5.  **Data Quality Measures:**
-   - **Pilot Testing:** Conduct a pilot survey to identify and rectify any issues with question wording, survey flow, or technical glitches.
-   - **Anonymity and Confidentiality:** Assure respondents that their answers are confidential and will be used solely for research purposes to encourage honest and accurate responses.
-   - **Response Validation:** Implement checks to prevent duplicate responses, ensure logical consistency, and verify the completeness of the survey.
+
+    -   **Pilot Testing:** Conduct a pilot survey to identify and
+        rectify any issues with question wording, survey flow, or
+        technical glitches.
+    -   **Anonymity and Confidentiality:** Assure respondents that their
+        answers are confidential and will be used solely for research
+        purposes to encourage honest and accurate responses.
+    -   **Response Validation:** Implement checks to prevent duplicate
+        responses, ensure logical consistency, and verify the
+        completeness of the survey.
 
 6.  **Data Analysis:**
-   - **Weighting:** Apply weighting adjustments to account for any sampling biases and ensure that the survey results accurately reflect the population demographics.
-   - **Statistical Techniques:** Use appropriate statistical methods to analyze the data, identifying significant predictors of donation behavior and uncovering underlying trends.
+
+    -   **Weighting:** Apply weighting adjustments to account for any
+        sampling biases and ensure that the survey results accurately
+        reflect the population demographics.
+    -   **Statistical Techniques:** Use appropriate statistical methods
+        to analyze the data, identifying significant predictors of
+        donation behavior and uncovering underlying trends.
 

### 2024-12-04 paper/paper.qmd
Update idealized survey

@@
 
-**State of Residence:** Dropdown list of all provinces and
-territories. - **Age:** Multiple-choice options (e.g., 18-24, 25-34,
-etc.). - **Gender:** Options including Male, Female, Non-binary/Other,
-Prefer not to say. - **Income Level:** Income brackets (e.g., Less than
-\$25,000, \$25,000-\$49,999, etc.). - **Education Level:** Highest
-degree obtained (e.g., High school, Bachelor’s, Master’s, etc.). -
-**Occupation:** Open-ended or categorized options. - **Political
-Affiliation:** Conservative, Liberal, NDP, Other.
-
-**Section 2: Donation Behavior** - **Are you currently registered to
-donate to political parties?** - Yes - No - **How frequently do you
-donate to political parties?** - Never - Rarely (once a year or less) -
-Occasionally (a few times a year) - Regularly (monthly or more) -
-**Average Donation Amount:** Slider or multiple-choice options (e.g.,
-Less than \$50, \$50-\$199, etc.). - **Preferred Donation Channels:**
-Multiple selections (e.g., Online platforms, Mail-in donations,
-In-person events, etc.).
-
-**Section 3: Motivations for Donating** - **What motivates you to donate
-to a political party?** (Select all that apply) - Support for specific
-policies or issues - Desire to influence election outcomes - Loyalty to
-the party or its leaders - Social pressure or community influence - Tax
-benefits - Other (please specify) - **Which factors most influence your
-decision to donate?** (Rank in order of importance) - Candidate's
-leadership qualities - Party's stance on key issues - Party's past
-performance - Recommendations from peers or influencers - Media
-coverage - Other (please specify)
-
-**Section 4: Impact of Campaigns and Events** - **Have recent political
-events influenced your donation behavior?** - Yes - No - **If yes,
-please specify the events and how they influenced your donations.**
-(Open-ended) - **How do you perceive the effectiveness of political
-campaigns in encouraging donations?** - Very ineffective - Ineffective -
-Neutral - Effective - Very effective
-
-**Section 5: Feedback and Suggestions** - **What could political parties
-do to encourage more donations from supporters?** (Open-ended) - **Any
-additional comments or suggestions regarding political donations?**
-(Open-ended)
-
-Survey Administration
-
-1.  **Sampling Method:** Utilize stratified random sampling to ensure
-    representation across different provinces, income levels, education
-    backgrounds, and political affiliations. Oversample underrepresented
-    groups to enhance data diversity.
-
-2.  **Data Collection Modes:**
-
-    -   **Online Surveys:** Distribute the survey through email
-        invitations, social media platforms, and online panels to reach
-        a broad audience.
-    -   **Telephone Surveys:** Conduct phone interviews to include
-        respondents who may have limited internet access or prefer
-        verbal communication.
-    -   **In-Person Interviews:** Organize face-to-face interviews in
-        selected regions to capture detailed qualitative data.
-
-3.  **Incentives:** Offer incentives such as entry into a sweepstakes,
-    gift cards, or charitable donations to encourage participation and
-    improve response rates.
-
-4.  **Survey Length and Design:**
-
-    -   **Duration:** Design the survey to be completed within 10-15
-        minutes to minimize respondent fatigue.
-    -   **Clarity:** Use clear and concise language, avoiding technical
-        jargon to ensure respondents understand each question.
-    -   **Question Types:** Employ a mix of multiple-choice, Likert
-        scales, and open-ended questions to capture both quantitative
-        and qualitative data.
-
-5.  **Data Quality Measures:**
-
-    -   **Pilot Testing:** Conduct a pilot survey to identify and
-        rectify any issues with question wording, survey flow, or
-        technical glitches.
-    -   **Anonymity and Confidentiality:** Assure respondents that their
-        answers are confidential and will be used solely for research
-        purposes to encourage honest and accurate responses.
-    -   **Response Validation:** Implement checks to prevent duplicate
-        responses, ensure logical consistency, and verify the
-        completeness of the survey.
-
-6.  **Data Analysis:**
-
-    -   **Weighting:** Apply weighting adjustments to account for any
-        sampling biases and ensure that the survey results accurately
-        reflect the population demographics.
-    -   **Statistical Techniques:** Use appropriate statistical methods
-        to analyze the data, identifying significant predictors of
-        donation behavior and uncovering underlying trends.
+**State of Residence:** - Dropdown list of all provinces and
+territories.
+
+**Age:** Multiple-choice options
+
+-   18-24
+
+-   25-34
+
+-   35-44
+
+-   45-54
+
+-   55-64
+
+**Gender:**
+
+-   Male
+
+-   Female
+
+-   Non-binary/Other
+
+-   Prefer not to say.
+
+**Income Level:**
+
+-   Less than \$25,000
+
+-   \$25,000-\$49,999
+
+-   \$50,000 - \$74,999
+
+-   \$75,000 - \$99,000
+
+-   \$100,000 - \$149,000
+
+-   \$150,000 or more
+
+**Education Level:** Highest degree obtained
+
+-   Less than high school
+
+-   High school diploma or equivalent
+
+-   Some college, no degree
+
+-   Associate degree
+
+-   Bachelor's degree
+
+-   Graduate or professional degree
+
+**Political Affiliation:**
+
+-   Conservative
+
+-   Liberal
+
+-   NDP
+
+-   Other (please specify)
+
+**Section 2: Donation Behavior**
+
+**Are you currently registered to donate to political parties?**
+
+-   Yes
+
+-   No
+
+**How frequently do you donate to political parties?**
+
+-   Never
+
+-   Rarely (once a year or less)
+
+-   Occasionally (a few times a year)
+
+-   Regularly (monthly or more)
+
+**Average Donation Amount:**
+
+-   Less than \$200
+
+-   \$200 - \$1,000
+
+-   \$1,001 - \$5,000
+
+-   \$5,001 - \$10,000
+
+-   More than \$10,000
+
+**Preferred Donation Channels:**
+
+-   Online platforms
+
+-   Mail-in donations
+
+-   In-person events
+
+**Section 3: Motivations for Donating**
+
+**What motivates you to donate to a political party?** (Select all that
+apply)
+
+-   Support for specific policies or issues
+
+-   Desire to influence election outcomes
+
+-   Loyalty to the party or its leaders
+
+-   Social pressure or community influence
+
+-   Tax benefits
+
+-   Other (please specify)
+
+**Which factors most influence your decision to donate?** (Select all
+that apply)
+
+-   Candidate's leadership qualities
+
+-   Party's stance on key issues
+
+-   Party's past performance
+
+-   Recommendations from peers or influencers
+
+-   Media coverage
+
+-   Other (please specify)
+
+**Section 4: Impact of Campaigns and Events**
+
+**Have recent political events influenced your donation behavior?**
+
+-   Yes
+
+-   No
+
+**If yes, please specify the events and how they influenced your
+donations.** (Open-ended)
+
+**How do you perceive the effectiveness of political campaigns in
+encouraging donations?**
+
+-   Very ineffective
+
+-   Ineffective
+
+-   Neutral
+
+-   Effective
+
+-   Very effective
+
+**Section 5: Feedback and Suggestions**
+
+**What could political parties do to encourage more donations from
+supporters?** (Open-ended) -
+
+**Any additional comments or suggestions regarding political
+donations?** (Open-ended)
 

### 2024-12-04 paper/paper.qmd
update llm usage

@@
 ---
-title: "Impact of Party Leadership on Donation Patterns"
-subtitle: "What Data from 2013 - 2024 Reveal About Party Funding Trends"
+title: "Donor Preferences in Canadian Political Financing"
+subtitle: "Understanding Donor Support for Ruling vs. Opposition Parties Between 2013 - 2024"
 author: 
@@
 library(dplyr)
-
 # Tidy models and combine
@@
   filter(term != "(Intercept)") %>%
-  mutate(term = recode(term, 
-                       "In_Power" = "In Power",
-                       "Party_Size" = "Party Size",
-                       "Election_Year" = "Election Year",
-                       "Party" = "Party (1=Conservative, 0=Liberal)",
-                       "In_Power:Party" = "In Power * Party"))
+  mutate(term = case_when(
+    term == "In_Power" ~ "In Power",
+    term == "Party_Size" ~ "Party Size",
+    term == "Election_Year" ~ "Election Year",
+    term == "Party" ~ "Party (1=Conservative, 0=Liberal)",
+    term == "In_Power:Party" ~ "In Power * Party",
+    TRUE ~ term  # Keeps other terms unchanged
+  ))
 

### 2024-12-04 paper/paper.qmd
cleaned up scripts

@@
 
-At the heart of politics is finance, as revealed by Traag's paper on the
-complex contagion of campaign donations @traag2016complex. Their study
-demonstrates that a majority of campaign contributions in the US
-originate from a few wealthy individuals through large social networks.
-With the constant battle between government and corporations, the
-findings in this study underscore the significant influence of financial
-power in shaping political dynamics.
+At the heart of politics is finance, as revealed by Traag's paper on the complex contagion of campaign donations @traag2016complex.
+Their study demonstrates that a majority of campaign contributions in the US originate from a few wealthy individuals through large social networks.
+With the constant battle between government and corporations, the findings in this study underscore the significant influence of financial power in shaping political dynamics.
 
 The landscape in Canada for political financing is growing each year.
-Results from studies demonstrate that individual contributors may prefer
-to support challengers for several reasons. An optimistic but naive
-explanation may be that individuals hope to foster competitive
-elections, given that donations depend not on political leaning but on
-who is currently in power. However, a more realistic explanation for
-this phenomenon is that some individuals want to make policy changes. It
-is no secret that politicians, particularly those out of power, often
-solicit funding and political support in exchange for advancing the
-interests of their stakeholders @mollona2022ethics. These two dynamics
-may be working in tandem to create an effect where parties eager to be
-elected rely more on funding from external sources, which can switch
-each election cycle.
-
-Moreover, these trends of political donations fluctuating based on the
-party in power are not unique to Canada. The study "Financing UK
-Democracy: A Stocktake of 20 Years of Political Donations" highlights a
-surge in private political donations, more specifically from
-superdonors, widening the resource gap between large parties
-@draca2022financing. Similar to Canada, the UK also sees large donations
-from a wealthy few influencing their elections each year. The parallel
-between Canada and the UK underscores a broader democratic issue of
-financial power being used to influence political outcomes. Though this
-concept is not novel, this study contributes to the already growing
-spotlight on the persistent lack of proper reform in political systems.
-
-This study provides valuable insights into donation behaviours across
-Canada but it its limitations. First, the analysis is applied to only
-federal and provincial donation records, overlooking regional variations
-found in other provinces and territories. The geographical limitations
-of this study hinder its ability to generalize across the entire
-country. Moreover, the study only spans from 2006 to 2024 which economic
-or political events could have influence over the donation patterns that
-the models do not account for.
-
-## Second discussion point
-
-## Third discussion point
+Results from studies demonstrate that individual contributors may prefer to support challengers for several reasons.
+An optimistic but naive explanation may be that individuals hope to foster competitive elections, given that donations depend not on political leaning but on who is currently in power.
+However, a more realistic explanation for this phenomenon is that some individuals want to make policy changes.
+It is no secret that politicians, particularly those out of power, often solicit funding and political support in exchange for advancing the interests of their stakeholders @mollona2022ethics}.
+These two dynamics may be working in tandem to create an effect where parties eager to be elected rely more on funding from external sources, which can switch each election cycle.
+
+Moreover, these trends of political donations fluctuating based on the party in power are not unique to Canada.
+The study "Financing UK Democracy: A Stocktake of 20 Years of Political Donations" highlights a surge in private political donations, more specifically from superdonors, widening the resource gap between large parties @draca2022financing}.
+Similar to Canada, the UK also sees large donations from a wealthy few influencing their elections each year.
+The parallel between Canada and the UK underscores a broader democratic issue of financial power being used to influence political outcomes.
+Though this concept is not novel, this study contributes to the already growing spotlight on the persistent lack of proper reform in political systems.
 
 ## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
+This study provides valuable insights into donation behaviours across Canada but it its limitations.
+First, the analysis is applied to only federal and provincial donation records, overlooking regional variations found in other provinces and territories.
+The geographical limitations of this study hinder its ability to generalize across the entire country.
+Moreover, the study only spans from 2006 to 2024 which economic or political events could have influence over the donation patterns that the models do not account for.
+Additionally qualitative factors such as personal beliefs, specific policy preferences and campaign effectiveness are not directly addressed in this study.
+Data quality issues such as potential missing values or inconsistencies could also have affected the accuracy of the findings here.
+Lastly, the linear regression models here assume a linear relationship between predictors and donations which can greatly oversimplify the complexities.
+
+Future research should not only aim to address these limitations, but expand the analysis to other provinces and territories across Canada.
+This would enhance the generalisability of the results and provide a more nuanced understanding of regional differences.
+Furthermore, utilizing nonlinear modelling techniques or looking at a broader set of predictor variables such as donor demographics or economic indicators could better capture the nuances of donor behaviour and also mitigate the risk of omitted variable bias.
+Comparative studies with other democratic countries, beyond Canada, could provide more general perspectives on how other political systems and regulatory frameworks influence or are influenced by donation behaviours.
+Other valuable areas of future exploration include assessing the impact of media coverage on donation trends.
+Finally, by addressing these areas, research can build on the current study's findings and contribute to a greater understanding of political donations.
 
@@
 
+[Purpose of survey]
+[Link to survey]
+
 **Section 1: Demographics**

### 2024-12-04 scripts/01-test_simulated_data.R

@@ -1,8 +1,7 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated political donation data 
-  #electoral divisions dataset.
+# Purpose: Tests the structure and validity of the simulated political donation data electoral divisions dataset.
 # Author: Maria Mangru
-# Date: 25 November 2024
+# Date: 12 December 2024
 # Contact: maria.mangru@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 

### 2024-12-04 scripts/02-download_data.py

@@ -1,5 +1,15 @@
+#### Preamble ####
+# Purpose: Download political donations data from IJF 
+# Author: Maria Mangru
+# Date: 12 December 2024
+# Contact: maria.mangru@mail.utoronto.ca
+# License: MIT
+
+
+#### Workspace setup ####
 import requests
 import pandas as pd
+
 # API Endpoint
 url = "https://ieypoqqymehdkesavuan.supabase.co/rest/v1/don_mv"
 
@@ -11,6 +21,7 @@ params = {
     "order": "donation_year.desc,amount.desc",
     "donation_year": "eq.2024"
 }
+
 # Headers
 # Replace # with the appropriate details
 headers = {

### 2024-12-04 scripts/03-clean_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans the raw political donations data
 # Author: Maria Mangru
-# Date: 25 November 2024
+# Date: 12 December 2024
 # Contact: maria.mangru@mail.utoronto.ca
 # License: MIT
 

### 2024-12-04 scripts/04-test_analysis_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests the structure and validity of the analysis dataset.
+# Author: Maria Mangru
+# Date: 12 December 2024
+# Contact: maria.mangru@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
 
 
 #### Workspace setup ####

### 2024-12-04 scripts/05-exploratory_data_analysis.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
+# Purpose: Explores the analysis dataset to determine any trends and patterns
 # Author: Maria Mangru
-# Date: 25 November 2024
+# Date: 12 December 2024
 # Contact: maria.mangru@mail.utoronto.ca
 # License: MIT
 

### 2024-12-04 scripts/06-model_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Models
+# Purpose: Generates regression for donation amount and power status
 # Author: Maria Mangru
-# Date: 25 November 2024
+# Date: 12 December 2024
 # Contact: maria.mangru@mail.utoronto.ca
 # License: MIT
 

### 2024-12-04 scripts/07-test_model.R

@@ -1,9 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
+# Purpose: Tests the validity of the models generated
 # Author: Maria Mangru
-# Date: 25 November 2024
+# Date: 12 December 2024
 # Contact: maria.mangru@mail.utoronto.ca
 # License: MIT
+# Pre-requisites: 
+# - The `tidyverse` package must be installed and loaded
 
 
 #### Workspace setup ####

### 2024-12-04 other/datasheet/datasheet.qmd
updated data section

---
title: "Datasheet for 'Donor Preferences in Canadian Political Financing'"
subtitle: "Understanding Donor Support for Ruling vs. Opposition Parties Between 2013 - 2024"
author: 
  - Maria Mangru
thanks: "Code and data are available at: [https://github.com/MariaMangru/Ontario_Political_Donors](https://github.com/MariaMangru/Ontario_Political_Donors)."
date: today
date-format: long
abstract: "This study investigates how the ruling political parties in Canada's federal and Ontario provincial governments affect the donation behaviors of individual contributors. By analyzing donation records from 2013 to 2024, the research explores whether a party being in power influences the total amount of financial support it receives. The findings reveal that opposition parties often receive more donations than those in power, highlighting donor preferences to support challengers. This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of Australian politicians. We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The Investigative Journalism Foundation created the dataset, they are an independent organization.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The IJF is funded by several foundations: Mitacs, Social Sciences and Humanities Research Council, 3858278 Canada Foundation, Canadian Internet Registration Authority, Data Driven Reporting Project, Avanti Foundation, Balsillie Family Foundation, C4C Canada, Data Science For Social Good, Toronto Metropolitan University and Individuals.
4. *Any other comments?*
    - NA

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances that comprise the dataset represent political donations made by individuals or entities to political parties or entities in Canada.
2. *How many instances are there in total (of each type, if appropriate)?*
	- There are 9,269,201 instances in total.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset does not contain all possible instances since it only covers from 1993 to the present for federal data, but British Columbia's data only spans from 2005 onwards. The larger set would be all public data that references donations both federally and provincially. The sample is not representative of the larger set as it only covers a very short period of time since record availability and disclosure practices have withheld lots of public data.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- The data comprising each instance is an processed json file with features, these features include region, date added, electoral event, political entity, political party, recipient, donor location, donation date, donation year, donor full name, donor type, amount and the monetary/non monetary amount as part of the total amount.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- There is no label or target associated with each instance.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Some records may have missing information due to data conversion (using OCR technology to convert pdf records into text), older recordes or the laws in different jurisdictions allowing corporate donations or not.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Relationships are not explicitly defined but can be inferred.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- The data does not come with recommended data splits.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Yes, inconsistent naming conventions, data entry mistakes and OCR errors (which is the process of converting pdf to text) can occur.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset does rely on public federal and provincial datasets to exist, but doesn't explicitly link to them. There are guarantees it will exist since it is compiled from publicly accessible records. There are archival version of the dataset that the IJF maintains. Also, the government of Canada and other provinces have archival versions as well. There are restrictions, which are fees to be able to access this data from teh website.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, the dataset is comprised of public records and information of political donations which are legally required to be disclosed and accessible to the public.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No it does not contains offensive, insulting, threatening or otherwise anxiety-inducing information.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The dataset does explicitly identify sub-populations for example specific regions and political parties and corporations which could be considered subpopulations. [X]
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- Yes it is possible to identify individuals directly from the dataset by reading their full name from the row. All donations that exceed a certain threshold must be identifiable.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- Yes political opinions can be inferred from the donation records as it shows who donated to which party.
16. *Any other comments?*
	- NA

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data associated with each instance was acquired from official records that are maintained by election agencies at municipal, territorial, provincial and federal levels. All of these records are publicly accessible and include the detailed information on political donations.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The mechanisms used to collect the data were automated retrieval, manual retrieval with OCR (Optical Character Recognition). These mechanisms were validated using manual verifications to correct any errors.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The sample strategy was to fetch as much information as legally possible.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The data collection team was done by the IJF team which are staff membesr and potentially contractors. Compensation for this was not publicly disclosed.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The timeframe was 1993 to 2024, the data is updated constantly as soon as its released publicly, so it does match the creation timeframe.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Any ethical review processes are not explicitly mentioned since this is scraped public data from different levels of government in Canada.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was collected by making API calls to their server, therefore it was obtained by third parties.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- The individuals were not notified of the data collection.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- The individuals did consent to the collection and use of their data, since it is publicly disclosed data mandated by the law. Moreover, for the IJF's use of this publicly available data, additional consent is not required for this dataset.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- NA
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No analysis has been done.
12. *Any other comments?*
	- NA

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Yes, filtered the data for Federal and Ontario, then ensured the columns are the right datatype, records with incomplete information with key variables were excluded. Names were sometimes combined into one a single name for consistency, binary variables were also introduced and others for normalization and logarithmic transformations.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, the raw data can be accessed at: https://github.com/MariaMangru/Ontario_Political_Donors/blob/main/data/01-raw_data/raw_data.csv
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The software was preprocessed and cleaned with RStudio https://posit.co/products/open-source/rstudio/
4. *Any other comments?*
	- NA

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset has been used for the current study, which investigated the impact of party leadership on donation patterns.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- This repository https://github.com/MariaMangru/Ontario_Political_Donors/tree/main is the only one that links to a paper that uses the dataset.
3. *What (other) tasks could the dataset be used for?*
	- The dataset could be used for policy analysis and data journalism such as creating stories to highlight donation patterns or analyzing political financing trends.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- A dataset consumer could conceal the names of the donors which donated large sums of money to political parties as it shows their full name or corporation name on the dataset.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used to harass or target individual donors.
6. *Any other comments?*
	- NA

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Yes the dataset is distributed to third parties through its website and public asccess to political donation records.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is mainly accessible by their website where users can search and view records. There is currently no digital object identifier on the dataset.
3. *When will the dataset be distributed?*
	- The dataset is already distributed as it is publicly available and updated with the most recent data.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- The full dataset is behind a paywall, but other than that since it is all publicly available data it will stay publicly availabe with no intellectual property license on it.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No third parties have imposed IP-based restrictions since it is public available government data that they use.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- Yes there are, you cannot download or export the dataset like you would on other sites, you can only view it through the table on the search page.
7. *Any other comments?*
	- NA

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- THe IJF maintains, supports and hosts the political donations dataset.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- They can be contacted through their contact page on their website which is a form for communication. Email addresses are not publicly listed.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- There is no mention of an erratum.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- The dataset is constantly updated to reflect new political donations, the IJF does not describe how often they do this.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- The dataset includes information on political donations which contains donor names and details for large enough sums. This information is publicly accessible from the government and maintined by election agencies which have their own respective data retention policies.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- The IJF does not mention whether older version of the dataset will continue to be supported.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- The IJF does not allow any mechanisms for external contributions to the dataset.
8. *Any other comments?*
	- NA
\newpage

# References

### 2024-12-04 paper/paper.qmd
updated data section

@@
 
-The dataset used in this analysis comprises political donation records
-from 2013 to 2024, covering both the federal level and the province of
-Ontario in Canada. These records were sourced from the
-@ijf_political_donors Political Donors Dataset, which systematically
-collects and maintains comprehensive data on political donations across
-Canadian jurisdictions. The IJF updates this dataset daily by monitoring
-election agency websites at the federal level and within each province
-and territory, ensuring the data remains current and reliable. The
-analysis was conducted using both R (@citeR) and Python (@python),
+The dataset analyzed in this study encompasses political donation records from 2006 to 2024, covering both the federal level and the province of Ontario in Canada.
+These records were sourced from the @ijf_political_donors Political Donors Dataset, which systematically collects and maintains comprehensive data on political donations across Canadian jurisdictions.
+The IJF updates this dataset daily by monitoring election agency websites at the federal level and within each province and territory, ensuring the data remains current and reliable.
+covering both the federal level and the province of
+Ontario in Canada. The analysis was conducted using both R (@citeR) and Python (@python),
 leveraging a range of libraries for data manipulation and visualization.
@@
 report donations exceeding specific thresholds. These disclosures vary
-by jurisdiction in terms of frequency and reporting formats. For
-instance, Ontario requires annual returns for general and by-elections,
-leadership races, and donations to parties, candidates, and constituency
-associations, all available in downloadable spreadsheet formats from
-2007 onward. This regulatory framework ensures that significant
+by jurisdiction in terms of frequency and reporting formats. This regulatory framework ensures that significant
 contributions are systematically documented and publicly accessible,
 providing a reliable foundation for this analysis.
+
+The period starting from 2006 is particularly significant because it marks the first federal election involving the newly formed Conservative Party of Canada.
+In 2003, the Canadian Alliance and the Progressive Conservative Party of Canada merged to create the modern Conservative Party.
+The 2006 election was the first in which this new entity participated, making it a pivotal point for analyzing donation patterns related to this party.
 
@@
 library(dplyr)
-library(gridExtra)
 library(scales)
@@
 
-# Convert donation_year to a properly ordered factor
-donations_by_year_ontario$donation_year <- factor(
-  donations_by_year_ontario$donation_year, 
-  levels = sort(unique(as.numeric(donations_by_year_ontario$donation_year)))
-)
-
-donations_by_year_federal$donation_year <- factor(
-  donations_by_year_federal$donation_year, 
-  levels = sort(unique(as.numeric(donations_by_year_federal$donation_year)))
-)
+# Ensure years are in correct order
+donations_by_year_ontario$donation_year <- as.integer(donations_by_year_ontario$donation_year)
+donations_by_year_federal$donation_year <- as.integer(donations_by_year_federal$donation_year)
 
@@
 
-# Plot using facets with a shared Y-axis
-p_combined <- ggplot(donations_combined, aes(x = donation_year, y = avg_amount, group = Region, color = Region)) +
+# Plot using facets
+ggplot(donations_combined, aes(x = donation_year, y = avg_amount, group = Region, color = Region)) +
   geom_line() +
@@
   labs(
-    title = "Average Donation Amount Over Years",
+    title = "Figure 1: Average Donation Amount Over Years",
     x = "Year",
@@
   ) +
+  scale_x_continuous(breaks = seq(2006, 2024, 2)) +
   scale_y_continuous(labels = dollar_format()) +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-print(p_combined)
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    legend.position = "none"
+  )
 
@@
 [Purpose of survey]
-[Link to survey]
+You can access the survey here: https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform?usp=sharing
 

### 2024-12-04 paper/paper.qmd
update outcome variables section

@@
 are higher, reflecting broader donor bases and different legal
-contribution limits.
-
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+contribution limits. The distribution of the average donation amount can be see in figure @fig-avg-donation.
+
+\newpage
+
+```{r fig-avg-donation, echo=FALSE, message=FALSE, warning=FALSE}
+#| label: fig-avg-donation
+#| tbl-cap: Figure Showing the Average Donation Amounts Over Years
+
 library(ggplot2)
@@
   labs(
-    title = "Figure 1: Average Donation Amount Over Years",
     x = "Year",

### 2024-12-04 paper/paper.qmd
finalize data section

@@
 covering both the federal level and the province of
-Ontario in Canada. The analysis was conducted using both R (@citeR) and Python (@python),
+Ontario in Canada.
+
+The analysis was conducted using both R (@citeR) and Python (@python),
 leveraging a range of libraries for data manipulation and visualization.
@@
 are higher, reflecting broader donor bases and different legal
-contribution limits. The distribution of the average donation amount can be see in figure @fig-avg-donation.
+contribution limits.
 
@@
 
+@fig-avg-donation shows the trend of average donation amounts over the years for both Ontario and Federal levels.
+The x-axis represents the years from 2006 to 2024, while the y-axis shows the average donation amount in Canadian dollars.
+The data reveals fluctuations in donation amounts over time, with notable during 2018 for Ontario and 2024 for Federal.
+
 ## Predictor Variables
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+```{r tbl-fed-donation, echo=FALSE, message=FALSE, warning=FALSE}
+#| label: tbl-fed-donation
+#| tbl-cap: Table Showing Donations Based on Recipient Power Status at Federal Level
 
@@
 library(kableExtra)
-library(scales) # For dollar formatting
+library(scales)
 
@@
 kable(donations_in_power_federal, 
-      caption = "Donations Based on Recipient Power Status - Federal", 
       col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+In @tbl-fed-donation, parties in power (indicated by 1) received both lower total donation amounts and lower average donation amounts compared to opposition parties.
+This pattern suggests that donors may prefer to support challengers over incumbents.
+
+```{r tbl-ontario-donation, echo=FALSE, message=FALSE, warning=FALSE}
+#| label: tbl-ontario-donation
+#| tbl-cap: Table Showing Donations Based on Recipient Power Status at Ontario Level
+
 ## Donations Based on Recipient Power Status - Ontario
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
-# Load required libraries
+@tbl-ontario-donation also shows a similar trend of parties in power receiving lower total and average donations.
+
+```{r fig-ontario-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
+#| label: fig-ontario-total-donation
+#| tbl-cap: Figure Showing Total Donations by Political Party at Ontario Level
+
 library(dplyr)
@@
 library(treemapify)
-
-# Summarize data for tree map
+library(scales)
+
+# Summarize data for treemap
 donations_by_party_ontario <- ontario_data %>%
@@
 
-# Create the tree map
+# Create the treemap
 ggplot(donations_by_party_ontario, aes(
@@
     place = "center", 
-    grow = TRUE
-  ) +
-  labs(
-    title = "Total Donations by Political Party - Ontario",
-    subtitle = "Size of rectangles represents total donation amounts",
-    caption = "Data source: Ontario Political Donations"
+    grow = FALSE,
+    reflow = TRUE,
+    size = 8
   ) +
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+@fig-ontario-total-donation displays a tree map of total donations received by each political party in Ontario.
+The size of each rectangle corresponds to the total amount donated to that party.
+Major parties like the Progressive Conservative Party of Ontario and the Liberal Party of Ontario dominate the donation landscape, reflecting their significant fundraising capabilities.
+
+```{r fig-federal-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
+#| label: fig-federal-total-donation
+#| tbl-cap: Figure Showing Total Donations by Political Party at Federal Level
+
 # Load required libraries
@@
 
-# Summarize data for tree map
+# Summarize data for treemap
 donations_by_party_federal <- federal_data %>%
@@
 
-# Create the tree map
+# Create the treemap
 ggplot(donations_by_party_federal, aes(
@@
     place = "center", 
-    grow = TRUE
-  ) +
-  labs(
-    title = "Total Donations by Political Party - Federal",
-    subtitle = "Size of rectangles represents total donation amounts",
-    caption = "Data source: Federal Political Donations"
+    grow = FALSE,
+    reflow = TRUE,
+    size = 8
   ) +
@@
 ```
+
+@fig-federal-total-donation presents a treemap for federal political parties.
+Similar to the provincial level, major parties such as the Conservative Party of Canada and the Liberal Party of Canada receive the largest amounts in donations.
+This visualization highlights the prominence of major parties in attracting financial support at the federal level.
 

### 2024-12-04 paper/paper.qmd
update predictor variable

@@
 pandas (@pandas), requests (@requests), and matplotlib (@matplotlib)
-were utilized for additional data processing and visualization tasks
+were utilized for gathering the data.
 
@@
 
+\newpage
+
 ## Predictor Variables
@@
 
+\newpage
 ```{r tbl-fed-donation, echo=FALSE, message=FALSE, warning=FALSE}
@@
 kable(donations_in_power_ontario, 
-      caption = "Donations Based on Recipient Power Status - Ontario", 
       col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
@@
 
+\newpage
 # Model {#sec-model}

### 2024-12-04 scripts/06-model_data.R
save models

@@ -9,6 +9,7 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(here)
 
 #### Read data ####
 analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
@@ -105,4 +106,16 @@ total_donations_cl <- total_donations_cl %>%
 model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)
 summary(model_cl)
 
-# Need to save the model
+#### Save model ####
+
+if (!dir.exists("models")) {
+  dir.create("models")
+}
+# Save the Ontario model
+saveRDS(model_ontario, file = "model/model_ontario.rds")
+
+# Save the Federal model
+saveRDS(model_federal, file = "model/model_federal.rds")
+
+# Save the Conservative and Liberal model
+saveRDS(model_cl, file = "model/model_cl.rds")

### 2024-12-04 paper/paper.qmd
updated summary tables

@@
 
+\newpage
 # Data {#sec-data}
@@
 
-\newpage
-
 ```{r fig-avg-donation, echo=FALSE, message=FALSE, warning=FALSE}
@@
 \newpage
+
 ```{r tbl-fed-donation, echo=FALSE, message=FALSE, warning=FALSE}
@@
       align = c("l", "c", "c", "c")) %>%
-  kable_styling(latex_options = c("striped", "hold_position"))
+  kable_styling(latex_options = c("striped"))
 
@@
       align = c("l", "c", "c", "c")) %>%
-  kable_styling(latex_options = c("striped", "hold_position"))
+  kable_styling(latex_options = c("striped"))
 ```
@@
 #| label: fig-ontario-total-donation
-#| tbl-cap: Figure Showing Total Donations by Political Party at Ontario Level
+#| fig-cap: Figure Showing Total Donations by Political Party at Ontario Level
 
@@
 #| label: fig-federal-total-donation
-#| tbl-cap: Figure Showing Total Donations by Political Party at Federal Level
+#| fig-cap: Figure Showing Total Donations by Political Party at Federal Level
 
@@
 Additionally, a focused regression model was created for the
-Conservative and Liberal parties to explore the interaction between
-power status and party size.
+Conservative and Liberal parties to explore the interaction between power status and party affiliation.
 
@@
 
-To analyze the impact of a political party's power status on donation
-patterns, we employed linear regression models for both Ontario and
-federal levels:
+The models are specified as follows for the Ontario and Federal Models:
 
@@
 the party (Party_Size), and whether the donation was made during an
-election year (Election_Year). The specialized model for Conservative
+election year (Election_Year).
+
+The specialized model for Conservative
 and Liberal parties includes an additional binary variable (Party) and
@@
 
-The models assume a linear relationship between the predictors and the
-logarithm of total donations, ensuring that the effects of the
-predictors are additive and proportional. Additionally, they assume that
-the residuals are normally distributed (normality of errors) and exhibit
-constant variance (homoscedasticity). The models also rely on the
-absence of multicollinearity, meaning that the predictors are not highly
-correlated, and they assume independence of observations, where each
-data point is independent of the others. Potential limitations of these
-models include the possibility of unobserved variables influencing
-donation amounts and the inherent assumption that past trends can
-reliably predict future behavior.
-
-### Model Validation \*\*\*\*\*\*
-
-Model validation was conducted to assess the predictive accuracy and
-generalizability of the regression models. This involved evaluating key
-metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square
-Error (RMSE) to determine how well the models explain the variance in
-Log_Total_Donations and their predictive performance.
-
-For the Ontario Model, the regression analysis yielded an R² of 0.645
-and an Adjusted R² of 0.635, indicating that approximately 64.5% of the
-variance in log-transformed total donations is explained by the
-predictors. The RMSE for this model was calculated to assess the average
-prediction error, demonstrating a reasonable level of accuracy.
-
-The Federal Model showed an R² of 0.515 and an Adjusted R² of 0.503,
-suggesting that 51.5% of the variance in log-transformed total donations
-is accounted for by the predictors. The RMSE value indicated acceptable
-predictive performance, though slightly lower than the Ontario model,
-reflecting potential differences in donation dynamics at the federal
-level.
-
-In the Conservative and Liberal Parties Model, the inclusion of the
-interaction term between In_Power and Party resulted in an R² of 0.630
-and an Adjusted R² of 0.600. This model effectively captures the
-combined effect of power status and party type on donation amounts, with
-the interaction term providing significant insights into how being in
-power influences donations differently for Conservative and Liberal
-parties. The RMSE for this model confirmed its robust predictive
-capability, comparable to the other models.
-
-Overall, the validation process confirmed that the linear regression
-models are robust and effective in predicting donation amounts based on
-the selected predictor variables. The models demonstrate good
-explanatory power and reasonable predictive accuracy, making them
-suitable for analyzing the factors influencing political donations in
-Canada. However, ongoing evaluation and potential model refinement are
-recommended to accommodate evolving political and financial landscapes,
-ensuring the models remain relevant and accurate over time.
+The models assume a linear relationship between the predictors and the logarithm of total donations, ensuring that the effects of the predictors are additive and proportional.
+Additionally, they assume that the residuals are normally distributed and exhibit constant variance (homoscedasticity).
+The models also rely on the absence of multicollinearity, meaning that the predictors are not highly correlated, and they assume independence of observations.
+
+Potential limitations include the possibility of unobserved variables influencing donation amounts and the inherent assumption that past trends can reliably predict future behavior.
+These factors could introduce bias or limit the generalizability of the findings.
+
+### Model Validation
+
+Model validation was conducted to assess the predictive accuracy and generalizability of the regression models.
+This involved evaluating key metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square Error (RMSE) to determine how well the models explain the variance in Log_Total_Donations and their predictive performance.
+
+The Ontario model yielded an R² of 0.5496 and an Adjusted R² of 0.5267, indicating that approximately 54.96% of the variance in the log-transformed total donations is explained by the predictors.
+The RMSE was 2.5919, reflecting the average prediction error.
+The Variance Inflation Factor (VIF) values for all predictors were below 2, suggesting no significant multicollinearity.
+
+The Federal model showed an R² of 0.6487 and an Adjusted R² of 0.6410, indicating that about 64.87% of the variance is explained by the model.
+The RMSE was 1.4909, and VIF values were again below 2, suggesting no significant multicollinearity.
+
+The Conservative/Liberal model achieved an R² of 0.1091 and an Adjusted R² of 0.0431, indicating a lower explanatory power compared to the previous models.
+The RMSE was 1.2814.
+VIF values suggested moderate multicollinearity due to the interaction term, but they were within acceptable limits.
+
+Overall, the validation process confirmed that the linear regression models are effective in predicting donation amounts based on the selected predictor variables.
+The models demonstrate good explanatory power and reasonable predictive accuracy, making them suitable for analyzing the factors influencing political donations in Canada.
 
@@
 
-Table @tbl-modelsummary presents the summary of key model estimates for
+@tbl-modelsummary presents the summary of key model estimates for
 the Ontario, Federal, and Conservative/Liberal models. These models

### 2024-12-04 paper/paper.qmd
update appendix

@@
 \newpage
+
 # Data {#sec-data}
@@
 
-The dataset analyzed in this study encompasses political donation records from 2006 to 2024, covering both the federal level and the province of Ontario in Canada.
-These records were sourced from the @ijf_political_donors Political Donors Dataset, which systematically collects and maintains comprehensive data on political donations across Canadian jurisdictions.
-The IJF updates this dataset daily by monitoring election agency websites at the federal level and within each province and territory, ensuring the data remains current and reliable.
-covering both the federal level and the province of
-Ontario in Canada.
+The dataset analyzed in this study encompasses political donation
+records from 2006 to 2024, covering both the federal level and the
+province of Ontario in Canada. These records were sourced from the
+@ijf_political_donors Political Donors Dataset, which systematically
+collects and maintains comprehensive data on political donations across
+Canadian jurisdictions. The IJF updates this dataset daily by monitoring
+election agency websites at the federal level and within each province
+and territory, ensuring the data remains current and reliable. covering
+both the federal level and the province of Ontario in Canada.
 
@@
 report donations exceeding specific thresholds. These disclosures vary
-by jurisdiction in terms of frequency and reporting formats. This regulatory framework ensures that significant
-contributions are systematically documented and publicly accessible,
-providing a reliable foundation for this analysis.
-
-The period starting from 2006 is particularly significant because it marks the first federal election involving the newly formed Conservative Party of Canada.
-In 2003, the Canadian Alliance and the Progressive Conservative Party of Canada merged to create the modern Conservative Party.
-The 2006 election was the first in which this new entity participated, making it a pivotal point for analyzing donation patterns related to this party.
+by jurisdiction in terms of frequency and reporting formats. This
+regulatory framework ensures that significant contributions are
+systematically documented and publicly accessible, providing a reliable
+foundation for this analysis.
+
+The period starting from 2006 is particularly significant because it
+marks the first federal election involving the newly formed Conservative
+Party of Canada. In 2003, the Canadian Alliance and the Progressive
+Conservative Party of Canada merged to create the modern Conservative
+Party. The 2006 election was the first in which this new entity
+participated, making it a pivotal point for analyzing donation patterns
+related to this party.
 
@@
 
-@fig-avg-donation shows the trend of average donation amounts over the years for both Ontario and Federal levels.
-The x-axis represents the years from 2006 to 2024, while the y-axis shows the average donation amount in Canadian dollars.
-The data reveals fluctuations in donation amounts over time, with notable during 2018 for Ontario and 2024 for Federal.
+@fig-avg-donation shows the trend of average donation amounts over the
+years for both Ontario and Federal levels. The x-axis represents the
+years from 2006 to 2024, while the y-axis shows the average donation
+amount in Canadian dollars. The data reveals fluctuations in donation
+amounts over time, with notable during 2018 for Ontario and 2024 for
+Federal.
 
@@
 
-In @tbl-fed-donation, parties in power (indicated by 1) received both lower total donation amounts and lower average donation amounts compared to opposition parties.
-This pattern suggests that donors may prefer to support challengers over incumbents.
+In @tbl-fed-donation, parties in power (indicated by 1) received both
+lower total donation amounts and lower average donation amounts compared
+to opposition parties. This pattern suggests that donors may prefer to
+support challengers over incumbents.
 
@@
 
-@tbl-ontario-donation also shows a similar trend of parties in power receiving lower total and average donations.
+@tbl-ontario-donation also shows a similar trend of parties in power
+receiving lower total and average donations.
 
@@
 
-@fig-ontario-total-donation displays a tree map of total donations received by each political party in Ontario.
-The size of each rectangle corresponds to the total amount donated to that party.
-Major parties like the Progressive Conservative Party of Ontario and the Liberal Party of Ontario dominate the donation landscape, reflecting their significant fundraising capabilities.
+@fig-ontario-total-donation displays a tree map of total donations
+received by each political party in Ontario. The size of each rectangle
+corresponds to the total amount donated to that party. Major parties
+like the Progressive Conservative Party of Ontario and the Liberal Party
+of Ontario dominate the donation landscape, reflecting their significant
+fundraising capabilities.
 
@@
 
-@fig-federal-total-donation presents a treemap for federal political parties.
-Similar to the provincial level, major parties such as the Conservative Party of Canada and the Liberal Party of Canada receive the largest amounts in donations.
-This visualization highlights the prominence of major parties in attracting financial support at the federal level.
+@fig-federal-total-donation presents a treemap for federal political
+parties. Similar to the provincial level, major parties such as the
+Conservative Party of Canada and the Liberal Party of Canada receive the
+largest amounts in donations. This visualization highlights the
+prominence of major parties in attracting financial support at the
+federal level.
 
 \newpage
+
 # Model {#sec-model}
@@
 Additionally, a focused regression model was created for the
-Conservative and Liberal parties to explore the interaction between power status and party affiliation.
+Conservative and Liberal parties to explore the interaction between
+power status and party affiliation.
 
@@
 
-The specialized model for Conservative
-and Liberal parties includes an additional binary variable (Party) and
-an interaction term (In_Power \* Party) to capture the combined effect
-of power status and party type on donation amounts.
+The specialized model for Conservative and Liberal parties includes an
+additional binary variable (Party) and an interaction term (In_Power \*
+Party) to capture the combined effect of power status and party type on
+donation amounts.
 
@@
 
-The models assume a linear relationship between the predictors and the logarithm of total donations, ensuring that the effects of the predictors are additive and proportional.
-Additionally, they assume that the residuals are normally distributed and exhibit constant variance (homoscedasticity).
-The models also rely on the absence of multicollinearity, meaning that the predictors are not highly correlated, and they assume independence of observations.
-
-Potential limitations include the possibility of unobserved variables influencing donation amounts and the inherent assumption that past trends can reliably predict future behavior.
-These factors could introduce bias or limit the generalizability of the findings.
+The models assume a linear relationship between the predictors and the
+logarithm of total donations, ensuring that the effects of the
+predictors are additive and proportional. Additionally, they assume that
+the residuals are normally distributed and exhibit constant variance
+(homoscedasticity). The models also rely on the absence of
+multicollinearity, meaning that the predictors are not highly
+correlated, and they assume independence of observations.
+
+Potential limitations include the possibility of unobserved variables
+influencing donation amounts and the inherent assumption that past
+trends can reliably predict future behavior. These factors could
+introduce bias or limit the generalizability of the findings.
 
@@
 
-Model validation was conducted to assess the predictive accuracy and generalizability of the regression models.
-This involved evaluating key metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square Error (RMSE) to determine how well the models explain the variance in Log_Total_Donations and their predictive performance.
-
-The Ontario model yielded an R² of 0.5496 and an Adjusted R² of 0.5267, indicating that approximately 54.96% of the variance in the log-transformed total donations is explained by the predictors.
-The RMSE was 2.5919, reflecting the average prediction error.
-The Variance Inflation Factor (VIF) values for all predictors were below 2, suggesting no significant multicollinearity.
-
-The Federal model showed an R² of 0.6487 and an Adjusted R² of 0.6410, indicating that about 64.87% of the variance is explained by the model.
-The RMSE was 1.4909, and VIF values were again below 2, suggesting no significant multicollinearity.
-
-The Conservative/Liberal model achieved an R² of 0.1091 and an Adjusted R² of 0.0431, indicating a lower explanatory power compared to the previous models.
-The RMSE was 1.2814.
-VIF values suggested moderate multicollinearity due to the interaction term, but they were within acceptable limits.
-
-Overall, the validation process confirmed that the linear regression models are effective in predicting donation amounts based on the selected predictor variables.
-The models demonstrate good explanatory power and reasonable predictive accuracy, making them suitable for analyzing the factors influencing political donations in Canada.
+Model validation was conducted to assess the predictive accuracy and
+generalizability of the regression models. This involved evaluating key
+metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square
+Error (RMSE) to determine how well the models explain the variance in
+Log_Total_Donations and their predictive performance.
+
+The Ontario model yielded an R² of 0.5496 and an Adjusted R² of 0.5267,
+indicating that approximately 54.96% of the variance in the
+log-transformed total donations is explained by the predictors. The RMSE
+was 2.5919, reflecting the average prediction error. The Variance
+Inflation Factor (VIF) values for all predictors were below 2,
+suggesting no significant multicollinearity.
+
+The Federal model showed an R² of 0.6487 and an Adjusted R² of 0.6410,
+indicating that about 64.87% of the variance is explained by the model.
+The RMSE was 1.4909, and VIF values were again below 2, suggesting no
+significant multicollinearity.
+
+The Conservative/Liberal model achieved an R² of 0.1091 and an Adjusted
+R² of 0.0431, indicating a lower explanatory power compared to the
+previous models. The RMSE was 1.2814. VIF values suggested moderate
+multicollinearity due to the interaction term, but they were within
+acceptable limits.
+
+Overall, the validation process confirmed that the linear regression
+models are effective in predicting donation amounts based on the
+selected predictor variables. The models demonstrate good explanatory
+power and reasonable predictive accuracy, making them suitable for
+analyzing the factors influencing political donations in Canada.
 
@@
 
-@tbl-modelsummary presents the summary of key model estimates for
-the Ontario, Federal, and Conservative/Liberal models. These models
-predict the logarithm of total donations received by political parties
-based on whether the party was in power, the size of the party, and
-whether the donation occurred during an election year. The
-Conservative/Liberal model includes an interaction term to explore how
-the effect of being in power varies between these two major parties.
+@tbl-modelsummary presents the summary of key model estimates for the
+Ontario, Federal, and Conservative/Liberal models. These models predict
+the logarithm of total donations received by political parties based on
+whether the party was in power (In_Power), the size of the party
+(Party_Size), and whether the donation occurred during an election year
+(Election_Year). The Conservative/Liberal model includes an interaction
+term (In_Power × Party) to explore how the effect of being in power
+varies between these two major parties.
 
@@
 
-At the heart of politics is finance, as revealed by Traag's paper on the complex contagion of campaign donations @traag2016complex.
-Their study demonstrates that a majority of campaign contributions in the US originate from a few wealthy individuals through large social networks.
-With the constant battle between government and corporations, the findings in this study underscore the significant influence of financial power in shaping political dynamics.
+At the heart of politics is finance, as revealed by Traag's paper on the
+complex contagion of campaign donations @traag2016complex. Their study
+demonstrates that a majority of campaign contributions in the US
+originate from a few wealthy individuals through large social networks.
+With the constant battle between government and corporations, the
+findings in this study underscore the significant influence of financial
+power in shaping political dynamics.
 
 The landscape in Canada for political financing is growing each year.
-Results from studies demonstrate that individual contributors may prefer to support challengers for several reasons.
-An optimistic but naive explanation may be that individuals hope to foster competitive elections, given that donations depend not on political leaning but on who is currently in power.
-However, a more realistic explanation for this phenomenon is that some individuals want to make policy changes.
-It is no secret that politicians, particularly those out of power, often solicit funding and political support in exchange for advancing the interests of their stakeholders @mollona2022ethics}.
-These two dynamics may be working in tandem to create an effect where parties eager to be elected rely more on funding from external sources, which can switch each election cycle.
-
-Moreover, these trends of political donations fluctuating based on the party in power are not unique to Canada.
-The study "Financing UK Democracy: A Stocktake of 20 Years of Political Donations" highlights a surge in private political donations, more specifically from superdonors, widening the resource gap between large parties @draca2022financing}.
-Similar to Canada, the UK also sees large donations from a wealthy few influencing their elections each year.
-The parallel between Canada and the UK underscores a broader democratic issue of financial power being used to influence political outcomes.
-Though this concept is not novel, this study contributes to the already growing spotlight on the persistent lack of proper reform in political systems.
+Results from studies demonstrate that individual contributors may prefer
+to support challengers for several reasons. An optimistic but naive
+explanation may be that individuals hope to foster competitive
+elections, given that donations depend not on political leaning but on
+who is currently in power. However, a more realistic explanation for
+this phenomenon is that some individuals want to make policy changes. It
+is no secret that politicians, particularly those out of power, often
+solicit funding and political support in exchange for advancing the
+interests of their stakeholders @mollona2022ethics}. These two dynamics
+may be working in tandem to create an effect where parties eager to be
+elected rely more on funding from external sources, which can switch
+each election cycle.
+
+Moreover, these trends of political donations fluctuating based on the
+party in power are not unique to Canada. The study "Financing UK
+Democracy: A Stocktake of 20 Years of Political Donations" highlights a
+surge in private political donations, more specifically from
+superdonors, widening the resource gap between large parties
+@draca2022financing}. Similar to Canada, the UK also sees large
+donations from a wealthy few influencing their elections each year. The
+parallel between Canada and the UK underscores a broader democratic
+issue of financial power being used to influence political outcomes.
+Though this concept is not novel, this study contributes to the already
+growing spotlight on the persistent lack of proper reform in political
+systems.
 
 ## Weaknesses and next steps
-This study provides valuable insights into donation behaviours across Canada but it its limitations.
-First, the analysis is applied to only federal and provincial donation records, overlooking regional variations found in other provinces and territories.
-The geographical limitations of this study hinder its ability to generalize across the entire country.
-Moreover, the study only spans from 2006 to 2024 which economic or political events could have influence over the donation patterns that the models do not account for.
-Additionally qualitative factors such as personal beliefs, specific policy preferences and campaign effectiveness are not directly addressed in this study.
-Data quality issues such as potential missing values or inconsistencies could also have affected the accuracy of the findings here.
-Lastly, the linear regression models here assume a linear relationship between predictors and donations which can greatly oversimplify the complexities.
-
-Future research should not only aim to address these limitations, but expand the analysis to other provinces and territories across Canada.
-This would enhance the generalisability of the results and provide a more nuanced understanding of regional differences.
-Furthermore, utilizing nonlinear modelling techniques or looking at a broader set of predictor variables such as donor demographics or economic indicators could better capture the nuances of donor behaviour and also mitigate the risk of omitted variable bias.
-Comparative studies with other democratic countries, beyond Canada, could provide more general perspectives on how other political systems and regulatory frameworks influence or are influenced by donation behaviours.
-Other valuable areas of future exploration include assessing the impact of media coverage on donation trends.
-Finally, by addressing these areas, research can build on the current study's findings and contribute to a greater understanding of political donations.
+
+This study provides valuable insights into donation behaviours across
+Canada but it its limitations. First, the analysis is applied to only
+federal and provincial donation records, overlooking regional variations
+found in other provinces and territories. The geographical limitations
+of this study hinder its ability to generalize across the entire
+country. Moreover, the study only spans from 2006 to 2024 which economic
+or political events could have influence over the donation patterns that
+the models do not account for. Additionally qualitative factors such as
+personal beliefs, specific policy preferences and campaign effectiveness
+are not directly addressed in this study. Data quality issues such as
+potential missing values or inconsistencies could also have affected the
+accuracy of the findings here. Lastly, the linear regression models here
+assume a linear relationship between predictors and donations which can
+greatly oversimplify the complexities.
+
+Future research should not only aim to address these limitations, but
+expand the analysis to other provinces and territories across Canada.
+This would enhance the generalisability of the results and provide a
+more nuanced understanding of regional differences. Furthermore,
+utilizing nonlinear modelling techniques or looking at a broader set of
+predictor variables such as donor demographics or economic indicators
+could better capture the nuances of donor behaviour and also mitigate
+the risk of omitted variable bias. Comparative studies with other
+democratic countries, beyond Canada, could provide more general
+perspectives on how other political systems and regulatory frameworks
+influence or are influenced by donation behaviours. Other valuable areas
+of future exploration include assessing the impact of media coverage on
+donation trends. Finally, by addressing these areas, research can build
+on the current study's findings and contribute to a greater
+understanding of political donations.
 
@@
 
-The dataset used in this analysis underwent a rigorous data cleaning
-process to ensure accuracy, consistency, and reliability. The raw
-donation records were sourced from the Investigative Journalism
-Foundation (IJF), encompassing political donations at both the federal
-and Ontario provincial levels from 2013 to 2024. The data cleaning
-process involved several critical steps, detailed below.
-
-Initially, the dataset was examined for missing or incomplete values in
-key variables such as Amount Donated, Political Party, Donation Year,
-and Region. Records with missing values in these essential fields were
-excluded from the analysis to maintain data integrity and prevent biases
-that could arise from incomplete data entries.
-
-Political party names often appeared in various formats due to
-inconsistencies in reporting. To address this, party names were
-standardized across all records. For instance, variations such as
-"Liberal Party of Ontario" and "Ontario Liberal Party" were consolidated
-under the standardized name "Ontario Liberal Party." This
-standardization ensured accurate attribution of donations to the correct
-parties.
-
-Donor information, including names and locations, was also standardized
-where possible. Typographical errors were corrected, and formatting was
-made consistent (e.g., "First Name Last Name" instead of "Last Name,
-First Name"). However, original donor names were retained in cases where
-standardization could introduce inaccuracies, particularly when dealing
-with individuals with similar names.
-
-The dataset included both monetary (amount_monetary) and non-monetary
-(amount_non_monetary) contributions. A new variable, Amount Donated, was
-created by summing these two fields for each record. This provided a
-comprehensive measure of the total value of donations received by each
-party. All monetary values were converted to a consistent currency
-format, and any discrepancies in currency symbols or formats were
-corrected to ensure accurate aggregation and comparison of donation
-amounts.
-
-To facilitate analysis at the party-year level, individual donation
-records were aggregated. Donations were summed for each political party
-within each year, resulting in a dataset where each entry represented
-the total donations received by a party in a specific year. This
-aggregation allowed for the examination of trends over time and
-assessment of the impact of factors such as power status and election
-years on donation amounts.
+Rigorous data cleaning was done to the dataset used in this
+investigation to assure correctness and dependability. The raw donation
+records were sourced from the Investigative Journalism Foundation (IJF)
+which include political donations at both federal and provincial levels,
+specifically Ontario, from 2013 to 2024. The details of the data
+cleaning processed are described below.
+
+First, missing and incomplete values were searched for in key variables
+such as Amount Donated, Donation Year, Political Party and Region.
+Excluding any records with missing values in these crucial fields to
+minimise biases created from incomplete data entries and to maintain
+data integrity.
+
+Inconsistencies in reporting caused political party names to appear in
+different formats. To fix this, party names were given constant and
+standardized names across all the records. For example, if there was a
+variation such as "Liberal Party of Ontario" and "Ontario Liberal Party"
+these were combined into a single name "Ontario Liberal Party" to ensure
+accurate attribution of donations to parties.
+
+Fields that were also standardized were the donor information which
+includes names and locations. Any typographical errors were corrected,
+and formatting was made consistent, for example the structure "First
+Name Last Name" instead of "Last Name, First Name". But, in cases where
+standardization could introduce inaccuracies, original donor names were
+kept to avoid dealing with individuals with similar names.
+
+Two columns, (amount_monetary) and (amount_non_monetary) which
+correspond to monetary and non-monetary contributes are included in the
+data set. Using these a new variable was created that sums these two
+fields together for every record. In doing this, the dataset now
+provides a comprehensive measure of the total value of donations
+received by a party. Moreover, monetary values were converted into a
+consistent currency format to disregard any discrepancies in the
+currency symbols or formatting and to ensure accurate aggregation and
+proper comparison.
+
+Individual donation records were aggregated in order to perform proper
+analysis at the party-year level. For each political party within each
+year donations were summed yielding a dataset where each entry
+represented the total donations received by a party for a specific year.
+Aggregating the data like this enabled the examination of trends over
+time and impact assessment factors like power status and election years
+on donation amounts.
 
@@
 Indicates whether the donation was made during an election year (1 for
-election year, 0 for non-election year). These variables were crucial
-for the regression analysis, allowing for the assessment of the
-influence of these factors on total donations.
+election year, 0 for non-election year). These variables were imperative
+for the regression analysis and allowed the assessment of the influence
+of these on total donations.
 
 The distribution of total donation amounts was highly skewed, with a
-small number of large donations and many smaller ones. To address this
-skewness and stabilize variance, a logarithmic transformation was
-applied to the Total Donations variable, creating Log_Total_Donations.
-This transformation also helped in handling zero values by adding a
-small constant (e.g., 1) before taking the logarithm.
-
-To ensure the accuracy of the cleaned dataset, cross-validation was
-performed against original records and official financial statements
-where available. This step involved spot-checking aggregated totals and
-individual records to confirm that the cleaning and aggregation
-processes did not introduce errors.
-
-Some challenges were encountered due to inconsistencies in the raw data,
-such as typographical errors, inconsistent reporting formats, and
-missing donor types. While efforts were made to standardize and correct
-these issues, some limitations remain due to the quality of the original
-data sources. In cases where donor type information was missing and
-legal restrictions indicated that only individuals could donate (e.g.,
-Ontario after 2017), the donor type was inferred as "Individual." While
-this assumption is reasonable based on the legal context, it may not
-capture all nuances in the data.
+small number of large donations and many smaller ones. In order to
+address this, a logarithmic transformation was applied to the
+Total_Donations variable creating a new variable called
+Log_Total_Donations. This transformation also helped with handling zero
+values by adding a small constant before taking the logarithm.
+
+Cross-validation was used against original records and official
+financial statements to ensure accuracy of the cleaned dataset. This
+step required spot-checking aggregated totals and individuals records to
+confirm that the cleaning and aggregation process did not create more
+errors.
+
+There were challenges that occurred due to inconsistencies in the raw
+data, these included typographical errors, reporting format errors and
+missing donor types. Some limitations remained despite the effort to
+standardize and correct these issues, due to the quality of the original
+data source. When donor type information was missing and legal
+restrictions indicated that only individuals could donate, the donor
+type was inferred as "Individual". Despite this assumption being legally
+reasonable in the context of the paper, it may not capture all nuances
+in the data.
 
@@
 approach to collect and compile political donation data across Canada.
-Understanding the IJF's data collection methodology is essential for
-evaluating the reliability and scope of the data used in this analysis.
-
-**Data Sources and Acquisition**
+Understanding the IJF's data collection methodology is crucial for
+assessing the reliability and scope of the data used in this analysis.
+
+Data Sources and Acquisition
 
@@
 
--   Parties: Registered political parties at the federal and provincial
-    levels.
+-   Parties: Federal and provincial registered political parties.
 
@@
 
-Contribution limits and the types of entities permitted to donate vary
-across jurisdictions. For instance, at the federal level, only
+Jurisdictions have varying limits on the donation amount and types of
+entities allowed to donate. For example, in the federal level, only
 individuals who are Canadian citizens or permanent residents can donate,
-with maximum annual contributions set at specified amounts. In contrast,
-some provinces allow donations from corporations, unions, and other
-organizations and may have different or no contribution limits.
+with a maximum annual contribution amount. Comparatively, there are some
+provinces that allow donations from corporations, organizations and
+unions and may have different or no contribution limits.
 
@@
 
-Extensive manual cleaning is performed to correct errors arising from
-OCR conversion and to standardize the data. Key steps in the data
-cleaning process include:
+Manual cleaning is extensively done to correct errors that arise from
+OCR conversion and to standardize the data. The key steps are:
 
@@
 
-Original records submitted by political entities may contain
-typographical errors or inaccuracies, which are reflected in the
-dataset. Incomplete or Inconsistent Data: Variations in reporting
-requirements and practices across jurisdictions can lead to
-inconsistencies in the data. OCR Limitations: Despite manual
-verification, OCR technology may not perfectly capture handwritten or
-poorly scanned documents, potentially introducing errors. Conclusion
-
-### Idealized Methodology
+Records submitted by political entities can contain typographical errors
+or inaccuracies which are seen in the dataset. Inconsistent or
+incomplete data can occur in reporting practices across different
+jurisdictions. Moreover, OCR technology is imperfect at capturing
+handwritings or poorly scanned documents introducing errors despite
+manual correction.
+
+**Idealized Methodology**
 
@@
 proportional representation from various demographics and regions. The
-sample will be stratified based on:
+sample will be stratified according to:
 
@@
 
-To maintain a national margin of error of ±3% at a 95% confidence level,
-a total sample size of approximately 1,500 respondents is required. This
-sample size allows for meaningful analysis of subgroups within the data.
-Given an estimated response rate of 10%, outreach efforts will target
-approximately 15,000 potential respondents across Canada.
+To maintain a national margin of error of plus or minus 3% at a 95%
+confidence level, a total sample size of approximately 1,500 respondents
+is required. This sample size allows for meaningful analysis of
+subgroups within the data. Given an estimated response rate of 10%,
+outreach efforts will target approximately 15,000 potential respondents
+across Canada.
 
@@
 
-To incentivize participation, all respondents will be offered a small
-monetary compensation (e.g., a \$10 electronic gift card) or entered
-into a national sweepstakes with a chance to win one of ten \$500 cash
-prizes.
+To incentivize participation, all respondents will be entered into a
+national sweepstakes with a chance to win one of ten \$500 cash prizes.
 
@@
 
--   nclude attention-check questions to identify inattentive or
+-   Include attention-check questions to identify inattentive or
     fraudulent responses.
@@
         standards. Additionally, funds account for bilingual translation
-        services (English and French).
+        services (English and French). This also covers the cost of the
+        potential monetary gifts given after completing the survey.
 
@@
 
-Timeline
+**Timeline**
 
@@
 
-### Idealized Survey Design
-
-Survey Structure
-
-[Purpose of survey]
-You can access the survey here: https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform?usp=sharing
+#### Idealized Survey Design
+
+**Survey Structure**
 
@@
 
-**State of Residence:** - Dropdown list of all provinces and
-territories.
-
-**Age:** Multiple-choice options
+State of Residence:
+
+-   Dropdown list of all provinces and territories.
+
+Age: Multiple-choice options
 
@@
 
-**Gender:**
+Gender:
 
@@
 
-**Income Level:**
+Income Level:
 
@@
 
-**Education Level:** Highest degree obtained
+Education Level: Highest degree obtained
 
@@
 
-**Political Affiliation:**
+Political Affiliation:
 
@@
 
-**Are you currently registered to donate to political parties?**
+Are you currently registered to donate to political parties?
 
@@
 
-**How frequently do you donate to political parties?**
+How frequently do you donate to political parties?
 
@@
 
-**Average Donation Amount:**
+Average Donation Amount:
 
@@
 
-**Preferred Donation Channels:**
+Preferred Donation Channels:
 
@@
 
-**What motivates you to donate to a political party?** (Select all that
+What motivates you to donate to a political party? (Select all that
 apply)
@@
 
-**Which factors most influence your decision to donate?** (Select all
-that apply)
+Which factors most influence your decision to donate? (Select all that
+apply)
 
@@
 
-**Have recent political events influenced your donation behavior?**
+Have recent political events influenced your donation behavior?
 
@@
 
-**If yes, please specify the events and how they influenced your
-donations.** (Open-ended)
-
-**How do you perceive the effectiveness of political campaigns in
-encouraging donations?**
+If yes, please specify the events and how they influenced your
+donations. (Open-ended)
+
+How do you perceive the effectiveness of political campaigns in
+encouraging donations?
 
@@
 
-**What could political parties do to encourage more donations from
-supporters?** (Open-ended) -
-
-**Any additional comments or suggestions regarding political
-donations?** (Open-ended)
+What could political parties do to encourage more donations from
+supporters? (Open-ended) -
+
+Any additional comments or suggestions regarding political donations?
+(Open-ended)
 

### 2024-12-04 paper/paper.qmd
Update results section

@@
 
+## Ontario Results 
+In the Ontario model, Party_Size is a significant predictor of total donations, with a coefficient of 2.956 (standard error = 0.485, p < 0.001).
+This indicates that major parties in Ontario receive significantly higher total donations than minor parties, highlighting the influence of party prominence on fundraising success.
+
+The In_Power variable has a coefficient of 0.884 (standard error = 0.540, p = 0.106), which is not statistically significant at the conventional 5% level.
+This suggests that being the ruling party in Ontario does not have a significant direct effect on the total amount of donations received.
+
+The Election_Year variable is also not significant in the Ontario model (coefficient = 0.239, standard error = 0.436, p = 0.586), indicating that donations are not substantially influenced by whether they occur during an election year.
+
+## Federal Results
+In the Federal model, Party_Size remains a significant predictor, with a coefficient of 3.450 (standard error = 0.259, p < 0.001).
+This reinforces the importance of party size in attracting donations at the federal level.
+
+The In_Power variable in the Federal model has a coefficient of 0.637 (standard error = 0.3882, p = 0.103), which is not statistically significant.
+Similar to the Ontario context, being in power at the federal level does not have a significant direct effect on the total donations received.
+
+The Election_Year variable is not significant in the Federal model (coefficient = -0.0752, standard error = 0.2209, p = 0.734), suggesting no substantial impact of election years on donation amounts at the federal level.
+
+## Conservative/Liberal Results
+In the Conservative/Liberal model, In_Power has a significant positive effect, with a coefficient of 1.202 (standard error = 0.497, p = 0.01827).
+Additionally, the interaction term In_Power × Party is significantly negative, with a coefficient of -1.718 (standard error = 0.711, p = 0.01845).
+This interaction indicates that the effect of being in power on total donations varies between the Conservative and Liberal parties.
+
+For the Conservative Party (Party = 1), the combined effect of being in power is 1.2023 + (-1.7175) = -0.5152, suggesting that being in power is associated with lower total donations for the Conservative Party.
+For the Liberal Party (Party = 0), the effect of being in power remains 1.2023, indicating that being in power is associated with higher total donations for the Liberal Party.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
@@
 
+This survey helps to gather data about those who donate to political
+parties and acquire more qualitative details about a sample of the
+population.
+
+The survey can be accessed using [this
+link](https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform).
+
 **Survey Structure**

### 2024-12-04 paper/paper.qmd
update results

@@
 
+The Party variable itself is significant (coefficient = 1.3648, standard error = 0.4968, p = 0.00774), suggesting that, overall, the Conservative Party tends to receive more donations than the Liberal Party when not considering the interaction effect.
+
+The Election_Year variable is not significant in the Conservative/Liberal model (coefficient = -0.0527, standard error = 0.3777, p = 0.88946).
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
@@
   labs(
-    title = "Coefficient Estimates for Ontario, Federal, and Conservative/Liberal Models",
     x = "Predictors",
@@
 ```
-
-The regression analysis reveals that Party Size is a significant
-predictor of total donations across both the Ontario and Federal models,
-indicating that larger parties receive more financial support.
-Specifically, in the Ontario model, Party Size has a coefficient of
-2.302 (p \< 0.001), while in the Federal model, it is 3.524 (p \<
-0.001). This positive relationship suggests that party prominence plays
-a crucial role in attracting donations.
-
-Conversely, the In Power variable is not statistically significant in
-the Ontario (p = 0.322) and Federal (p = 0.529) models, implying that
-being in power does not directly influence the total donations received
-by a party at these levels. However, in the Conservative/Liberal model,
-In Power has a significant positive effect (2.513, p \< 0.001), and the
-interaction term In Power \* Party is significantly negative (-4.865, p
-\< 0.001). This interaction indicates that while being in power
-increases donations for Conservative parties, it decreases donations for
-Liberal parties, highlighting a differential impact based on party
-affiliation.
-
-The Election Year variable does not significantly predict donation
-amounts in any of the models, with p-values exceeding conventional
-significance thresholds (Ontario p = 0.616, Federal p = 0.975,
-Conservative/Liberal p = 0.325). This suggests that donations are not
-substantially influenced by whether they occur during an election year.
-
-Overall, the models explain a moderate portion of the variance in total
-donations, with R² values ranging from 0.3046 (Ontario) to 0.5825
-(Federal), and the Conservative/Liberal model achieving an R² of 0.4481.
-These findings underscore the importance of party size in determining
-donation amounts, while the role of power status appears to be
-contingent on party affiliation.
+@fig-estimates highlights the consistently strong positive effect of Party_Size on total donations in both the Ontario and Federal models, emphasizing that major parties attract significantly more donations than minor parties.
+It also illustrates the significant interaction effect between In_Power and Party in the Conservative/Liberal model, demonstrating the differing impact of being in power on donations for the two major parties.
+
+In summary, the analyses reveal that party size is a significant determinant of total donations at both provincial and federal levels, while the effect of being in power on donations is contingent upon party affiliation, particularly between the Conservative and Liberal parties.
 

### 2024-12-04 paper/paper.qmd
bolded a few words

@@
 
-The models are specified as follows for the Ontario and Federal Models:
+The models are specified as follows for the **Ontario and Federal
+Models**:
 
@@
 ```
-For the Conservative and Liberal parties, we included an interaction
+For the **Conservative and Liberal parties**, we included an interaction
 term:
@@
 Where:
+
 - \(\text{Party}_i\): A binary variable indicating 1 for Conservative, 0 for Liberal.\\
 ```
-Each model uses the logarithm of total donations (Log_Total_Donations)
-as the dependent variable to stabilize variance and normalize the
-distribution of donation amounts. Predictor variables include binary
-indicators for whether the party was in power (In_Power), the size of
-the party (Party_Size), and whether the donation was made during an
-election year (Election_Year).
-
 The specialized model for Conservative and Liberal parties includes an
@@
 ## Ontario Results
-In the Ontario model, Party_Size is a significant predictor of total donations, with a coefficient of 2.956 (standard error = 0.485, p < 0.001).
-This indicates that major parties in Ontario receive significantly higher total donations than minor parties, highlighting the influence of party prominence on fundraising success.
-
-The In_Power variable has a coefficient of 0.884 (standard error = 0.540, p = 0.106), which is not statistically significant at the conventional 5% level.
-This suggests that being the ruling party in Ontario does not have a significant direct effect on the total amount of donations received.
-
-The Election_Year variable is also not significant in the Ontario model (coefficient = 0.239, standard error = 0.436, p = 0.586), indicating that donations are not substantially influenced by whether they occur during an election year.
+
+In the Ontario model, Party_Size is a significant predictor of total
+donations, with a coefficient of 2.956 (standard error = 0.485, p \<
+0.001). This indicates that major parties in Ontario receive
+significantly higher total donations than minor parties, highlighting
+the influence of party prominence on fundraising success.
+
+The In_Power variable has a coefficient of 0.884 (standard error =
+0.540, p = 0.106), which is not statistically significant at the
+conventional 5% level. This suggests that being the ruling party in
+Ontario does not have a significant direct effect on the total amount of
+donations received.
+
+The Election_Year variable is also not significant in the Ontario model
+(coefficient = 0.239, standard error = 0.436, p = 0.586), indicating
+that donations are not substantially influenced by whether they occur
+during an election year.
 
 ## Federal Results
-In the Federal model, Party_Size remains a significant predictor, with a coefficient of 3.450 (standard error = 0.259, p < 0.001).
-This reinforces the importance of party size in attracting donations at the federal level.
-
-The In_Power variable in the Federal model has a coefficient of 0.637 (standard error = 0.3882, p = 0.103), which is not statistically significant.
-Similar to the Ontario context, being in power at the federal level does not have a significant direct effect on the total donations received.
-
-The Election_Year variable is not significant in the Federal model (coefficient = -0.0752, standard error = 0.2209, p = 0.734), suggesting no substantial impact of election years on donation amounts at the federal level.
+
+In the Federal model, Party_Size remains a significant predictor, with a
+coefficient of 3.450 (standard error = 0.259, p \< 0.001). This
+reinforces the importance of party size in attracting donations at the
+federal level.
+
+The In_Power variable in the Federal model has a coefficient of 0.637
+(standard error = 0.3882, p = 0.103), which is not statistically
+significant. Similar to the Ontario context, being in power at the
+federal level does not have a significant direct effect on the total
+donations received.
+
+The Election_Year variable is not significant in the Federal model
+(coefficient = -0.0752, standard error = 0.2209, p = 0.734), suggesting
+no substantial impact of election years on donation amounts at the
+federal level.
 
 ## Conservative/Liberal Results
-In the Conservative/Liberal model, In_Power has a significant positive effect, with a coefficient of 1.202 (standard error = 0.497, p = 0.01827).
-Additionally, the interaction term In_Power × Party is significantly negative, with a coefficient of -1.718 (standard error = 0.711, p = 0.01845).
-This interaction indicates that the effect of being in power on total donations varies between the Conservative and Liberal parties.
-
-For the Conservative Party (Party = 1), the combined effect of being in power is 1.2023 + (-1.7175) = -0.5152, suggesting that being in power is associated with lower total donations for the Conservative Party.
-For the Liberal Party (Party = 0), the effect of being in power remains 1.2023, indicating that being in power is associated with higher total donations for the Liberal Party.
-
-The Party variable itself is significant (coefficient = 1.3648, standard error = 0.4968, p = 0.00774), suggesting that, overall, the Conservative Party tends to receive more donations than the Liberal Party when not considering the interaction effect.
-
-The Election_Year variable is not significant in the Conservative/Liberal model (coefficient = -0.0527, standard error = 0.3777, p = 0.88946).
+
+In the Conservative/Liberal model, In_Power has a significant positive
+effect, with a coefficient of 1.202 (standard error = 0.497, p =
+0.01827). Additionally, the interaction term In_Power × Party is
+significantly negative, with a coefficient of -1.718 (standard error =
+0.711, p = 0.01845). This interaction indicates that the effect of being
+in power on total donations varies between the Conservative and Liberal
+parties.
+
+For the Conservative Party (Party = 1), the combined effect of being in
+power is 1.2023 + (-1.7175) = -0.5152, suggesting that being in power is
+associated with lower total donations for the Conservative Party. For
+the Liberal Party (Party = 0), the effect of being in power remains
+1.2023, indicating that being in power is associated with higher total
+donations for the Liberal Party.
+
+The Party variable itself is significant (coefficient = 1.3648, standard
+error = 0.4968, p = 0.00774), suggesting that, overall, the Conservative
+Party tends to receive more donations than the Liberal Party when not
+considering the interaction effect.
+
+The Election_Year variable is not significant in the
+Conservative/Liberal model (coefficient = -0.0527, standard error =
+0.3777, p = 0.88946).
 
@@
 ```
-@fig-estimates highlights the consistently strong positive effect of Party_Size on total donations in both the Ontario and Federal models, emphasizing that major parties attract significantly more donations than minor parties.
-It also illustrates the significant interaction effect between In_Power and Party in the Conservative/Liberal model, demonstrating the differing impact of being in power on donations for the two major parties.
-
-In summary, the analyses reveal that party size is a significant determinant of total donations at both provincial and federal levels, while the effect of being in power on donations is contingent upon party affiliation, particularly between the Conservative and Liberal parties.
+
+@fig-estimates highlights the consistently strong positive effect of
+Party_Size on total donations in both the Ontario and Federal models,
+emphasizing that major parties attract significantly more donations than
+minor parties. It also illustrates the significant interaction effect
+between In_Power and Party in the Conservative/Liberal model,
+demonstrating the differing impact of being in power on donations for
+the two major parties.
+
+In summary, the analyses reveal that party size is a significant
+determinant of total donations at both provincial and federal levels,
+while the effect of being in power on donations is contingent upon party
+affiliation, particularly between the Conservative and Liberal parties.
 

### 2024-12-04 scripts/03-clean_data.R
update cleaning data

@@ -9,6 +9,8 @@
 library(dplyr)
 library(lubridate)
 library(readr)
+library(here)
+library(arrow)  
 
 #### Clean data ####
 raw_data <- read_csv("data/01-raw_data/raw_data.csv")
@@ -130,3 +132,6 @@ donations_filtered <- donations_filtered %>%
 
 #### Save data ####
 write_csv(donations_filtered, "data/02-analysis_data/analysis_data.csv")
+
+write_parquet(x = donations_filtered, sink = here("data", "02-analysis_data", "analysis_data.parquet"))
+

### 2024-12-04 paper/paper.qmd
update discussion heading

@@
 
-## The Influence of Party Size on Donation Amounts {#sec-first-point}
+## Financial influence shapes political landscapes
 
@@
 solicit funding and political support in exchange for advancing the
-interests of their stakeholders @mollona2022ethics}. These two dynamics
+interests of their stakeholders @mollona2022ethics. These two dynamics
 may be working in tandem to create an effect where parties eager to be
@@
 superdonors, widening the resource gap between large parties
-@draca2022financing}. Similar to Canada, the UK also sees large
+@draca2022financing. Similar to Canada, the UK also sees large
 donations from a wealthy few influencing their elections each year. The

### 2024-12-14 paper/paper.qmd
update discussion and references

@@
 often benefit from more donations, particularly from interest groups
-seeking to maintain influence (@fouirnaies2014financial). However, there
+seeking to maintain influence [@fouirnaies2014financial]. However, there
 has been less focus on whether this holds true in Canada, where the
@@
 
-The analysis was conducted using both R (@citeR) and Python (@python),
+The analysis was conducted using both R [@citeR] and Python [@python],
 leveraging a range of libraries for data manipulation and visualization.
-In R, key packages included dplyr (@dplyr), lubridate (@lubridate),
-tidyverse (@tidyverse), ggplot2 (@ggplot), readr (@readr), tibble
-(@tibble), and rstanarm (@rstanarm) for statistical modeling. In Python,
-pandas (@pandas), requests (@requests), and matplotlib (@matplotlib)
+In R, key packages included dplyr [@dplyr], lubridate [@lubridate],
+tidyverse [@tidyverse], ggplot2 [@ggplot], readr [@readr], tibble
+[@tibble], and rstanarm [@rstanarm] for statistical modeling. In Python,
+pandas [@pandas], requests [@requests], and matplotlib [@matplotlib]
 were utilized for gathering the data.
@@
 
-At the heart of politics is finance, as revealed by Traag's paper on the
-complex contagion of campaign donations @traag2016complex. Their study
-demonstrates that a majority of campaign contributions in the US
-originate from a few wealthy individuals through large social networks.
-With the constant battle between government and corporations, the
-findings in this study underscore the significant influence of financial
-power in shaping political dynamics.
-
-The landscape in Canada for political financing is growing each year.
-Results from studies demonstrate that individual contributors may prefer
-to support challengers for several reasons. An optimistic but naive
-explanation may be that individuals hope to foster competitive
-elections, given that donations depend not on political leaning but on
-who is currently in power. However, a more realistic explanation for
-this phenomenon is that some individuals want to make policy changes. It
-is no secret that politicians, particularly those out of power, often
-solicit funding and political support in exchange for advancing the
-interests of their stakeholders @mollona2022ethics. These two dynamics
-may be working in tandem to create an effect where parties eager to be
-elected rely more on funding from external sources, which can switch
-each election cycle.
-
-Moreover, these trends of political donations fluctuating based on the
-party in power are not unique to Canada. The study "Financing UK
-Democracy: A Stocktake of 20 Years of Political Donations" highlights a
-surge in private political donations, more specifically from
-superdonors, widening the resource gap between large parties
-@draca2022financing. Similar to Canada, the UK also sees large
-donations from a wealthy few influencing their elections each year. The
-parallel between Canada and the UK underscores a broader democratic
-issue of financial power being used to influence political outcomes.
-Though this concept is not novel, this study contributes to the already
-growing spotlight on the persistent lack of proper reform in political
-systems.
+At the heart of politics is finance, as revealed by Traag’s research on the complex contagion of campaign donations in the United States [@traag2016complex].
+Traag’s study found that most campaign contributions come from a small number of wealthy donors linked by large social networks, demonstrating how a few individuals can shape political outcomes.
+Similar dynamics appear to be at play in Canada.
+Here, donors often prefer to support parties not currently in power.
+One reason might be the hope that challengers will bring needed policy changes.
+More realistically, as noted by Mollona [@mollona2022ethics], politicians, especially those out of power, often seek funding in exchange for promoting donors’ interests.
+Over time, this can lead to a pattern where donors shift their support to whoever best serves their goals, fostering continuous political competition.
+A key finding of this study is that at both the Ontario and federal levels, the Conservative Party generally attracts more financial backing than other parties.
+This steady flow of resources may help the Conservatives maintain a strong presence in the political arena, influencing agenda-setting and public debate.
+With more funds, they can more easily finance voter outreach, shape their messaging, and solidify their support networks.
+While this financial advantage may help sustain dynamic election contests, it also underscores concerns about whose interests gain the most traction.
+When one party consistently out-raises others, the policy conversation may lean toward those who provide the most money, potentially sidelining issues important to groups with fewer financial resources.
+
+## International Comparisons and the Importance of Regulation
+These trends are not confined to Canada.
+In the United Kingdom, for example, large private donations, often from very wealthy “superdonors”, widen the gap between major parties and smaller competitors [@draca2022financing].
+Such patterns highlight a broader democratic challenge, as money can concentrate power and influence political direction in ways that may not always reflect broad public preferences.
+Rather than serving purely as a resource that encourages healthy competition, political donations can also become a tool that benefits certain parties and voices more than others.
+Regulation and transparency are essential.
+Clearer rules and better public reporting help citizens understand who funds whom and why.
+Drawing lessons from other countries may guide reforms aimed at limiting undue influence.
+Stronger oversight, spending caps, and easy-to-access donor databases can empower voters to make informed choices.
+Without these safeguards, the already powerful, whether individuals, industries, or interest groups, can gain even more sway, risking an erosion of trust in democratic institutions.
 
@@
 accuracy of the findings here. Lastly, the linear regression models here
-assume a linear relationship between predictors and donations which can
-greatly oversimplify the complexities.
+assume a linear relationship between predictors and donations which can oversimplify the complexities.
 
@@
 expand the analysis to other provinces and territories across Canada.
-This would enhance the generalisability of the results and provide a
+This would enhance the generalizability of the results and provide a
 more nuanced understanding of regional differences. Furthermore,
@@
 
-Data Sources and Acquisition
+**Data Sources and Acquisition**
 
@@
 
-**Idealized Methodology**
+##Idealized Methodology
 

### 2024-12-14 paper/paper.qmd
finalized idealized survey

@@
 This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
-format: pdf
+format: 
+  pdf:
+    toc: true
+    toc-depth: 3
 number-sections: true
@@
 
-## Financial influence shapes political landscapes
-
-At the heart of politics is finance, as revealed by Traag’s research on the complex contagion of campaign donations in the United States [@traag2016complex].
-Traag’s study found that most campaign contributions come from a small number of wealthy donors linked by large social networks, demonstrating how a few individuals can shape political outcomes.
-Similar dynamics appear to be at play in Canada.
-Here, donors often prefer to support parties not currently in power.
-One reason might be the hope that challengers will bring needed policy changes.
-More realistically, as noted by Mollona [@mollona2022ethics], politicians, especially those out of power, often seek funding in exchange for promoting donors’ interests.
-Over time, this can lead to a pattern where donors shift their support to whoever best serves their goals, fostering continuous political competition.
-A key finding of this study is that at both the Ontario and federal levels, the Conservative Party generally attracts more financial backing than other parties.
-This steady flow of resources may help the Conservatives maintain a strong presence in the political arena, influencing agenda-setting and public debate.
-With more funds, they can more easily finance voter outreach, shape their messaging, and solidify their support networks.
-While this financial advantage may help sustain dynamic election contests, it also underscores concerns about whose interests gain the most traction.
-When one party consistently out-raises others, the policy conversation may lean toward those who provide the most money, potentially sidelining issues important to groups with fewer financial resources.
+## Financial Influence Shapes Political Landscapes
+
+At the heart of politics is finance, as revealed by Traag’s research on
+the complex contagion of campaign donations in the United States
+[@traag2016complex]. Traag’s study found that most campaign
+contributions come from a small number of wealthy donors linked by large
+social networks, demonstrating how a few individuals can shape political
+outcomes. Similar dynamics appear to be at play in Canada. Here, donors
+often prefer to support parties not currently in power. One reason might
+be the hope that challengers will bring needed policy changes. More
+realistically, as noted by Mollona [@mollona2022ethics], politicians,
+especially those out of power, often seek funding in exchange for
+promoting donors’ interests. Over time, this can lead to a pattern where
+donors shift their support to whoever best serves their goals, fostering
+continuous political competition. A key finding of this study is that at
+both the Ontario and federal levels, the Conservative Party generally
+attracts more financial backing than other parties. This steady flow of
+resources may help the Conservatives maintain a strong presence in the
+political arena, influencing agenda-setting and public debate. With more
+funds, they can more easily finance voter outreach, shape their
+messaging, and solidify their support networks. While this financial
+advantage may help sustain dynamic election contests, it also
+underscores concerns about whose interests gain the most traction. When
+one party consistently out-raises others, the policy conversation may
+lean toward those who provide the most money, potentially sidelining
+issues important to groups with fewer financial resources.
 
 ## International Comparisons and the Importance of Regulation
-These trends are not confined to Canada.
-In the United Kingdom, for example, large private donations, often from very wealthy “superdonors”, widen the gap between major parties and smaller competitors [@draca2022financing].
-Such patterns highlight a broader democratic challenge, as money can concentrate power and influence political direction in ways that may not always reflect broad public preferences.
-Rather than serving purely as a resource that encourages healthy competition, political donations can also become a tool that benefits certain parties and voices more than others.
-Regulation and transparency are essential.
-Clearer rules and better public reporting help citizens understand who funds whom and why.
-Drawing lessons from other countries may guide reforms aimed at limiting undue influence.
-Stronger oversight, spending caps, and easy-to-access donor databases can empower voters to make informed choices.
-Without these safeguards, the already powerful, whether individuals, industries, or interest groups, can gain even more sway, risking an erosion of trust in democratic institutions.
+
+These trends are not confined to Canada. In the United Kingdom, for
+example, large private donations, often from very wealthy “superdonors”,
+widen the gap between major parties and smaller competitors
+[@draca2022financing]. Such patterns highlight a broader democratic
+challenge, as money can concentrate power and influence political
+direction in ways that may not always reflect broad public preferences.
+Rather than serving purely as a resource that encourages healthy
+competition, political donations can also become a tool that benefits
+certain parties and voices more than others. Regulation and transparency
+are essential. Clearer rules and better public reporting help citizens
+understand who funds whom and why. Drawing lessons from other countries
+may guide reforms aimed at limiting undue influence. Stronger oversight,
+spending caps, and easy-to-access donor databases can empower voters to
+make informed choices. Without these safeguards, the already powerful,
+whether individuals, industries, or interest groups, can gain even more
+sway, risking an erosion of trust in democratic institutions.
 
@@
 accuracy of the findings here. Lastly, the linear regression models here
-assume a linear relationship between predictors and donations which can oversimplify the complexities.
+assume a linear relationship between predictors and donations which can
+oversimplify the complexities.
 
@@
 
-##Idealized Methodology
+## Appendix C: Idealized Methodology
 
@@
 
--   Details: Expenses related to accessing online panel respondents,
-    purchasing targeted advertising on social media, and collaborating
-    with panel providers. Funds also cover the services of statisticians
-    to ensure the sampling aligns with the population distribution.
+    -   Expenses related to accessing online panel respondents,
+        purchasing targeted advertising on social media, and
+        collaborating with panel providers. Funds also cover the
+        services of statisticians to ensure the sampling aligns with the
+        population distribution.
 
@@
 
--   Details: Allocated for compensating participants with \$10
-    electronic gift cards or funding the national sweepstakes prizes.
-    Incentives are crucial for improving response rates and attracting a
-    diverse respondent pool.
+    -   Allocated for compensating participants with \$10 electronic
+        gift cards or funding the national sweepstakes prizes.
+        Incentives are crucial for improving response rates and
+        attracting a diverse respondent pool.
 
@@
 
--   Details: Operational costs for administering the survey, including
-    online platform fees, telephone survey expenses, interviewer
-    training, and call center operations. This ensures data collection
-    is efficient and reaches a broad audience.
+    -   Operational costs for administering the survey, including online
+        platform fees, telephone survey expenses, interviewer training,
+        and call center operations. This ensures data collection is
+        efficient and reaches a broad audience.
 
@@
 
--   Details: Funds dedicated to analyzing the collected data, including
-    hiring data analysts proficient in statistical software, conducting
-    comprehensive analyses, and preparing detailed reports and
-    visualizations to communicate findings effectively.
+    -   Funds dedicated to analyzing the collected data, including
+        hiring data analysts proficient in statistical software,
+        conducting comprehensive analyses, and preparing detailed
+        reports and visualizations to communicate findings effectively.
 
@@
 
--   Details: Covers the costs of obtaining ethical approvals, ensuring
+    -   Covers the costs of obtaining ethical approvals, ensuring
         compliance with privacy legislation (e.g., Personal Information
-    Protection and Electronic Documents Act), and implementing robust
-    data security measures to protect respondent information.
+        Protection and Electronic Documents Act), and implementing
+        robust data security measures to protect respondent information.
 
@@
 parties and acquire more qualitative details about a sample of the
-population.
-
-The survey can be accessed using [this
+population. The survey can be accessed using [this
 link](https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform).
@@
 
+**Introduction**
+
+Thank you for participating in this important survey. Your responses
+will help us understand how Canadians make decisions about donating to
+political parties at both the federal and Ontario provincial levels. By
+sharing your perspectives, you will contribute to a deeper understanding
+of what motivates political donations, what factors shape donor
+decisions, and how parties can improve their engagement with supporters.
+Your answers are completely confidential and will be used solely for
+research purposes.
+
 **Section 1: Demographics**
@@
 
-Age: Multiple-choice options
+Age:
 
@@
 
+-   65+
+
 Gender:
@@
 
-Income Level:
+Annual Household Income Level:
 
@@
 
-Education Level: Highest degree obtained
+-   Prefer not to say
+
+Highest Level of Education Completed:
 
@@
 
+-   Green
+
 -   Other (please specify)
 
+-   No specific affiliation
+
 **Section 2: Donation Behavior**
 
-Are you currently registered to donate to political parties?
-
--   Yes
-
--   No
+Have you ever donated to a political party?
+
+-   Yes, currently donate
+
+-   Yes, but not currently
+
+-   No, never donated
 
@@
 
-Average Donation Amount:
+Average Annual Donation Amount (to all political parties combined):
 
@@
 
-Preferred Donation Channels:
-
--   Online platforms
+-   Prefer not to say
+
+Preferred Donation Channels: (Select all that apply)
+
+-   Online platforms (e.g., party websites, crowdfunding sites)
 
@@
 
--   In-person events
+-   In-person events (e.g., fundraisers, rallies)
+
+-   Phone solicitations
+
+-   Other (please specify)
 
@@
 
+-   Personal connection to a candidate
+
 -   Other (please specify)
@@
 
--   Candidate's leadership qualities
-
--   Party's stance on key issues
-
--   Party's past performance
-
--   Recommendations from peers or influencers
-
--   Media coverage
+-   Candidate’s leadership qualities
+
+-   Party’s stance on key issues
+
+-   Party’s track record and past performance
+
+-   Recommendations from peers, family, or community leaders
+
+-   Media coverage and public opinion polls
+
+-   Urgency or importance of the upcoming election
 
@@
 
-Have recent political events influenced your donation behavior?
+-   Have recent political events (e.g., elections, leadership changes,
+    policy announcements) influenced your donation behavior?
 
@@
 
-If yes, please specify the events and how they influenced your
-donations. (Open-ended)
-
-How do you perceive the effectiveness of political campaigns in
+    If yes, please specify the event(s) and how they influenced your
+    donations:\
+    *(Open-ended)*
+
+-   How do you perceive the effectiveness of political campaigns in
     encouraging donations?
@@
 
+-   Have you ever stopped donating or reduced donations because of a
+    party’s perceived performance or power status (e.g., losing seats,
+    scandals)?
+
+    -   Yes (please explain why) *(Open-ended)*
+
+    -   No
+
 **Section 5: Feedback and Suggestions**
 
-What could political parties do to encourage more donations from
-supporters? (Open-ended) -
-
-Any additional comments or suggestions regarding political donations?
-(Open-ended)
+-   In your opinion, what could political parties do to encourage more
+    donations from supporters?\
+    *(Open-ended)*
+
+-   Any additional comments or suggestions regarding political
+    donations?\
+    *(Open-ended)*
+
+**Thank you for participating!** Your feedback is valuable and will help
+researchers and political parties understand the dynamics of political
+giving in Canada. Your insights contribute to more informed policies and
+practices, ultimately shaping a more responsive and transparent
+political landscape.
+
+\newpage
 

### 2024-12-15 paper/paper.qmd
renamed folder
update model to use real_amount

@@
 
-\newpage
-
 ## Predictor Variables
@@
   summarize(
-    Total_Donations = sum(amount, na.rm = TRUE),
+    Total_Donations = sum(real_amount, na.rm = TRUE),
     In_Power = first(recipient_in_power),
@@
   summarize(
-    Total_Donations = sum(amount, na.rm = TRUE),
+    Total_Donations = sum(real_amount, na.rm = TRUE),
     In_Power = first(In_Power),
@@
 In the Ontario model, Party_Size is a significant predictor of total
-donations, with a coefficient of 2.956 (standard error = 0.485, p \<
+donations, with a coefficient of 3.046 (standard error = 0.499, p \<
 0.001). This indicates that major parties in Ontario receive
@@
 
-The In_Power variable has a coefficient of 0.884 (standard error =
-0.540, p = 0.106), which is not statistically significant at the
+The In_Power variable has a coefficient of 0.886 (standard error =
+0.547, p = 0.116), which is not statistically significant at the
 conventional 5% level. This suggests that being the ruling party in
@@
 The Election_Year variable is also not significant in the Ontario model
-(coefficient = 0.239, standard error = 0.436, p = 0.586), indicating
+(coefficient = 0.274, standard error = 0.449, p = 0.544), indicating
 that donations are not substantially influenced by whether they occur
@@
 
-In the Federal model, Party_Size remains a significant predictor, with a
-coefficient of 3.450 (standard error = 0.259, p \< 0.001). This
-reinforces the importance of party size in attracting donations at the
-federal level.
-
-The In_Power variable in the Federal model has a coefficient of 0.637
-(standard error = 0.3882, p = 0.103), which is not statistically
-significant. Similar to the Ontario context, being in power at the
-federal level does not have a significant direct effect on the total
-donations received.
-
-The Election_Year variable is not significant in the Federal model
-(coefficient = -0.0752, standard error = 0.2209, p = 0.734), suggesting
-no substantial impact of election years on donation amounts at the
-federal level.
+At the federal level, Party_Size once again emerges as a significant determinant of total donations, with a coefficient of 3.452 (standard error = 0.259, p \< 0.001).
+This confirms that larger, more established parties are more successful at fundraising on the national stage, similar to the provincial setting.
+
+The In_Power variable is not statistically significant in the federal model (coefficient = 0.637, standard error = 0.389, p = 0.103), implying that being the governing party at the federal level does not confer a clear advantage in terms of donation totals.
+
+Similarly, Election_Year does not significantly impact federal donations (coefficient = -0.042, standard error = 0.221, p = 0.849), indicating no substantial change in donor behavior during election years at the federal level.
 
@@
 
-In the Conservative/Liberal model, In_Power has a significant positive
-effect, with a coefficient of 1.202 (standard error = 0.497, p =
-0.01827). Additionally, the interaction term In_Power × Party is
-significantly negative, with a coefficient of -1.718 (standard error =
-0.711, p = 0.01845). This interaction indicates that the effect of being
-in power on total donations varies between the Conservative and Liberal
-parties.
-
-For the Conservative Party (Party = 1), the combined effect of being in
-power is 1.2023 + (-1.7175) = -0.5152, suggesting that being in power is
-associated with lower total donations for the Conservative Party. For
-the Liberal Party (Party = 0), the effect of being in power remains
-1.2023, indicating that being in power is associated with higher total
-donations for the Liberal Party.
-
-The Party variable itself is significant (coefficient = 1.3648, standard
-error = 0.4968, p = 0.00774), suggesting that, overall, the Conservative
-Party tends to receive more donations than the Liberal Party when not
-considering the interaction effect.
-
-The Election_Year variable is not significant in the
-Conservative/Liberal model (coefficient = -0.0527, standard error =
-0.3777, p = 0.88946).
+In the Conservative/Liberal model, the In_Power variable is now positively and significantly related to total donations (coefficient = 1.201, standard error = 0.513, p = 0.022), while the interaction In_Power × Party is significantly negative (coefficient = -1.711, standard error = 0.734, p = 0.023).
+This interaction means that the effect of being in power depends on whether the party is Conservative or Liberal.
+
+For the Conservative Party (Party = 1), the combined effect of being in power is approximately 1.201 + (-1.711) = -0.510, suggesting that being in power is associated with lower total donations for the Conservative Party.
+For the Liberal Party (Party = 0), the effect of being in power remains 1.201, indicating that being in power is linked to higher donation totals for the Liberal Party.
+
+The Party variable itself is also significant (coefficient = 1.364, standard error = 0.513, p = 0.0099), indicating that, when not considering the interaction, the Conservative Party tends to attract more donations than the Liberal Party.
+The Election_Year variable remains non-significant (coefficient = -0.023, standard error = 0.390, p = 0.953), suggesting no direct election-year influence on donations within this subset.
 
@@
 
-@fig-estimates highlights the consistently strong positive effect of
-Party_Size on total donations in both the Ontario and Federal models,
-emphasizing that major parties attract significantly more donations than
-minor parties. It also illustrates the significant interaction effect
-between In_Power and Party in the Conservative/Liberal model,
-demonstrating the differing impact of being in power on donations for
-the two major parties.
-
-In summary, the analyses reveal that party size is a significant
-determinant of total donations at both provincial and federal levels,
-while the effect of being in power on donations is contingent upon party
-affiliation, particularly between the Conservative and Liberal parties.
+@fig-estimates continues to highlight the consistently strong positive effect of Party_Size on total donations in both the Ontario and Federal models, affirming that well-established parties attract more donor support.
+It also illustrates the significant interaction effect between In_Power and Party in the Conservative/Liberal model, demonstrating that the influence of being in power on donations differs notably between the two major parties.
+
+In summary, the updated results show that party size consistently drives higher donations across jurisdictions, while the impact of being in power on donation amounts depends on the party’s ideological affiliation, especially when contrasting Conservative and Liberal parties.
 

### 2024-12-15 scripts/03-clean_data.R

@@ -130,6 +130,27 @@ donations_filtered <- donations_filtered %>%
     )
   )
 
+
+#  Convert Nominal Amounts to Real Amounts Manually
+
+cpi_data <- data.frame(
+  year = 2006:2024,
+  CPI = c(109.1, 111.5, 114.1, 114.4, 116.5, 119.9, 121.7, 122.8, 125.2, 
+          126.6, 128.4, 130.4, 133.4, 136.0, 137.0, 141.6, 151.2, 157.1, 161.80)
+)
+
+# Specify base year
+base_year <- 2024
+
+# Extract CPI for the base year
+cpi_base <- cpi_data$CPI[cpi_data$year == base_year]
+
+# Merge to add CPI to each donation record by its year
+donations_filtered <- merge(donations_filtered, cpi_data, by.x = "donation_year", by.y = "year", all.x = TRUE)
+
+# Convert nominal amounts to real amounts in 2024 dollars
+donations_filtered$real_amount <- donations_filtered$amount * (cpi_base / donations_filtered$CPI)
+
 #### Save data ####
 write_csv(donations_filtered, "data/02-analysis_data/analysis_data.csv")
 

### 2024-12-15 scripts/05-exploratory_data_analysis.R

@@ -36,6 +36,7 @@ summary_stats_ontario <- data.frame(
 )
 write_csv(summary_stats_ontario, "data/02-analysis_data/02.1-exploration_data/Ontario/summary_statistics_ontario.csv")
 
+
 #### Donations by Year ####
 donations_by_year_ontario <- ontario_data %>%
   group_by(donation_year) %>%

### 2024-12-15 scripts/06-model_data.R

@@ -106,6 +106,98 @@ total_donations_cl <- total_donations_cl %>%
 model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)
 summary(model_cl)
 
+
+### Regressions Using Real Amount ###
+# Filter data for Ontario
+ontario_data2 <- analysis_data %>% 
+  filter(region == "Ontario")
+
+# Aggregate total donations per party per year in Ontario
+total_donations_ontario2 <- ontario_data2 %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(real_amount, na.rm = TRUE),
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Log-transform the dependent variable to stabilize variance
+total_donations_ontario2 <- total_donations_ontario2 %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
+
+# Run Regression for Ontario with Controls
+model_ontario2 <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario2)
+summary(model_ontario2)
+
+#### Model for Federal ####
+
+# Filter data for Federal
+federal_data2 <- analysis_data %>% 
+  filter(region == "Federal")
+
+# Aggregate total donations per party per year at Federal level
+total_donations_federal2 <- federal_data2 %>%
+  group_by(political_party, donation_year) %>%
+  summarize(
+    Total_Donations = sum(real_amount, na.rm = TRUE),
+    In_Power = first(recipient_in_power),
+    Party_Size = first(Party_Size),
+    Election_Year = first(Election_Year)
+  ) %>%
+  ungroup()
+
+# Log-transform the dependent variable to stabilize variance
+total_donations_federal2 <- total_donations_federal2 %>%
+  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)
+
+# Run Regression for Federal with Controls
+model_federal2 <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal2)
+summary(model_federal2)
+
+#### Conservative and Liberal Analysis ####
+
+# Filter for Conservative and Liberal parties
[…]

### 2024-12-15 paper/paper.qmd
rendered paper

@@
 then aggregated at the party-year level to analyze total annual
-contributions received by each party. Binary variables, including
+contributions received by each party.
+
+To account for inflation and ensure comparability over time in the modeling process, the amount donated was adjusted using the Consumer Price Index (CPI), with 2024 set as the base year.
+This adjustment facilitated accurate regression analysis by normalizing donation amounts across different years.
+For the analysis section, which focuses on presenting either total or average donations, the unadjusted donation amounts were retained to provide a straightforward depiction of contribution figures.
+
+Binary variables, including
 Recipient_in_Power, Party_Size, and Election_Year, were created to

### 2024-12-15 scripts/06-model_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Generates regression for donation amount and power status
 # Author: Maria Mangru
-# Date: 12 December 2024
+# Date: 14 December 2024
 # Contact: maria.mangru@mail.utoronto.ca
 # License: MIT
 
@@ -16,7 +16,6 @@ analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
 
 #### Model for Ontario ####
-
 # Filter data for Ontario
 ontario_data <- analysis_data %>% 
   filter(region == "Ontario")
@@ -25,7 +24,7 @@ ontario_data <- analysis_data %>%
 total_donations_ontario <- ontario_data %>%
   group_by(political_party, donation_year) %>%
   summarize(
-    Total_Donations = sum(amount, na.rm = TRUE),
+    Total_Donations = sum(real_amount, na.rm = TRUE),
     In_Power = first(recipient_in_power),
     Party_Size = first(Party_Size),
     Election_Year = first(Election_Year)
@@ -38,10 +37,10 @@ total_donations_ontario <- total_donations_ontario %>%
 
 # Run Regression for Ontario with Controls
 model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)
-summary(model_ontario)
+summary(model_ontario2)
+
 
 #### Model for Federal ####
-
 # Filter data for Federal
 federal_data <- analysis_data %>% 
   filter(region == "Federal")
@@ -50,7 +49,7 @@ federal_data <- analysis_data %>%
 total_donations_federal <- federal_data %>%
   group_by(political_party, donation_year) %>%
   summarize(
-    Total_Donations = sum(amount, na.rm = TRUE),
+    Total_Donations = sum(real_amount, na.rm = TRUE),
     In_Power = first(recipient_in_power),
     Party_Size = first(Party_Size),
     Election_Year = first(Election_Year)
@@ -63,7 +62,7 @@ total_donations_federal <- total_donations_federal %>%
 
 # Run Regression for Federal with Controls
 model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)
-summary(model_federal)
+summary(model_federal2)
 
 #### Conservative and Liberal Analysis ####
 
@@ -91,7 +90,7 @@ cl_data <- cl_data %>%
 total_donations_cl <- cl_data %>%
[…]

### 2024-12-19 other/datasheet/datasheet.qmd
Updated datasheet

@@
 1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
-    - The dataset was created to enable analysis of Australian politicians.
-We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
+    - The dataset was created to analyze political donations in Canada, specifically to investigate whether the status of a political party, being in power or not, affects the level of donations it receives.
 2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
